### 2021-08-16 publications/dora_outlier_detection_iaai.tex

\def\year{2021}\relax
%File: formatting-instructions-latex-2021.tex
%release 2021.2
\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
\usepackage{aaai21}  % DO NOT CHANGE THIS
\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet} % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\usepackage{graphicx} % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\usepackage{natbib}  % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\usepackage{caption} % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\frenchspacing  % DO NOT CHANGE THIS
\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS
%\nocopyright
%PDF Info Is REQUIRED.
% For /Author, add all authors within the parentheses, separated by commas. No accents or commands.
% For /Title, add Title in Mixed Case. No accents or commands. Retain the parentheses.
\pdfinfo{
/Title (AAAI Press Formatting Instructions for Authors Using LaTeX -- A Guide)
/Author (AAAI Press Staff, Pater Patel Schneider, Sunil Issar, J. Scott Penberthy, George Ferguson, Hans Guesgen, Francisco Cruz, Marc Pujol-Gonzalez)
/TemplateVersion (2021.2)
} %Leave this
% /Title ()
% Put your actual complete title (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case
% Leave the space between \Title and the beginning parenthesis alone
% /Author ()
% Put your actual complete list of authors (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case.
% Each author should be only by a comma. If the name contains accents, remove them. If there are any LaTeX commands,
% remove them.

% DISALLOWED PACKAGES
% \usepackage{authblk} -- This package is specifically forbidden
% \usepackage{balance} -- This package is specifically forbidden
% \usepackage{color (if used in text)
% \usepackage{CJK} -- This package is specifically forbidden
% \usepackage{float} -- This package is specifically forbidden
% \usepackage{flushend} -- This package is specifically forbidden
% \usepackage{fontenc} -- This package is specifically forbidden
% \usepackage{fullpage} -- This package is specifically forbidden
% \usepackage{geometry} -- This package is specifically forbidden
% \usepackage{grffile} -- This package is specifically forbidden
% \usepackage{hyperref} -- This package is specifically forbidden
% \usepackage{navigator} -- This package is specifically forbidden
% (or any other package that embeds links such as navigator or hyperref)
% \indentfirst} -- This package is specifically forbidden
% \layout} -- This package is specifically forbidden
% \multicol} -- This package is specifically forbidden
% \nameref} -- This package is specifically forbidden
% \usepackage{savetrees} -- This package is specifically forbidden
% \usepackage{setspace} -- This package is specifically forbidden
% \usepackage{stfloats} -- This package is specifically forbidden
% \usepackage{tabu} -- This package is specifically forbidden
% \usepackage{titlesec} -- This package is specifically forbidden
% \usepackage{tocbibind} -- This package is specifically forbidden
% \usepackage{ulem} -- This package is specifically forbidden
% \usepackage{wrapfig} -- This package is specifically forbidden
% DISALLOWED COMMANDS
% \nocopyright -- Your paper will not be published if you use this command
% \addtolength -- This command may not be used
% \balance -- This command may not be used
% \baselinestretch -- Your paper will not be published if you use this command
% \clearpage -- No page breaks of any kind may be used for the final version of your paper
% \columnsep -- This command may not be used
% \newpage -- No page breaks of any kind may be used for the final version of your paper
% \pagebreak -- No page breaks of any kind may be used for the final version of your paperr
% \pagestyle -- This command may not be used
% \tiny -- This is not an acceptable font size.
% \vspace{- -- No negative value may be used in proximity of a caption, figure, table, section, subsection, subsubsection, or reference
% \vskip{- -- No negative value may be used to alter spacing above or below a caption, figure, table, section, subsection, subsubsection, or reference

\setcounter{secnumdepth}{0} %May be changed to 1 or 2 if section numbers are desired.

% The file aaai21.sty is the style file for AAAI Press
% proceedings, working notes, and technical reports.
%

% Title

% Your title must be in mixed case, not sentence case.
% That means all verbs (including short verbs like be, is, using,and go),
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\iffalse
\title{AAAI Press Formatting Instructions \\for Authors Using \LaTeX{} --- A Guide }
\author{
    %Authors
    % All authors must be in the same font size and format.
    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
    AAAI Style Contributions by Pater Patel Schneider,
    Sunil Issar,  \\
    J. Scott Penberthy,
    George Ferguson,
    Hans Guesgen,
    Francisco Cruz,
    Marc Pujol-Gonzalez
    \\
}
\affiliations{
    %Afiliations
    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
    %If you have multiple authors and multiple affiliations
    % use superscripts in text and roman font to identify them.
    %For example,

    % Sunil Issar, \textsuperscript{\rm 2}
    % J. Scott Penberthy, \textsuperscript{\rm 3}
    % George Ferguson,\textsuperscript{\rm 4}
    % Hans Guesgen, \textsuperscript{\rm 5}.
    % Note that the comma should be placed BEFORE the superscript for optimum readability

    2275 East Bayshore Road, Suite 160\\
    Palo Alto, California 94303\\
    % email address must be in roman text type, not monospace or sans serif
    publications21@aaai.org

    % See more examples next
}
\fi
\iffalse
%Example, Single Author, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it
\title{My Publication Title --- Single Author}
\author {
    % Author
    Author Name \\
}

\affiliations{
    Affiliation \\
    Affiliation Line 2 \\
    name@example.com
}
\fi

%Example, Multiple Authors, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it
\title{Domain-agnostic Outlier Ranking Algorithms (DORA)---A Configurable Pipeline for Facilitating Outlier Detection in Scientific Datasets}
\author {
    % Authors
    First Author Name,\textsuperscript{\rm 1}
    Second Author Name, \textsuperscript{\rm 2}
    Third Author Name \textsuperscript{\rm 1} \\
}
\affiliations {
    % Affiliations
    \textsuperscript{\rm 1} Affiliation 1 \\
    \textsuperscript{\rm 2} Affiliation 2 \\
    firstAuthor@affiliation1.com, secondAuthor@affilation2.com, thirdAuthor@affiliation1.com
}

\begin{document}

\maketitle

\begin{abstract}
AAAI creates proceedings, working notes, and technical reports directly from electronic source furnished by the authors. To ensure that all papers in the publication have a uniform appearance, authors must adhere to the following instructions.
\end{abstract}

\section{Introduction}
\begin{itemize}
\item Framing of outlier detection as a domain-agnostic problem that is well-suited to cross-disciplinary applications
\item Example applications of outlier detection across science application domains
\item Challenges/barriers to entry for domain scientists who could benefit from outlier detection 
\item DORA as a solution to lower these barriers to entry and help integrate outlier detection methods into existing scientific workflows across variety of domains
\end{itemize}

\section{Related Work}
\begin{itemize}
\item Methods for outlier detection
\item Applications of outlier detection related to our focus domains (e.g., \citet{kerner2020comparison})
\item Tools for applying outlier detection or other methods across domains?

\end{itemize}

\section{Methods}

\begin{itemize}
\item DORA configurable pipeline approach 
\item data types
\item feature extraction
\item ranking algorithms
\item results modules
\end{itemize}

\section{Experiments}

\subsection{Datasets}

Four use case datasets; additional benchmark datasets (e.g. FashionMNIST/MNIST)?

\subsection{Experiment setup}

Experiments for each of the use cases; evaluation criteria

\section{Results}

Experiment results

\section{Discussion}
\begin{itemize}
\item Discuss results from experiments
\item Challenges of evaluating results in outlier detection?
\item Downstream use of results in science applications
\end{itemize}

\section{Conclusions}
Summarize conclusions and plans for future work

Path to deployment

\bibliography{dora_references}

\section{ Acknowledgments}
JPL CIF, SMD etc.

\end{document}

### 2021-08-17 dora_exp_pipeline/conv_pae_outlier_detection.py
Expose training params; use random seed for model training; update fmnist benchmark sample data config file

@@ -31,7 +31,8 @@ class ConvPAEOutlierDetection(OutlierDetection):
         super(ConvPAEOutlierDetection, self).__init__('conv_pae')
 
     def _rank_internal(self, data_to_fit, data_to_score, data_to_score_ids,
-                       top_n, seed, latent_dim):
+                       top_n, seed, latent_dim, max_epochs=1000, patience=10,
+                       val_split=0.25, verbose=0):
         if data_to_fit is None:
             data_to_fit = deepcopy(data_to_score)
 
@@ -53,9 +54,16 @@ class ConvPAEOutlierDetection(OutlierDetection):
                                f'must be <= number of features '
                                f'({num_features})')
 
+        # Set tensorflow logging level
+        os.environ['TF_CPP_MIN_LOG_LEVEL'] = '3' if verbose == 0 else '0'
+
+        # Set seed
+        tf.random.set_seed(seed)
+
         # Rank targets
         scores = train_and_run_conv_PAE(data_to_fit, data_to_score, latent_dim,
-                                        image_shape, seed)
+                                        image_shape, seed, max_epochs, patience,
+                                        val_split, verbose)
         selection_indices = np.argsort(scores)[::-1]
 
         results = dict()
@@ -70,17 +78,19 @@ class ConvPAEOutlierDetection(OutlierDetection):
         return results
 
 
-def train_and_run_conv_PAE(train, test, latent_dim, image_shape, seed):
+def train_and_run_conv_PAE(train, test, latent_dim, image_shape, seed, 
+                           max_epochs, patience, val_split, verbose):
     # Make tensorflow datasets
     channels = image_shape[2]
-    train_ds, val_ds, test_ds = get_train_val_test(train, test, seed, channels)
+    train_ds, val_ds, test_ds = get_train_val_test(train, test, seed, channels, 
+                                                   val_split)
 
     # Train autoencoder
     autoencoder = ConvAutoencoder(latent_dim, image_shape)
     autoencoder.compile(optimizer='adam', loss=losses.MeanSquaredError())
-    callback = EarlyStopping(monitor='val_loss', patience=5)
-    autoencoder.fit(x=train_ds, validation_data=val_ds, verbose=0, epochs=1000,
-                    callbacks=[callback])
+    callback = EarlyStopping(monitor='val_loss', patience=patience)
+    autoencoder.fit(x=train_ds, validation_data=val_ds, verbose=verbose, 
[…]

### 2021-08-17 dora_exp_pipeline/dora_exp.py

@@ -78,9 +78,6 @@ def start(config_file: str, out_dir: str, log_file=None, seed=1234):
     if log_file is not None:
         logger = LogUtil('dora_exp', log_file)
 
-    # Supress Tensorflow progress output
-    os.environ['TF_CPP_MIN_LOG_LEVEL'] = '3'
-
     config = DoraConfig(config_file, logger)
     if out_dir is not None:
         config.out_dir = out_dir

### 2021-08-17 dora_exp_pipeline/pae_outlier_detection.py

@@ -24,7 +24,8 @@ class PAEOutlierDetection(OutlierDetection):
         super(PAEOutlierDetection, self).__init__('pae')
 
     def _rank_internal(self, data_to_fit, data_to_score, data_to_score_ids,
-                       top_n, seed, latent_dim):
+                       top_n, seed, latent_dim, max_epochs=500, patience=3, 
+                       val_split=0.25, verbose=0):
         if data_to_fit is None:
             data_to_fit = deepcopy(data_to_score)
 
@@ -40,9 +41,16 @@ class PAEOutlierDetection(OutlierDetection):
                                f'must be <= number of features '
                                f'({num_features})')
 
+        # Set tensorflow logging level
+        os.environ['TF_CPP_MIN_LOG_LEVEL'] = '3' if verbose == 0 else '0'
+
+        # Set seed
+        tf.random.set_seed(seed)
+
         # Rank targets
         scores = train_and_run_PAE(data_to_fit, data_to_score, latent_dim,
-                                   num_features)
+                                   num_features, seed, max_epochs, patience, 
+                                   val_split, verbose)
         selection_indices = np.argsort(scores)[::-1]
 
         results = dict()
@@ -57,21 +65,23 @@ class PAEOutlierDetection(OutlierDetection):
         return results
 
 
-def train_and_run_PAE(train, test, latent_dim, num_features):
+def train_and_run_PAE(train, test, latent_dim, num_features, seed, max_epochs, 
+                      patience, val_split, verbose):
     # Train autoencoder
     autoencoder = Autoencoder(latent_dim, num_features)
     autoencoder.compile(optimizer='adam', loss=losses.MeanSquaredError())
-    callback = EarlyStopping(monitor='val_loss', patience=3)
-    autoencoder.fit(train, train, epochs=500, verbose=0, callbacks=[callback],
-                    validation_split=0.25)
+    callback = EarlyStopping(monitor='val_loss', patience=patience)
+    autoencoder.fit(train, train, epochs=max_epochs, verbose=verbose, 
+                    callbacks=[callback], validation_split=val_split)
 
     # Train flow
     encoded_train = autoencoder.encoder(train).numpy()
     flow = NormalizingFlow(latent_dim)
     flow.compile(optimizer='adam', loss=lambda y, rv_y: -rv_y.log_prob(y))
-    callback = EarlyStopping(monitor='val_loss', patience=3)
[…]

### 2021-08-17 dora_exp_pipeline/conv_pae_outlier_detection.py
Fix lint errors

@@ -78,18 +78,18 @@ class ConvPAEOutlierDetection(OutlierDetection):
         return results
 
 
-def train_and_run_conv_PAE(train, test, latent_dim, image_shape, seed, 
+def train_and_run_conv_PAE(train, test, latent_dim, image_shape, seed,
                            max_epochs, patience, val_split, verbose):
     # Make tensorflow datasets
     channels = image_shape[2]
-    train_ds, val_ds, test_ds = get_train_val_test(train, test, seed, channels, 
+    train_ds, val_ds, test_ds = get_train_val_test(train, test, seed, channels,
                                                    val_split)
 
     # Train autoencoder
     autoencoder = ConvAutoencoder(latent_dim, image_shape)
     autoencoder.compile(optimizer='adam', loss=losses.MeanSquaredError())
     callback = EarlyStopping(monitor='val_loss', patience=patience)
-    autoencoder.fit(x=train_ds, validation_data=val_ds, verbose=verbose, 
+    autoencoder.fit(x=train_ds, validation_data=val_ds, verbose=verbose,
                     epochs=max_epochs, callbacks=[callback])
 
     # Encode datasets
@@ -101,8 +101,8 @@ def train_and_run_conv_PAE(train, test, latent_dim, image_shape, seed,
     flow = NormalizingFlow(latent_dim)
     flow.compile(optimizer='adam', loss=lambda y, rv_y: -rv_y.log_prob(y))
     callback = EarlyStopping(monitor='val_loss', patience=patience)
-    flow.fit(np.zeros((len(encoded_train), 0)), encoded_train, 
-             verbose=verbose, epochs=max_epochs, callbacks=[callback], 
+    flow.fit(np.zeros((len(encoded_train), 0)), encoded_train,
+             verbose=verbose, epochs=max_epochs, callbacks=[callback],
              validation_split=val_split)
 
     # Calculate scores
@@ -243,7 +243,7 @@ class ConvAutoencoder(Model):
                 kernel_size=3,
                 strides=1,
                 padding='same'),
-            layers.experimental.preprocessing.Rescaling(255) 
+            layers.experimental.preprocessing.Rescaling(255)
         ])
 
         self.encoder = keras.Sequential(self._encoder_layers)

### 2021-08-17 dora_exp_pipeline/pae_outlier_detection.py

@@ -9,6 +9,7 @@
 # Date created: August 16, 2021
 
 from dora_exp_pipeline.outlier_detection import OutlierDetection
+import os
 import numpy as np
 from copy import deepcopy
 import tensorflow as tf
@@ -24,7 +25,7 @@ class PAEOutlierDetection(OutlierDetection):
         super(PAEOutlierDetection, self).__init__('pae')
 
     def _rank_internal(self, data_to_fit, data_to_score, data_to_score_ids,
-                       top_n, seed, latent_dim, max_epochs=500, patience=3, 
+                       top_n, seed, latent_dim, max_epochs=500, patience=3,
                        val_split=0.25, verbose=0):
         if data_to_fit is None:
             data_to_fit = deepcopy(data_to_score)
@@ -49,7 +50,7 @@ class PAEOutlierDetection(OutlierDetection):
 
         # Rank targets
         scores = train_and_run_PAE(data_to_fit, data_to_score, latent_dim,
-                                   num_features, seed, max_epochs, patience, 
+                                   num_features, seed, max_epochs, patience,
                                    val_split, verbose)
         selection_indices = np.argsort(scores)[::-1]
 
@@ -65,13 +66,13 @@ class PAEOutlierDetection(OutlierDetection):
         return results
 
 
-def train_and_run_PAE(train, test, latent_dim, num_features, seed, max_epochs, 
+def train_and_run_PAE(train, test, latent_dim, num_features, seed, max_epochs,
                       patience, val_split, verbose):
     # Train autoencoder
     autoencoder = Autoencoder(latent_dim, num_features)
     autoencoder.compile(optimizer='adam', loss=losses.MeanSquaredError())
     callback = EarlyStopping(monitor='val_loss', patience=patience)
-    autoencoder.fit(train, train, epochs=max_epochs, verbose=verbose, 
+    autoencoder.fit(train, train, epochs=max_epochs, verbose=verbose,
                     callbacks=[callback], validation_split=val_split)
 
     # Train flow
@@ -79,8 +80,8 @@ def train_and_run_PAE(train, test, latent_dim, num_features, seed, max_epochs,
     flow = NormalizingFlow(latent_dim)
     flow.compile(optimizer='adam', loss=lambda y, rv_y: -rv_y.log_prob(y))
     callback = EarlyStopping(monitor='val_loss', patience=patience)
-    flow.fit(np.zeros((len(encoded_train), 0)), encoded_train, 
[…]

### 2021-08-17 dora_exp_pipeline/conv_pae_outlier_detection.py
Expose optimizer parameter

@@ -32,7 +32,7 @@ class ConvPAEOutlierDetection(OutlierDetection):
 
     def _rank_internal(self, data_to_fit, data_to_score, data_to_score_ids,
                        top_n, seed, latent_dim, max_epochs=1000, patience=10,
-                       val_split=0.25, verbose=0):
+                       val_split=0.25, verbose=0, optimizer='adam'):
         if data_to_fit is None:
             data_to_fit = deepcopy(data_to_score)
 
@@ -63,7 +63,7 @@ class ConvPAEOutlierDetection(OutlierDetection):
         # Rank targets
         scores = train_and_run_conv_PAE(data_to_fit, data_to_score, latent_dim,
                                         image_shape, seed, max_epochs, patience,
-                                        val_split, verbose)
+                                        val_split, verbose, optimizer)
         selection_indices = np.argsort(scores)[::-1]
 
         results = dict()
@@ -79,7 +79,7 @@ class ConvPAEOutlierDetection(OutlierDetection):
 
 
 def train_and_run_conv_PAE(train, test, latent_dim, image_shape, seed,
-                           max_epochs, patience, val_split, verbose):
+                           max_epochs, patience, val_split, verbose, optimizer):
     # Make tensorflow datasets
     channels = image_shape[2]
     train_ds, val_ds, test_ds = get_train_val_test(train, test, seed, channels,
@@ -87,7 +87,7 @@ def train_and_run_conv_PAE(train, test, latent_dim, image_shape, seed,
 
     # Train autoencoder
     autoencoder = ConvAutoencoder(latent_dim, image_shape)
-    autoencoder.compile(optimizer='adam', loss=losses.MeanSquaredError())
+    autoencoder.compile(optimizer=optimizer, loss=losses.MeanSquaredError())
     callback = EarlyStopping(monitor='val_loss', patience=patience)
     autoencoder.fit(x=train_ds, validation_data=val_ds, verbose=verbose,
                     epochs=max_epochs, callbacks=[callback])
@@ -99,7 +99,7 @@ def train_and_run_conv_PAE(train, test, latent_dim, image_shape, seed,
 
     # Train flow
     flow = NormalizingFlow(latent_dim)
-    flow.compile(optimizer='adam', loss=lambda y, rv_y: -rv_y.log_prob(y))
+    flow.compile(optimizer=optimizer, loss=lambda y, rv_y: -rv_y.log_prob(y))
[…]

### 2021-08-17 dora_exp_pipeline/pae_outlier_detection.py

@@ -26,7 +26,7 @@ class PAEOutlierDetection(OutlierDetection):
 
     def _rank_internal(self, data_to_fit, data_to_score, data_to_score_ids,
                        top_n, seed, latent_dim, max_epochs=500, patience=3,
-                       val_split=0.25, verbose=0):
+                       val_split=0.25, verbose=0, optimizer='adam'):
         if data_to_fit is None:
             data_to_fit = deepcopy(data_to_score)
 
@@ -51,7 +51,7 @@ class PAEOutlierDetection(OutlierDetection):
         # Rank targets
         scores = train_and_run_PAE(data_to_fit, data_to_score, latent_dim,
                                    num_features, seed, max_epochs, patience,
-                                   val_split, verbose)
+                                   val_split, verbose, optimizer)
         selection_indices = np.argsort(scores)[::-1]
 
         results = dict()
@@ -67,10 +67,10 @@ class PAEOutlierDetection(OutlierDetection):
 
 
 def train_and_run_PAE(train, test, latent_dim, num_features, seed, max_epochs,
-                      patience, val_split, verbose):
+                      patience, val_split, verbose, optimizer):
     # Train autoencoder
     autoencoder = Autoencoder(latent_dim, num_features)
-    autoencoder.compile(optimizer='adam', loss=losses.MeanSquaredError())
+    autoencoder.compile(optimizer=optimizer, loss=losses.MeanSquaredError())
     callback = EarlyStopping(monitor='val_loss', patience=patience)
     autoencoder.fit(train, train, epochs=max_epochs, verbose=verbose,
                     callbacks=[callback], validation_split=val_split)
@@ -78,7 +78,7 @@ def train_and_run_PAE(train, test, latent_dim, num_features, seed, max_epochs,
     # Train flow
     encoded_train = autoencoder.encoder(train).numpy()
     flow = NormalizingFlow(latent_dim)
-    flow.compile(optimizer='adam', loss=lambda y, rv_y: -rv_y.log_prob(y))
+    flow.compile(optimizer=optimizer, loss=lambda y, rv_y: -rv_y.log_prob(y))
     callback = EarlyStopping(monitor='val_loss', patience=patience)
     flow.fit(np.zeros((len(encoded_train), 0)), encoded_train,
              epochs=max_epochs, verbose=verbose, callbacks=[callback],

### 2021-08-18 dora_exp_pipeline/conv_pae_outlier_detection.py
Update versions; add tqdm training progress bar

@@ -19,6 +19,7 @@ from PIL import Image
 from itertools import accumulate
 from copy import deepcopy
 import tensorflow as tf
+import tensorflow_addons as tfa
 from tensorflow import keras
 from tensorflow.keras import layers, losses
 from tensorflow.keras.models import Model
@@ -85,12 +86,17 @@ def train_and_run_conv_PAE(train, test, latent_dim, image_shape, seed,
     train_ds, val_ds, test_ds = get_train_val_test(train, test, seed, channels,
                                                    val_split)
 
+    # Initialize callbacks
+    early_stop_callback = EarlyStopping(monitor='val_loss', patience=patience)
+    tqdm_callback = tfa.callbacks.TQDMProgressBar(leave_overall_progress=False,
+                                                  show_epoch_progress=False)
+    callbacks = [early_stop_callback, tqdm_callback]
+
     # Train autoencoder
     autoencoder = ConvAutoencoder(latent_dim, image_shape)
     autoencoder.compile(optimizer=optimizer, loss=losses.MeanSquaredError())
-    callback = EarlyStopping(monitor='val_loss', patience=patience)
     autoencoder.fit(x=train_ds, validation_data=val_ds, verbose=verbose,
-                    epochs=max_epochs, callbacks=[callback])
+                    epochs=max_epochs, callbacks=callbacks)
 
     # Encode datasets
     enc_train = autoencoder.encoder.predict(train_ds)
@@ -100,9 +106,8 @@ def train_and_run_conv_PAE(train, test, latent_dim, image_shape, seed,
     # Train flow
     flow = NormalizingFlow(latent_dim)
     flow.compile(optimizer=optimizer, loss=lambda y, rv_y: -rv_y.log_prob(y))
-    callback = EarlyStopping(monitor='val_loss', patience=patience)
     flow.fit(np.zeros((len(encoded_train), 0)), encoded_train,
-             verbose=verbose, epochs=max_epochs, callbacks=[callback],
+             verbose=verbose, epochs=max_epochs, callbacks=callbacks,
              validation_split=val_split)
 
     # Calculate scores

### 2021-08-18 setup.py

@@ -22,8 +22,9 @@ setup(
         'pandas==1.1.5',
         'rasterio==1.2.6',
         'tqdm==4.62.0',
-        'tensorflow==2.4.2',
-        'tensorflow-probability==0.12.2'
+        'tensorflow==2.5.1',
+        'tensorflow-probability==0.13.0',
+        'tensorflow-addons==0.13.0'
     ],
     provide=[
         'dora_exp_pipeline'

### 2021-08-30 dora_exp_pipeline/version.py
update version #59

@@ -1,4 +1,4 @@
-__version__ = '0.0.7'
+__version__ = '0.0.8'
 
 
 # Copyright (c) 2021 California Institute of Technology ("Caltech").

### 2021-08-30 publications/dora_outlier_detection_iaai.tex

@@
 \begin{abstract}
-AAAI creates proceedings, working notes, and technical reports directly from electronic source furnished by the authors.
-To ensure that all papers in the publication have a uniform appearance, authors must adhere to the following instructions.
+Automatic detection of out-of-distribution (OOD) samples or features is
+universally needed when working with scientific datasets. For example,
+OOD detection methods can be used for cleaning datasets or flagging
+novel samples in order to guide instrument acquisition or scientific analysis.
+We present Domain-agnostic Outlier Ranking Algorithms (DORA) to provide
+a configurable pipeline that makes it easy for scientists to apply and evaluate
+outlier detection methods in a variety of domains. DORA allows users to 
+configure outlier detection experiments by specifying the location of their 
+dataset(s), the input data type, feature extraction methods, and which 
+algorithms should be applied. DORA supports image, raster, time series, 
+or feature vector input data types and outlier detection methods that include
+ Isolation Forest, Discovery via Eigenbasis Modeling of Uninteresting Data 
+ (DEMUD), principal component analysis (PCA), Reed-Xiaoli (RX) detector, 
+ Local RX, negative sampling, and probabilistic autoencoder. Given a set of
+  data samples, each algorithm assigns an outlier score to each sample. DORA
+   provides results organization and visualization modules to help users 
+   process the results, including sorting samples by outlier score, evaluating 
+   outlier recall for a set of known/labeled outliers, clustering groups of 
+   similar outliers together, and causal graphs. To demonstrate the 
+   cross-domain applicability of DORA, we conducted outlier detection
+    experiments for four use cases: 1) cleaning noisy ground-truth datasets
+     (Earth Science), 2) detecting volcanic thermal anomalies (Earth Science), 
+     3) flagging non-astrophysical objects in the Dark Energy Survey 
+     (astrophysics), and 4) novelty-guided onboard targeting of instruments
+      for Mars rovers (planetary science). [We showed that ...]
 \end{abstract}
@@
 \begin{itemize}
-\item Framing of outlier detection as a domain-agnostic problem that is well-suited to cross-disciplinary applications
-\item Example applications of outlier detection across science application domains
-\item Challenges/barriers to entry for domain scientists who could benefit from outlier detection 
-\item DORA as a solution to lower these barriers to entry and help integrate outlier detection methods into existing scientific workflows across variety of domains
+\item Framing of outlier detection as a domain-agnostic problem that is well-
+suited to cross-disciplinary applications
+\item Example applications of outlier detection across science application 
+domains
+\item Challenges/barriers to entry for domain scientists who could benefit from
+ outlier detection 
+\item DORA as a solution to lower these barriers to entry and help integrate 
+outlier detection methods into existing scientific workflows across variety of 
+domains
 \end{itemize}
@@
 \item Methods for outlier detection
-\item Applications of outlier detection related to our focus domains (e.g., \citet{kerner2020comparison})
+\item Applications of outlier detection related to our focus domains 
+(e.g., \citet{kerner2020comparison})
 \item Tools for applying outlier detection or other methods across domains?
@@
 
-\begin{itemize}
-\item DORA configurable pipeline approach 
-\item data types
-\item feature extraction
-\item ranking algorithms
-\item results modules
-\end{itemize}
+\subsection{DORA configurable pipeline} Steven please add details about the 
+approach/design choices for this pipeline. Mention that additional data types, 
+feature extraction methods, ranking algorithms, and results organization and 
+visualization modules can easily be added and integrated if they follow the API. 
+Hannah to add architecture figure. Add descriptions of each layer of modules: 
+\paragraph{Data types} Hannah
+\paragraph{Feature extraction} Hannah 
+\paragraph{Ranking algorithms} We implemented 7 algorithms for scoring and 
+ranking samples by outlierness. [Describe each algorithm briefly and how they 
+complement each other.]
+\paragraph{Results organization} Eric, Umaa, Steven, Vinay
+\paragraph{Outlier visualization} Jake?
 
@@
 \subsection{Datasets}
+Describe the datasets overall and our motivation for choosing these. Maybe add
+ a table with number of samples, number of reference known outliers (if 
+ available), etc, if room?
 
-Four use case datasets; additional benchmark datasets (e.g. FashionMNIST/MNIST)?
+\subsubsection{Astrophysical objects in Dark Energy Survey}
+Umaa
 
-\subsection{Experiment setup}
+\subsubsection{Targets in Mars rover images}
+Kiri
 
-Experiments for each of the use cases; evaluation criteria
+\subsubsection{Satellite time series for ground-truth observations}
+Hannah: Describe the Earth crop field data set and applications
 
-\section{Results}
+\subsubsection{Volcanic thermal anomalies}
 
-Experiment results
+\subsubsection{MNIST and FashionMNIST}
+We added this to have a traditional benchmark and use case demonstration.
+
+\subsection{Results}
+Discuss the overall approach to the setup of the experiments. In each 
+subsection discuss specific setup details (including evaluation metrics) for 
+each dataset then the results from the experiments.
+
+\subsubsection{Astrophysical objects in Dark Energy Survey}
+Umaa
+
+\subsubsection{Targets in Mars rover images}
+Kiri
+
+\subsubsection{Satellite time series for ground-truth observations}
+Hannah: Describe the Earth crop field data set and applications
+
+\subsubsection{Volcanic thermal anomalies}
+
+\subsubsection{MNIST and FashionMNIST}
+We added this to have a traditional benchmark and use case demonstration.
 
@@
 \item Challenges of evaluating results in outlier detection?
-\item Downstream use of results in science applications
+\item Downstream use of results in science applications/Path to deployment
 \end{itemize}
@@
 \section{Conclusions}
-Summarize conclusions and plans for future work
+Summarize conclusions and plans for future work, including public 
+pip-installable package
 
-Path to deployment
+[Note: these datasets could also be used by other researchers as OOD
+ benchmark tasks?]
 
@@
 \section{ Acknowledgments}
-JPL CIF, SMD etc.
+Funding/support: JPL CIF, SMD etc. Acknowledge people who contributed to 
+datasets e.g. FAO, Matt Pritchard, DES, other CIF team members?
 

### 2021-08-30 test/test_planetary.py
Update pytests to specify number of trees in iForest. (#59,#14)

@@ -31,7 +31,7 @@ class TestPlanetary(TestCase):
 
         # Check results for each algorithm
         for file_base in ['demud-k=5/selections-demud.csv',
-                          'iforest/selections-iforest.csv',
+                          'iforest-n_trees=100/selections-iforest.csv',
                           'lrx-inner_window=3-outer_window=9-bands=1/selections-lrx.csv',
                           'negative_sampling-percent_increase=20/selections-negative_sampling.csv',
                           'pca-k=5/selections-pca.csv'

### 2021-08-30 test/test_planetary.py
Add 'random' alg to functional pytest. (#14)

@@ -34,7 +34,8 @@ class TestPlanetary(TestCase):
                           'iforest-n_trees=100/selections-iforest.csv',
                           'lrx-inner_window=3-outer_window=9-bands=1/selections-lrx.csv',
                           'negative_sampling-percent_increase=20/selections-negative_sampling.csv',
-                          'pca-k=5/selections-pca.csv'
+                          'pca-k=5/selections-pca.csv',
+                          'random/selections-random.csv'
                           ]:
             correct_file = 'test/ref-results/planetary_rover/functional/%s' % file_base
             output_file = '%s/%s' % (self.outdir, file_base)

### 2021-08-30 dora_exp_pipeline/version.py
add SOM clustering result organization method #38

@@ -1,4 +1,4 @@
-__version__ = '0.0.8'
+__version__ = '0.0.9'
 
 
 # Copyright (c) 2021 California Institute of Technology ("Caltech").

### 2021-09-01 dora_exp_pipeline/conv_pae_outlier_detection.py
commit before merging files

@@ -12,6 +12,10 @@
 # Date created: August 16, 2021
 
 from dora_exp_pipeline.outlier_detection import OutlierDetection
+from dora_exp_pipeline.pae_outlier_detection import (
+    NormalizingFlow,
+    make_tf_callbacks
+)
 import os
 import math
 import numpy as np
@@ -23,8 +27,7 @@ import tensorflow_addons as tfa
 from tensorflow import keras
 from tensorflow.keras import layers, losses
 from tensorflow.keras.models import Model
-from tensorflow.keras.callbacks import EarlyStopping
-from tensorflow_probability import distributions, bijectors, layers as tfpl
+from tensorflow.keras.callbacks import EarlyStopping 
 
 
 class ConvPAEOutlierDetection(OutlierDetection):
@@ -33,7 +36,8 @@ class ConvPAEOutlierDetection(OutlierDetection):
 
     def _rank_internal(self, data_to_fit, data_to_score, data_to_score_ids,
                        top_n, seed, latent_dim, max_epochs=1000, patience=10,
-                       val_split=0.25, verbose=0, optimizer='adam'):
+                       val_split=0.25, verbose=0, optimizer='adam', 
+                       log_dir=None):
         if data_to_fit is None:
             data_to_fit = deepcopy(data_to_score)
 
@@ -272,25 +276,6 @@ class ConvAutoencoder(Model):
         encoded = self.encoder(x)
         decoded = self.decoder(encoded)
         return decoded
-
-
-class NormalizingFlow(Model):
-    def __init__(self, latent_dim):
-        super(NormalizingFlow, self).__init__()
-
-        self.dist = keras.Sequential([
-            layers.InputLayer(input_shape=(0,), dtype=tf.float32),
-            tfpl.DistributionLambda(
-                lambda t: distributions.MultivariateNormalDiag(
-                    loc=tf.zeros(tf.concat([tf.shape(t)[:-1],
-                                           [latent_dim]],
-                                           axis=0)))),
-            tfpl.AutoregressiveTransform(bijectors.AutoregressiveNetwork(
-                params=2, hidden_units=[10, 10], activation='relu')),
-        ])
-
-    def call(self, x):
-        return self.dist(x)
 
 
 # Copyright (c) 2021 California Institute of Technology ("Caltech").

### 2021-09-01 dora_exp_pipeline/dora_exp.py

@@ -93,6 +93,10 @@ def start(config_file: str, out_dir: str, log_file=None, seed=1234):
             logger.text(f'Created out_dir: '
                         f'{os.path.abspath(config.out_dir)}')
 
+    # Limit tensorflow to single GPU
+    os.environ['CUDA_VISIBLE_DEVICES'] = '0'
+    os.environ['TF_FORCE_GPU_ALLOW_GROWTH'] = 'True'
+
     # Register all ranking algorithms supported
     register_od_algs()
 

### 2021-09-01 dora_exp_pipeline/pae_outlier_detection.py

@@ -17,7 +17,7 @@ import tensorflow_addons as tfa
 from tensorflow import keras
 from tensorflow.keras import layers, losses
 from tensorflow.keras.models import Model
-from tensorflow.keras.callbacks import EarlyStopping
+from tensorflow.keras.callbacks import EarlyStopping, TensorBoard
 from tensorflow_probability import distributions, bijectors, layers as tfpl
 
 
@@ -27,7 +27,8 @@ class PAEOutlierDetection(OutlierDetection):
 
     def _rank_internal(self, data_to_fit, data_to_score, data_to_score_ids,
                        top_n, seed, latent_dim, max_epochs=500, patience=3,
-                       val_split=0.25, verbose=0, optimizer='adam'):
+                       val_split=0.25, verbose=0, optimizer='adam', 
+                       log_dir=None):
         if data_to_fit is None:
             data_to_fit = deepcopy(data_to_score)
 
@@ -68,39 +69,21 @@ class PAEOutlierDetection(OutlierDetection):
 
 
 def train_and_run_PAE(train, test, latent_dim, num_features, seed, max_epochs,
-                      patience, val_split, verbose, optimizer):
-    # Progress bar formatting
-    autoencoder_lbar = 'Autoencoder training: {percentage:3.0f}%|{bar} '
-    flow_lbar = 'Flow training: {percentage:3.0f}%|{bar} '
-    rbar = '{n_fmt}/{total_fmt} ETA: {remaining}s,  {rate_fmt}{postfix}'
-
-    # Initialize callbacks
-    autoencoder_callbacks = [
-        EarlyStopping(monitor='val_loss', patience=patience),
-        tfa.callbacks.TQDMProgressBar(
-            show_epoch_progress=False,
-            overall_bar_format=autoencoder_lbar + rbar
-        )]
-
-    flow_callbacks = [
-        EarlyStopping(monitor='val_loss', patience=patience),
-        tfa.callbacks.TQDMProgressBar(
-            show_epoch_progress=False,
-            overall_bar_format=flow_lbar + rbar
-        )]
-
+                      patience, val_split, verbose, optimizer, log_dir):
     # Train autoencoder
     autoencoder = Autoencoder(latent_dim, num_features)
     autoencoder.compile(optimizer=optimizer, loss=losses.MeanSquaredError())
     autoencoder.fit(train, train, epochs=max_epochs, verbose=verbose,
[…]

### 2021-09-01 dora_exp_pipeline/conv_pae_outlier_detection.py
Merge conv/non-conv into single file; limit gpu access; make pae-specific progress bars disappear; log training progress for use with Tensorboard

@@ -1,305 +1,0 @@
-# This script implements a convolutional probabilistic autoencoder, drawing on
-# descriptions from the following papers:
-# Böhm, Vanessa, and Uroš Seljak. “Probabilistic Auto-Encoder.”
-# ArXiv:2006.05479 [Cs, Stat], Oct. 2020.
-# Prochaska, J. Xavier, Peter C Cornillon, and David M Reiman. "Deep Learning of
-# Sea Surface Temperature Patterns to Identify Ocean Extremes." Remote Sensing,
-# v. 13,.4 doi: 10.3390/rs13040744
-#
-# See copyright notice at end.
-#
-# Author: Bryce Dubayah
-# Date created: August 16, 2021
-
-from dora_exp_pipeline.outlier_detection import OutlierDetection
-from dora_exp_pipeline.pae_outlier_detection import (
-    NormalizingFlow,
-    make_tf_callbacks
-)
-import os
-import math
-import numpy as np
-from PIL import Image
-from itertools import accumulate
-from copy import deepcopy
-import tensorflow as tf
-import tensorflow_addons as tfa
-from tensorflow import keras
-from tensorflow.keras import layers, losses
-from tensorflow.keras.models import Model
-from tensorflow.keras.callbacks import EarlyStopping 
-
-
-class ConvPAEOutlierDetection(OutlierDetection):
-    def __init__(self):
-        super(ConvPAEOutlierDetection, self).__init__('conv_pae')
-
-    def _rank_internal(self, data_to_fit, data_to_score, data_to_score_ids,
-                       top_n, seed, latent_dim, max_epochs=1000, patience=10,
-                       val_split=0.25, verbose=0, optimizer='adam', 
-                       log_dir=None):
-        if data_to_fit is None:
-            data_to_fit = deepcopy(data_to_score)
-
-        if latent_dim < 1:
-            raise RuntimeError('The dimensionality of the latent space must be '
-                               '>= 1')
-
-        # Check if directory of images was passed in
-        if not is_list_of_images(data_to_fit):
-            raise RuntimeError('The convolutional PAE must be used with the'
-                               'Imagedir data loader')
-
-        # Check that the number of hidden layers <= number of features
-        image_shape = get_image_dimensions(data_to_fit)
-        num_features = np.prod(image_shape)
-        if latent_dim > num_features:
-            raise RuntimeError(f'The dimensionality of the latent space'
[…]

### 2021-09-01 dora_exp_pipeline/dora_exp.py

@@ -23,7 +23,6 @@ from dora_exp_pipeline.random_outlier_detection import RandomOutlierDetection
 from dora_exp_pipeline.negative_sampling_outlier_detection import \
     NegativeSamplingOutlierDetection
 from dora_exp_pipeline.pae_outlier_detection import PAEOutlierDetection
-from dora_exp_pipeline.conv_pae_outlier_detection import ConvPAEOutlierDetection
 from dora_exp_pipeline.util import LogUtil
 from dora_exp_pipeline.dora_feature import extract_feature
 from dora_exp_pipeline.dora_feature import z_score_normalize
@@ -63,10 +62,6 @@ def register_od_algs():
     pae_outlier_detection = PAEOutlierDetection()
     register_od_alg(pae_outlier_detection)
 
-    # Register Convolutional PAE outlier detection algorithm in the pool
-    conv_pae_outlier_detection = ConvPAEOutlierDetection()
-    register_od_alg(conv_pae_outlier_detection)
-
 
 def start(config_file: str, out_dir: str, log_file=None, seed=1234):
     if not os.path.exists(config_file):
@@ -93,9 +88,10 @@ def start(config_file: str, out_dir: str, log_file=None, seed=1234):
             logger.text(f'Created out_dir: '
                         f'{os.path.abspath(config.out_dir)}')
 
-    # Limit tensorflow to single GPU
+    # Configure tensorflow
     os.environ['CUDA_VISIBLE_DEVICES'] = '0'
     os.environ['TF_FORCE_GPU_ALLOW_GROWTH'] = 'True'
+    os.environ['TF_CPP_MIN_LOG_LEVEL'] = '3'
 
     # Register all ranking algorithms supported
     register_od_algs()

### 2021-09-01 dora_exp_pipeline/pae_outlier_detection.py

@@ -1,7 +1,10 @@
-# This script implements a probabilistic autoencoder, as described in the
-# following paper:
+# This script implements a convolutional probabilistic autoencoder, drawing on
+# descriptions from the following papers:
 # Böhm, Vanessa, and Uroš Seljak. “Probabilistic Auto-Encoder.”
 # ArXiv:2006.05479 [Cs, Stat], Oct. 2020.
+# Prochaska, J. Xavier, Peter C Cornillon, and David M Reiman. "Deep Learning of
+# Sea Surface Temperature Patterns to Identify Ocean Extremes." Remote Sensing,
+# v. 13,.4 doi: 10.3390/rs13040744
 #
 # See copyright notice at end.
 #
@@ -10,7 +13,10 @@
 
 from dora_exp_pipeline.outlier_detection import OutlierDetection
 import os
+import math
 import numpy as np
+from PIL import Image
+from itertools import accumulate
 from copy import deepcopy
 import tensorflow as tf
 import tensorflow_addons as tfa
@@ -26,9 +32,8 @@ class PAEOutlierDetection(OutlierDetection):
         super(PAEOutlierDetection, self).__init__('pae')
 
     def _rank_internal(self, data_to_fit, data_to_score, data_to_score_ids,
-                       top_n, seed, latent_dim, max_epochs=500, patience=3,
-                       val_split=0.25, verbose=0, optimizer='adam', 
-                       log_dir=None):
+                       top_n, seed, latent_dim, max_epochs=1000, patience=10,
+                       val_split=0.25, optimizer='adam', log_dir=None):
         if data_to_fit is None:
             data_to_fit = deepcopy(data_to_score)
 
@@ -36,24 +41,28 @@ class PAEOutlierDetection(OutlierDetection):
             raise RuntimeError('The dimensionality of the latent space must be '
                                '>= 1')
 
-        # Check that the number of hidden layers <= number of features
-        num_features = data_to_fit.shape[1]
+        # Autoencoder used in convolutional mode if list of images passed in
+        if is_list_of_images(data_to_fit):
+            sample_shape = get_image_dimensions(data_to_fit)
+            train_fn = train_and_run_conv_PAE
+        else:
+            sample_shape = data_to_fit.shape[1]
+            train_fn = train_and_run_PAE
+
+        num_features = np.prod(sample_shape)
         if latent_dim > num_features:
             raise RuntimeError(f'The dimensionality of the latent space'
[…]

### 2021-09-01 dora_exp_pipeline/lrx_outlier_detection.py
Add LRX progress bar. #46

@@ -8,6 +8,7 @@
 # Steven Lu, July 13, 2021, updated to be compatible with DORA pipeline.
 
 import numpy as np
+from tqdm import tqdm
 from dora_exp_pipeline.outlier_detection import OutlierDetection
 
 
@@ -59,7 +60,8 @@ def get_LRX_scores(images, w_inner, w_outer, bands):
     scores = np.zeros([rows, im_height, im_width])
     s = int(w_outer/2)
     # for each image, compute the LRX score in each pixel
-    for idx in range(images.shape[0]):
+    for idx in tqdm(range(images.shape[0]),
+                    desc='LRX'):
         im = images[idx]
         for i in range(s, im.shape[0]-s):
             for j in range(s, im.shape[1]-s):

### 2021-09-01 dora_exp_pipeline/dora_exp.py
Update tensorboard config; re-run fmnist experiment

@@ -10,6 +10,7 @@
 
 import os
 import sys
+import logging
 from tqdm import tqdm
 from dora_exp_pipeline.dora_config import DoraConfig
 from dora_exp_pipeline.dora_data_loader import get_data_loader_by_name
@@ -92,6 +93,7 @@ def start(config_file: str, out_dir: str, log_file=None, seed=1234):
     os.environ['CUDA_VISIBLE_DEVICES'] = '0'
     os.environ['TF_FORCE_GPU_ALLOW_GROWTH'] = 'True'
     os.environ['TF_CPP_MIN_LOG_LEVEL'] = '3'
+    logging.getLogger("tensorflow").setLevel(logging.ERROR)
 
     # Register all ranking algorithms supported
     register_od_algs()

### 2021-09-01 dora_exp_pipeline/outlier_detection.py

@@ -93,7 +93,7 @@ class OutlierDetection(object):
         """
         ret_string = ''
         for k, v in params_dict.items():
-            ret_string += '-%s=%s' % (k, v)
+            ret_string += '-%s=%s' % (k, str(v).replace('/', ''))
 
         return ret_string
 

### 2021-09-01 dora_exp_pipeline/pae_outlier_detection.py

@@ -62,7 +62,7 @@ class PAEOutlierDetection(OutlierDetection):
         # Rank targets
         scores = train_fn(data_to_fit, data_to_score, latent_dim,
                           sample_shape, seed, max_epochs, patience, val_split,
-                          optimizer)
+                          optimizer, log_dir)
         selection_indices = np.argsort(scores)[::-1]
 
         results = dict()
@@ -152,7 +152,8 @@ def make_tf_callbacks(name, patience, log_dir):
             overall_bar_format=name + lbar + rbar
         )]
     if log_dir:
-        callbacks.append(TensorBoard(log_dir=log_dir, histogram_freq=1))
+        logs_path = os.path.join(log_dir, name.replace(' ', '_'))
+        callbacks.append(TensorBoard(log_dir=logs_path, histogram_freq=5))
     return callbacks
 
 

### 2021-09-01 publications/dora_outlier_detection_iaai.tex
Add draft Intro text, mostly from proposal. #18

@@
    cross-domain applicability of DORA, we conducted outlier detection
-    experiments for four use cases: 1) cleaning noisy ground-truth datasets
-     (Earth Science), 2) detecting volcanic thermal anomalies (Earth Science), 
-     3) flagging non-astrophysical objects in the Dark Energy Survey 
-     (astrophysics), and 4) novelty-guided onboard targeting of instruments
+    experiments for three real-world use cases: 1) cleaning noisy ground-truth 
+    datasets (Earth Science), 
+     2) flagging non-astrophysical objects in the Dark Energy Survey 
+     (astrophysics), and 3) novelty-guided onboard targeting of instruments
       for Mars rovers (planetary science). [We showed that ...]
@@
 \section{Introduction}
-\begin{itemize}
-\item Framing of outlier detection as a domain-agnostic problem that is well-
-suited to cross-disciplinary applications
-\item Example applications of outlier detection across science application 
-domains
-\item Challenges/barriers to entry for domain scientists who could benefit from
- outlier detection 
-\item DORA as a solution to lower these barriers to entry and help integrate 
-outlier detection methods into existing scientific workflows across variety of 
-domains
-\end{itemize}
+The ability to automatically detect out-of-distribution samples in large data 
+sets is of interest for a wide variety of scientific domains. Depending on the
+ application setting, this capability is also commonly referred to as anomaly
+  detection, outlier detection, or novelty detection. More broadly, this is 
+  referred to as out-of-distribution (OOD) detection. In general, the goal of 
+  OOD detection systems is to identify samples that deviate from the majority
+   of samples in a dataset in an unsupervised manner (Pimentel et al., 2014). 
+   In machine learning, these methods are commonly used for identifying 
+   mislabeled or otherwise invalid samples in a dataset 
+   (e.g., Liang et al., 2018; Böhm et al., 2020).
+
+   When working with science datasets, OOD detection can be used for 
+   cleaning datasets, e.g., flagging ground-truth labels with GPS or human
+    entry error or identifying wrongly categorized objects in a catalog. 
+    It could also be used for discovery, e.g., to flag novel samples in order 
+    to guide instrument acquisition or scientific analysis. Another application
+   is the detection of rare objects that are known to exist but the known
+   examples are too few to create a large enough labeled dataset for 
+   supervised classification algorithms. In astrophysics, 
+    outlier detection methods have been employed for identifying astrophysical
+     objects with unique characteristics (Xiong et al., 2010; 
+     Storey-Fisher et al., 2020; Hayat et al., 2020) as well as 
+     data/modeling artifacts (Wagstaff et al., 2020; Lochner et al., 2020) in 
+     astronomical surveys. Examples of outlier detection applications in 
+     Earth science include detecting anomalous objects or materials 
+     (Chang and Chiang, 2002; Zhou et al., 2016), data artifacts/noise 
+     (Liu et al., 2017; Alvera-Azcárate et al., 2012), change (Touati et al.,
+      2020; Zhou et al., 2016),  and ocean extremes (Prochaska et al.). 
+      Planetary science applications of outlier detection have mostly focused 
+      on prioritizing samples with novel geologic or geochemical features for
+       follow-up targeting or analysis (Kerner et al., 2020a; Kerner et al., 
+       2020b; Wagstaff et al., 2013).
+
+Despite wide differences in applications, data types, and dimensionality,
+ the same underlying machine learning algorithms can be employed across 
+ all of these domains. A challenge for applying them however is that domain
+ scientists do not always have the programming or machine learning background
+ to apply the algorithms themselves using existing tools. Given the widespread 
+ applicability and transferability of OOD methods, the scientific community 
+ would benefit from a tool that made it easy for them to apply popular outlier
+ detection algorithms to their science datasets. We created DORA (Domain-
+ agnostic Outlier Ranking Algorithms) to provide a tool for applying outlier 
+ detection algorithms to a variety of scientific data sets with minimal coding
+ required. Users need only to specify details for their data/application 
+ including the data type, location, and algorithms to run in an experiment
+ configuration file. DORA supports image, raster, time series, 
+or feature vector input data types and outlier detection methods that include
+ Isolation Forest, Discovery via Eigenbasis Modeling of Uninteresting Data 
+ (DEMUD), principal component analysis (PCA), Reed-Xiaoli (RX) detector, 
+ Local RX, negative sampling, and probabilistic autoencoder. Given a set of
+  data samples, each algorithm assigns an outlier score to each sample. DORA
+   provides results organization and visualization modules to help users 
+   process the results, including sorting samples by outlier score, evaluating 
+   outlier recall for a set of known/labeled outliers, clustering groups of 
+   similar outliers together, and causal graphs. To demonstrate the 
+   cross-domain applicability of DORA, we conducted outlier detection
+    experiments for three real-world use cases: 1) cleaning noisy ground-truth 
+    datasets (Earth Science), 
+     2) flagging non-astrophysical objects in the Dark Energy Survey 
+     (astrophysics), and 3) novelty-guided onboard targeting of instruments
+      for Mars rovers (planetary science). [We showed that ...]
 
@@
 \item Tools for applying outlier detection or other methods across domains?
-
 \end{itemize}
@@
 
-\subsubsection{Volcanic thermal anomalies}
+% Decided to cut this use case to make the presentation simpler (only one
+% Earth science use case, and we already have a 1-band image dataset
+% (Navcam). 
+%\subsubsection{Volcanic thermal anomalies}
 
@@
 
-\subsubsection{Volcanic thermal anomalies}
+%\subsubsection{Volcanic thermal anomalies}
 

### 2021-09-01 publications/dora_outlier_detection_iaai.tex
Move text about specific studies to Related Work. #42

@@
    examples are too few to create a large enough labeled dataset for 
-   supervised classification algorithms. In astrophysics, 
-    outlier detection methods have been employed for identifying astrophysical
-     objects with unique characteristics (Xiong et al., 2010; 
-     Storey-Fisher et al., 2020; Hayat et al., 2020) as well as 
-     data/modeling artifacts (Wagstaff et al., 2020; Lochner et al., 2020) in 
-     astronomical surveys. Examples of outlier detection applications in 
-     Earth science include detecting anomalous objects or materials 
-     (Chang and Chiang, 2002; Zhou et al., 2016), data artifacts/noise 
-     (Liu et al., 2017; Alvera-Azcárate et al., 2012), change (Touati et al.,
-      2020; Zhou et al., 2016),  and ocean extremes (Prochaska et al.). 
-      Planetary science applications of outlier detection have mostly focused 
-      on prioritizing samples with novel geologic or geochemical features for
-       follow-up targeting or analysis (Kerner et al., 2020a; Kerner et al., 
-       2020b; Wagstaff et al., 2013).
-
+   supervised classification algorithms. 
   Despite wide differences in applications, data types, and dimensionality,
@@
 \section{Related Work}
+In astrophysics, 
+    outlier detection methods have been employed for identifying astrophysical
+     objects with unique characteristics (Xiong et al., 2010; 
+     Storey-Fisher et al., 2020; Hayat et al., 2020) as well as 
+     data/modeling artifacts (Wagstaff et al., 2020; Lochner et al., 2020) in 
+     astronomical surveys. Examples of outlier detection applications in 
+     Earth science include detecting anomalous objects or materials 
+     (Chang and Chiang, 2002; Zhou et al., 2016), data artifacts/noise 
+     (Liu et al., 2017; Alvera-Azcárate et al., 2012), change (Touati et al.,
+      2020; Zhou et al., 2016),  and ocean extremes (Prochaska et al.). 
+      Planetary science applications of outlier detection have mostly focused 
+      on prioritizing samples with novel geologic or geochemical features for
+       follow-up targeting or analysis (Kerner et al., 2020a; Kerner et al., 
+       2020b; Wagstaff et al., 2013).
+
 \begin{itemize}

### 2021-09-02 publications/dora_outlier_detection_iaai.tex
Add references for DES anomaly and DEMUD papers. (#42)

--- 2021-09-01 publications/dora_outlier_detection_iaai.tex
@@
 \section{Related Work}
+In astrophysics, 
+    outlier detection methods have been employed for identifying astrophysical
+     objects with unique characteristics (Xiong et al., 2010; 
+     Storey-Fisher et al., 2020; Hayat et al., 2020) as well as 
+     data/modeling artifacts~\citep{wagstaff:des-anom20} (Lochner et al., 2020) in 
+     astronomical surveys. Examples of outlier detection applications in 
+     Earth science include detecting anomalous objects or materials 
+     (Chang and Chiang, 2002; Zhou et al., 2016), data artifacts/noise 
+     (Liu et al., 2017; Alvera-Azcárate et al., 2012), change (Touati et al.,
+      2020; Zhou et al., 2016),  and ocean extremes (Prochaska et al.). 
+      Planetary science applications of outlier detection have mostly focused 
+      on prioritizing samples with novel geologic or geochemical features for
+       follow-up targeting or analysis (Kerner et al., 2020a; Kerner et al., 
+       2020b) \citep{wagstaff:demud13}.
+
 \begin{itemize}

### 2021-09-02 publications/dora_outlier_detection_iaai.tex
Add reference to PyOD. (#42)
Add description of DEMUD. (#42)

@@
  Isolation Forest, Discovery via Eigenbasis Modeling of Uninteresting Data 
- (DEMUD), principal component analysis (PCA), Reed-Xiaoli (RX) detector, 
+ (DEMUD)~\citep{wagstaff:demud13}, principal component analysis (PCA),
+   Reed-Xiaoli (RX) detector,  
  Local RX, negative sampling, and probabilistic autoencoder. Given a set of
@@
 (e.g., \citet{kerner2020comparison})
-\item Tools for applying outlier detection or other methods across domains?
+\item Outlier detection toolkits: PyOD Python
+package~\citep{zhao:pyod19}, \url{https://github.com/yzhao062/pyod} 
+\item Tools for applying outlier detection or other methods across
+domains?
 \end{itemize}
@@
 complement each other.]
+
+The DEMUD algorithm~\citep{wagstaff:demud13} differs from other
+outlier ranking methods: instead of independently scoring all
+observations, DEMUD incrementally identifies the most unusual
+remaining item, then incorporates it into the model of ``known''
+(non-outlier) observations before selecting the next most unusual
+item.  DEMUD's goal is to identify diverse outliers and avoid
+redundant selections.  Once an outlier is encountered, repeated
+occurrences of that outlier are deprioritized.  Methods that score
+outliers independently maximize coverage of outliers, while DEMUD
+maximizes the fast discovery of distinct outlier types.
+
 \paragraph{Results organization} Eric, Umaa, Steven, Vinay
@@
 \end{document}
+
+% LocalWords:  tex OOD Eigenbasis DEMUD PCA Xiaoli RX autoencoder al
+% LocalWords:  Pimentel mislabeled Liang Böhm wagstaff demud Xiong
+% LocalWords:  Storey Hayat des anom Lochner Chiang Zhou Liu Alvera
+% LocalWords:  Azcárate Touati Prochaska Kerner kerner PyOD zhao pyod
+% LocalWords:  API Umaa Vinay Navcam MNIST FashionMNIST installable
+% LocalWords:  dora CIF SMD FAO Pritchard

### 2021-09-02 publications/dora_outlier_detection_iaai.tex
Add description of rover target data set and relevant updates to acknowledgements (#42)

@@
 \setcounter{secnumdepth}{0} %May be changed to 1 or 2 if section numbers are desired.
+
+\usepackage{siunitx}
+\usepackage[dvipsnames]{xcolor}
+\newcommand{\todo}[1]{\textcolor{blue}{#1}}
 
@@
 \paragraph{Feature extraction} Hannah 
-\paragraph{Ranking algorithms} We implemented 7 algorithms for scoring and 
+\paragraph{Ranking algorithms.} We implemented 7 algorithms for scoring and 
 ranking samples by outlierness. [Describe each algorithm briefly and how they 
@@
 
-\subsubsection{Targets in Mars rover images}
-Kiri
+\subsubsection{Targets in Mars rover images.}
+Mars exploration is fundamentally an exercise in discovery with the
+goal of increasing our understanding of the history, evolution,
+composition, and currently active processes on Mars.  Outliers
+identified in Mars observations can inspire new discoveries and inform
+the choice of which areas merit follow-up, deeper
+investigation~\cite{kerner2020comparison,wagstaff:rover-novelty20}.
+We collected \todo{XXX} images from the Mars Science Laboratory's
+Navigation camera (Navcam) and employed the Rockster~\cite{burl:rockster16}
+method used by onboard rover software to identify candidate rock
+targets with an area of at least 100 pixels, yielding \num{6005}
+targets.  We cropped out a \num{64}$\times$\num{64} pixel image
+centered on each target.
+%
+We collaborated with an MSL science planning team member to
+independently review the targets and identify those that were
+considered novel by the mission ($n_{novel} = 108$).
 
@@
 
-\subsubsection{Targets in Mars rover images}
+\subsubsection{Targets in Mars rover images.}
 Kiri
-
+% and organized them chronologically.
 \subsubsection{Satellite time series for ground-truth observations}
@@
 
-\section{ Acknowledgments}
-Funding/support: JPL CIF, SMD etc. Acknowledge people who contributed to 
-datasets e.g. FAO, Matt Pritchard, DES, other CIF team members?
+\section{Acknowledgments}
+\todo{Funding/support: JPL CIF, SMD etc.}
+We thank the NASA Center Innovation Fund for supporting this work.
+
+\todo{Acknowledge people who contributed to 
+datasets e.g. FAO, Matt Pritchard, DES, other CIF team members?}
+
+We thank Raymond Francis for analyzing the Navcam rover images to
+identify targets considered to be ``novel'' for evaluating outlier
+detection in the planetary rover scenario.
+
+Part of this research was carried out at the Jet Propulsion
+Laboratory, California Institute of Technology, under a contract with
+the National Aeronautics and Space Administration.
 

### 2021-09-02 dora_exp_pipeline/combined_plot_script.py
Added script to create combined plots

@@ -1,0 +1,94 @@
+import os
+import sys
+#sys.path.append('.')
+import numpy as np
+import matplotlib.pyplot as plt
+import argparse
+
+def alg_indexes(filename):
+    with open(filename, 'r') as f:
+        text = f.read().split("\n")[:-1]
+        scores = sorted(text, key=lambda x:float(x.split(", ")[-1]))
+        scores = [int(i.split(", ")[1]) for i in scores]
+
+    return scores
+
+def filenames_and_labels(root):
+    filesAndFolders = os.listdir(root)
+    labels = []
+    filenames = []
+    for i in filesAndFolders:
+        if('.' in i):
+            continue
+        name = i
+        if('-' in name):
+            name = name[:name.index('-')]
+        labels.append(name.replace('_', ' '))
+        path = os.path.join(root, i, 'selections-{}.csv'.format(name))
+        filenames.append(path)
+
+    return filenames, labels
+
+def validation_labels(filename):
+    with open(filename, 'r') as f:
+        text = f.read().split("\n")[:-1]
+
+    labels = {}
+    for i in text:
+        line = i.split(",")
+        labels[int(line[0])] = int(line[1])
+
+    return labels
+
+def get_curve(scores, validationLabels):
+    x = list(range(1, len(scores)+1))
+    y = []
+    numOutliers = 0
+
+    for i in range(len(scores)):
+        if(validationLabels[scores[i]] == 1):
+            numOutliers += 1
+        y.append(numOutliers)
+
+    return x, y
+
+def combine_plots(root, validationDir):
+    filenames, labels = filenames_and_labels(root)
+
+    maxX, maxY = -1, -1
+    fig, axes = plt.subplots()
+
+    validationLabels = validation_labels(validationDir)
+    for i in range(len(filenames)):
+        scores = alg_indexes(filenames[i])
+
+        x,y = get_curve(scores, validationLabels)
+        maxX = max(maxX, x[-1])
+        maxY = max(maxY, y[-1])
+
+        index = x.index(y[-1])
+        #area = np.trapz(y, x)
+        area = np.trapz(y[:index+1], x[:index+1])
+
+        plt.plot(x, y, label="{} Area: {}".format(labels[i], area))
+
+    bestLine = list(range(1, maxX+1))
+    plt.plot(bestLine, bestLine, label='Best Line', color='darkgreen')
+    plt.title('Correct Outliers vs Selected Outliers')
[…]

### 2021-09-02 dora_exp_pipeline/dora_results_organization.py

@@ -99,14 +99,13 @@ class SaveComparisonPlot(ResultsOrganization):
 
         # Outliers will be 1s and inliers will be 0s.
         labels = self._get_validation_labels(validation_dir)
-        scores = np.argsort(dts_scores)[::-1]
 
-        x = list(range(1, len(scores)+1))
+        x = list(range(1, len(dts_sels)+1))
         y = []
         numOutliers = 0
 
-        for i in range(len(scores)):
-            if(labels[scores[i]] == 1):
+        for i in range(len(dts_sels)):
+            if(labels[dts_sels[i]] == 1):
                 numOutliers += 1
             y.append(numOutliers)
 
@@ -114,7 +113,8 @@ class SaveComparisonPlot(ResultsOrganization):
         index = x.index(y[-1])
         area = np.trapz(y[:index+1], x[:index+1])
 
-        plt.plot(x, y, label=outlier_alg_name)
+        plt.plot(x, y, label=alg_name)
+        plt.plot(x, x, label='Best Line')
         plt.plot([], [], ' ', label=f'Area: {area}')
         plt.title('Correct Outliers vs Selected Outliers')
         plt.xlabel('Number of Outliers Selected')
@@ -122,7 +122,7 @@ class SaveComparisonPlot(ResultsOrganization):
         plt.legend()
         axes.set_xlim(1, x[-1])
         axes.set_ylim(1, y[-1])
-        plt.savefig(f'{out_dir}/comparison_plot_{outlier_alg_name}.png')
+        plt.savefig(f'{out_dir}/comparison_plot_{alg_name}.png')
 
     def _get_validation_labels(self, validation_dir):
         with open(validation_dir, 'r') as f:
@@ -270,6 +270,30 @@ class ReshapeRaster(ResultsOrganization):
 reshape_raster = ReshapeRaster()
 register_org_method(reshape_raster)
 
+class SaveHistogram(ResultsOrganization):
+    def __init__(self):
+        super(SaveHistogram, self).__init__('histogram')
+
+    def _run(self, data_ids, dts_scores, dts_sels, data_to_score, alg_name,
+             out_dir, logger, seed, bins):
+        if(not(os.path.exists(out_dir))):
+            os.makedirs(out_dir)
+
+        scores = sorted(dts_scores)
+        fig, axs = plt.subplots()
+        #numBins = int((scores[-1]-scores[0])/increment)+1
+        #print(scores[0], scores[-1], increment, numBins)
+
[…]

### 2021-09-02 publications/dora_outlier_detection_iaai.tex

@@
 independently review the targets and identify those that were
-considered novel by the mission ($n_{novel} = 108$).
+considered novel by the mission ($n_{novel} = 108$).  Our goal for
+this application is to assess how well the selections made by each
+algorithm correlate with human novelty judgments to determine which
+methods would be most suitable for informing onboard decisions about
+follow-up observations.
 
@@
 \subsubsection{Targets in Mars rover images.}
-Kiri
-% and organized them chronologically.
+We simulated the operational setting in which the rover has observed
+targets up through mission day (sol) $s$ and the goal is to rank all
+subsequent targets (after sol $s$) to inform which recent targets
+merit further study.  We organized the Mars target images
+chronologically and partitioned them into the corresponding ``prior''
+set $D_s$ and ``assessment'' set $D_{s+}$. We initialized each
+algorithm with $D_s$ and requested a ranking of targets in $D_{s+}$.
+\todo{Add table showing number (or percentage?) of novel targets in
+the top $N$ selections, where $N$ = known number of novel targets in
+$D_{s+}$.  And/or, report average rank of novel targets if alg ranks all
+$|D_{s+}|$ targets. If time/room permits, do this for different values
+of $s$.}
+
 \subsubsection{Satellite time series for ground-truth observations}

### 2021-09-03 dora_exp_pipeline/combined_plot_script.py
Add pytests to github actions; cache dependencies for actions; fix lint errors

@@ -1,17 +1,19 @@
 import os
-import sys
-#sys.path.append('.')
+# import sys
+# sys.path.append('.')
 import numpy as np
 import matplotlib.pyplot as plt
 import argparse
 
+
 def alg_indexes(filename):
     with open(filename, 'r') as f:
         text = f.read().split("\n")[:-1]
-        scores = sorted(text, key=lambda x:float(x.split(", ")[-1]))
+        scores = sorted(text, key=lambda x: float(x.split(", ")[-1]))
         scores = [int(i.split(", ")[1]) for i in scores]
 
     return scores
+
 
 def filenames_and_labels(root):
     filesAndFolders = os.listdir(root)
@@ -29,6 +31,7 @@ def filenames_and_labels(root):
 
     return filenames, labels
 
+
 def validation_labels(filename):
     with open(filename, 'r') as f:
         text = f.read().split("\n")[:-1]
@@ -39,6 +42,7 @@ def validation_labels(filename):
         labels[int(line[0])] = int(line[1])
 
     return labels
+
 
 def get_curve(scores, validationLabels):
     x = list(range(1, len(scores)+1))
@@ -52,6 +56,7 @@ def get_curve(scores, validationLabels):
 
     return x, y
 
+
 def combine_plots(root, validationDir):
     filenames, labels = filenames_and_labels(root)
 
@@ -62,12 +67,12 @@ def combine_plots(root, validationDir):
     for i in range(len(filenames)):
         scores = alg_indexes(filenames[i])
 
-        x,y = get_curve(scores, validationLabels)
+        x, y = get_curve(scores, validationLabels)
         maxX = max(maxX, x[-1])
         maxY = max(maxY, y[-1])
 
         index = x.index(y[-1])
-        #area = np.trapz(y, x)
+        # area = np.trapz(y, x)
         area = np.trapz(y[:index+1], x[:index+1])
 
         plt.plot(x, y, label="{} Area: {}".format(labels[i], area))
@@ -82,6 +87,7 @@ def combine_plots(root, validationDir):
     axes.set_ylim(1, maxY)
     plt.savefig(f'{root}/comparison_plot_combined.png')
 
+
 def main():
     parser = argparse.ArgumentParser()
     parser.add_argument('root', type=str, help='Path to the results directory')
@@ -90,5 +96,6 @@ def main():
 
     args = parser.parse_args()
     combine_plots(**vars(args))
+
 
 main()

### 2021-09-03 dora_exp_pipeline/dora_results_organization.py

@@ -113,7 +113,7 @@ class SaveComparisonPlot(ResultsOrganization):
         index = x.index(y[-1])
         area = np.trapz(y[:index+1], x[:index+1])
 
-        plt.plot(x, y, label=alg_name)
+        plt.plot(x, y, label=outlier_alg_name)
         plt.plot(x, x, label='Best Line')
         plt.plot([], [], ' ', label=f'Area: {area}')
         plt.title('Correct Outliers vs Selected Outliers')
@@ -122,7 +122,7 @@ class SaveComparisonPlot(ResultsOrganization):
         plt.legend()
         axes.set_xlim(1, x[-1])
         axes.set_ylim(1, y[-1])
-        plt.savefig(f'{out_dir}/comparison_plot_{alg_name}.png')
+        plt.savefig(f'{out_dir}/comparison_plot_{outlier_alg_name}.png')
 
     def _get_validation_labels(self, validation_dir):
         with open(validation_dir, 'r') as f:
@@ -270,6 +270,7 @@ class ReshapeRaster(ResultsOrganization):
 reshape_raster = ReshapeRaster()
 register_org_method(reshape_raster)
 
+
 class SaveHistogram(ResultsOrganization):
     def __init__(self):
         super(SaveHistogram, self).__init__('histogram')
@@ -281,16 +282,16 @@ class SaveHistogram(ResultsOrganization):
 
         scores = sorted(dts_scores)
         fig, axs = plt.subplots()
-        #numBins = int((scores[-1]-scores[0])/increment)+1
-        #print(scores[0], scores[-1], increment, numBins)
+        # numBins = int((scores[-1]-scores[0])/increment)+1
+        # print(scores[0], scores[-1], increment, numBins)
 
-        yVals, bins, patches = axs.hist(scores, bins, density=True,
-                                            alpha=0.5)
+        yVals, bins, patches = axs.hist(scores, bins, density=True, alpha=0.5)
 
         plt.title('Histogram of Anomaly Scores')
         plt.xlabel('Score')
         plt.ylabel('Frequency')
         plt.savefig(f'{out_dir}/histogram_bar_graph-{alg_name}.png')
+
 
 save_histogram = SaveHistogram()
 register_org_method(save_histogram)

### 2021-09-05 publications/dora_outlier_detection_iaai.tex
Add description of Earth science dataset. #42

@@
 \subsection{Datasets}
-Describe the datasets overall and our motivation for choosing these. Maybe add
+We constructed three datasets to evaluate the utility of DORA and the
+performance of each ranking algorithm for a variety of scientific domains
+ (astrophysics, planetary science, and Earth science). We also included a 
+ benchmark dataset that uses MNIST and FashionMNIST. Here we describe
+ the details of each dataset.
+\todo{Maybe add
  a table with number of samples, number of reference known outliers (if 
- available), etc, if room?
+ available), etc, if room? }
+ \todo{Can we make any of these datasets publicly available with the paper?}
 
@@
 identified in Mars observations can inspire new discoveries and inform
-the choice of which areas merit follow-up, deeper
+the choice of which areas merit follow-up or deeper
 investigation~\cite{kerner2020comparison,wagstaff:rover-novelty20}.
-We collected \todo{XXX} images from the Mars Science Laboratory's
-Navigation camera (Navcam) and employed the Rockster~\cite{burl:rockster16}
-method used by onboard rover software to identify candidate rock
+We collected \todo{XXX} images from the Navigation camera (Navcam) 
+onboard the Mars Science Laboratory (MSL) rover
+and employed the Rockster~\cite{burl:rockster16}
+method currently used by onboard rover software to identify candidate rock
 targets with an area of at least 100 pixels, yielding \num{6005}
@@
 \subsubsection{Satellite time series for ground-truth observations}
-Hannah: Describe the Earth crop field data set and applications
+Many Earth science applications using satellite Earth observation (EO) data
+ require ground-truth observations for understanding and modeling ground-
+ identified objects in the satellite observations. These ground observations
+ also serve as labels that are paired with satellite data inputs for machine 
+ learning models. For example, a model trained to classify crop types in
+ satellite observations requires geo-referenced labels of crop type annotated
+ in the field~\cite{tseng2021learning}. A widespread challenge of using ground-
+ annotated datasets is that there are often points with erroneous location or 
+ label information (e.g., due to GPS location error or human entry error) that
+ need to be cleaned before the dataset can be used for machine learning or
+ other downstream use cases. Automatically detecting these outliers could save 
+ substantial time required for cleaning datasets and improve the performance of
+ downstream analyses that rely on high-quality datasets. 
+ %
+ We used a dataset of ground
+ observations of fall army worm (FAW) in maize crops collected by the UN Food
+ and Agriculture Organization (FAO). This dataset includes \todo{7XXX} samples
+ with location (latitude/longitude) metadata covering \todo{a few countries}. 
+ Most locations coincide with crop
+ fields but there are many outliers that coincide with other land cover types 
+ such as water, buildings, or forests. We constructed an evaluation set 
+ consisting of all samples in Kenya ($n=XX$) and manually annotated whether 
+ each sample was considered an inlier (i.e., located in a crop field) or not 
+ using high-resolution Planet satellite images in Collect Earth 
+ Online~\cite{planetdata} ($n_{inlier}=XX$, $n_{outlier}=XX$). We used Google
+ Earth Engine to export the Sentinel-1 synthetic aperture radar (SAR) monthly
+ time series for each sample location, using data from year each sample was 
+ collected. We used SAR data because it is sensitive to ground texture and
+  penetrates clouds, which is important for the often-cloudy region covered
+  by the dataset. Our goal for this application is to assess how well the 
+  selections made by each algorithm correlate with outliers determined by
+  visual inspection of the satellite images.
+ \todo{If we have time, I want to replace this with a different dataset and 
+ experiment that is in progress that evaluates the downstream classifier
+ performance before and after applying DORA.}
 
@@
 \section{Acknowledgments}
-\todo{Funding/support: JPL CIF, SMD etc.}
-We thank the NASA Center Innovation Fund for supporting this work.
+We thank the NASA Center Innovation Fund and the NASA Science Mission 
+Directorate (under the NASA Cross-Divisional AI Use Case Demonstration grants)
+ for supporting this work.
 
-\todo{Acknowledge people who contributed to 
-datasets e.g. FAO, Matt Pritchard, DES, other CIF team members?}
+\todo{Acknowledge DES?}
+We thank the UN Food and Agriculture Organization (FAO) and Dr. Ritvik Sahajpal
+(University of Maryland) for providing the fall army worm ground observation
+dataset.
 

### 2021-09-05 publications/dora_outlier_detection_iaai.tex
Update missing information about Earth dataset. #42

@@
  observations of fall army worm (FAW) in maize crops collected by the UN Food
- and Agriculture Organization (FAO). This dataset includes \todo{7XXX} samples
- with location (latitude/longitude) metadata covering \todo{a few countries}. 
- Most locations coincide with crop
+ and Agriculture Organization (FAO). This dataset includes 6,757 samples
+ with location (latitude/longitude) metadata primarily in Africa and Southeast
+ Asia. Most locations coincide with crop
  fields but there are many outliers that coincide with other land cover types 
  such as water, buildings, or forests. We constructed an evaluation set 
- consisting of all samples in Kenya ($n=XX$) and manually annotated whether 
+ consisting of all samples in Kenya ($n=113$) and manually annotated whether 
  each sample was considered an inlier (i.e., located in a crop field) or not 
  using high-resolution Planet satellite images in Collect Earth 
- Online~\cite{planetdata} ($n_{inlier}=XX$, $n_{outlier}=XX$). We used Google
+ Online~\cite{planetdata} ($n_{inlier}=76$, $n_{outlier}=37$). We used Google
  Earth Engine to export the Sentinel-1 synthetic aperture radar (SAR) monthly
@@
   visual inspection of the satellite images.
- \todo{If we have time, I want to replace this with a different dataset and 
- experiment that is in progress that evaluates the downstream classifier
- performance before and after applying DORA.}
+ \todo{If we have time, I want to also evaluate the downstream classifier
+ performance before and after removing top N outliers with DORA.}
 

### 2021-09-05 dora_exp_pipeline/combined_plot_script.py
Minor (mostly formatting) changes to combined plot script. #53

@@ -15,8 +15,8 @@ def alg_indexes(filename):
     return scores
 
 
-def filenames_and_labels(root):
-    filesAndFolders = os.listdir(root)
+def filenames_and_labels(resultsdir):
+    filesAndFolders = os.listdir(resultsdir)
     labels = []
     filenames = []
     for i in filesAndFolders:
@@ -26,7 +26,7 @@ def filenames_and_labels(root):
         if('-' in name):
             name = name[:name.index('-')]
         labels.append(name.replace('_', ' '))
-        path = os.path.join(root, i, 'selections-{}.csv'.format(name))
+        path = os.path.join(resultsdir, i, 'selections-{}.csv'.format(name))
         filenames.append(path)
 
     return filenames, labels
@@ -57,13 +57,13 @@ def get_curve(scores, validationLabels):
     return x, y
 
 
-def combine_plots(root, validationDir):
-    filenames, labels = filenames_and_labels(root)
+def combine_plots(resultsdir, label_path):
+    filenames, labels = filenames_and_labels(resultsdir)
 
     maxX, maxY = -1, -1
     fig, axes = plt.subplots()
 
-    validationLabels = validation_labels(validationDir)
+    validationLabels = validation_labels(label_path)
     for i in range(len(filenames)):
         scores = alg_indexes(filenames[i])
 
@@ -75,24 +75,26 @@ def combine_plots(root, validationDir):
         # area = np.trapz(y, x)
         area = np.trapz(y[:index+1], x[:index+1])
 
-        plt.plot(x, y, label="{} Area: {}".format(labels[i], area))
+        plt.plot(x, y, label="{} (AUC: {})".format(labels[i], area))
 
     bestLine = list(range(1, maxX+1))
-    plt.plot(bestLine, bestLine, label='Best Line', color='darkgreen')
-    plt.title('Correct Outliers vs Selected Outliers')
-    plt.xlabel('Number of Outliers Selected')
-    plt.ylabel('Number of True Outliers')
+    plt.plot(bestLine, bestLine, label='Best Line', linestyle='--', color='k')
+    plt.title('True Outliers vs. Algorithm Selections')
+    plt.xlabel('Number of selections')
+    plt.ylabel('Number of true outliers')
     plt.legend(loc='upper left')
     axes.set_xlim(1, maxX)
     axes.set_ylim(1, maxY)
-    plt.savefig(f'{root}/comparison_plot_combined.png')
+    plt.savefig(f'{resultsdir}/comparison_plot_combined.png')
[…]

### 2021-09-05 dora_exp_pipeline/combined_plot_script.py
Add option to plot in terms of precision at N. #53

@@ -44,7 +44,7 @@ def validation_labels(filename):
     return labels
 
 
-def get_curve(scores, validationLabels):
+def get_selections_curve(scores, validationLabels):
     x = list(range(1, len(scores)+1))
     y = []
     numOutliers = 0
@@ -57,36 +57,61 @@ def get_curve(scores, validationLabels):
     return x, y
 
 
-def combine_plots(resultsdir, label_path):
+def get_precision_curve(scores, validationLabels):
+    x = list(range(1, len(scores)+1))
+    y = []
+    numOutliers = 0
+    for i in range(len(scores)):
+        if(validationLabels[scores[i]] == 1):
+            numOutliers += 1
+        y.append(float(numOutliers)/(i+1))
+
+    return x, y
+
+
+def combine_plots(resultsdir, label_path, precision_at_n):
     filenames, labels = filenames_and_labels(resultsdir)
 
-    maxX, maxY = -1, -1
     fig, axes = plt.subplots()
 
     validationLabels = validation_labels(label_path)
+    num_outliers = sum([x for x in validationLabels.values()])
+
+    if precision_at_n:
+        plt.axhline(y=1, label='Best line', linestyle='--', color='k')
+        plt.title('Precision at N')
+        plt.xlabel('Number of selections (N)')
+        plt.ylabel('Precision at N')
+    else:
+        bestLine = list(range(1, num_outliers+1))
+        plt.plot(bestLine, bestLine, label='Best line', linestyle='--',
+                 color='k')
+        plt.title('True Outliers vs. Algorithm Selections')
+        plt.xlabel('Number of selections')
+        plt.ylabel('Number of true outliers')
+
     for i in range(len(filenames)):
         scores = alg_indexes(filenames[i])
 
-        x, y = get_curve(scores, validationLabels)
-        maxX = max(maxX, x[-1])
-        maxY = max(maxY, y[-1])
+        if precision_at_n:
+            x, y = get_precision_curve(scores, validationLabels)
+            plt.plot(x, y, label="{}".format(labels[i]))
+        else:
+            x, y = get_selections_curve(scores, validationLabels)
+            index = x.index(y[-1])
+            area = np.trapz(y[:index+1], x[:index+1])
+            plt.plot(x, y, label="{} (AUC: {})".format(labels[i], area))
 
-        index = x.index(y[-1])
-        # area = np.trapz(y, x)
[…]

### 2021-09-05 dora_exp_pipeline/combined_plot_script.py
Print the precision at N=num true outliers. #53

@@ -58,6 +58,7 @@ def get_selections_curve(scores, validationLabels):
 
 
 def get_precision_curve(scores, validationLabels):
+    tot_outliers = sum([x for x in validationLabels.values()])
     x = list(range(1, len(scores)+1))
     y = []
     numOutliers = 0
@@ -65,6 +66,9 @@ def get_precision_curve(scores, validationLabels):
         if(validationLabels[scores[i]] == 1):
             numOutliers += 1
         y.append(float(numOutliers)/(i+1))
+        if (i+1) == tot_outliers:
+            print('Precision at N=%d: %f' %
+                  (tot_outliers, float(numOutliers)/(i+1)))
 
     return x, y
 
@@ -94,6 +98,7 @@ def combine_plots(resultsdir, label_path, precision_at_n):
         scores = alg_indexes(filenames[i])
 
         if precision_at_n:
+            print(labels[i])
             x, y = get_precision_curve(scores, validationLabels)
             plt.plot(x, y, label="{}".format(labels[i]))
         else:

### 2021-09-05 publications/iaai21/dora_outlier_detection_iaai.tex
Add results for Earth science time series experiment. #42

--- 2021-09-05 publications/dora_outlier_detection_iaai.tex
@@
 
-\subsubsection{Satellite time series for ground-truth observations}
+\subsubsection{Satellite time series for ground-truth observations.}
 Many Earth science applications using satellite Earth observation (EO) data
@@
 
-\subsubsection{Satellite time series for ground-truth observations}
-Hannah: Describe the Earth crop field data set and applications
+\subsubsection{Satellite time series for ground-truth observations.}
+We used the full dataset of 6,757 samples to fit the ranking algorithms and
+used the independently-labeled set of 113 points in Kenya for evaluation.
+Table \ref{fig:faw_results} compares the performance of each ranking 
+algorithm by plotting the number of true outliers detected out of the top
+$N$ selections. We reported the area under the curve (AUC) in the legend
+for each algorithm to give a quantitative comparison of the curves.
+Table \ref{tab:faw_results} gives the precision at $N$ where
+$N=n_{outlier}=37$. The negative sampling algorithm had the best performance
+in both metrics. RX had the lowest precision at $N=n_{outlier}$ while PCA had
+the lowest AUC of all algorithms except random.
+
+\begin{table}
+  \caption{Precision at $N=n_{outlier}$ for Earth science dataset. }
+  \label{tab:faw_results}
+  \centering
+  \begin{tabular}{cc}
+    \hline
+    Algorithm & Precision at $N=n_{outlier}$ \\
+    \hline
+    RX & 0.30 \\
+    Random & 0.32 \\
+    DEMUD & 0.32 \\
+    PCA & 0.35 \\
+    Isolation Forest & 0.46 \\
+    \textbf{Negative Sampling} & \textbf{0.54} \\
+    \hline
+  \end{tabular}
+\end{table}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{figures/faw_combined_plot.pdf}
+    \caption{Plot of number of true outliers ranked in top $N$ selections for 
+    Earth science dataset.}
+    \label{fig:faw_results}
+\end{figure}
 

### 2021-09-06 publications/iaai21/dora_outlier_detection_iaai.tex
Add to methods, including architecture diagram. #42

@@
 
-\subsection{DORA configurable pipeline} Steven please add details about the 
-approach/design choices for this pipeline. Mention that additional data types, 
-feature extraction methods, ranking algorithms, and results organization and 
-visualization modules can easily be added and integrated if they follow the API. 
-Hannah to add architecture figure. Add descriptions of each layer of modules: 
-\paragraph{Data types} Hannah
-\paragraph{Feature extraction} Hannah 
-\paragraph{Ranking algorithms.} We implemented 7 algorithms for scoring and 
-ranking samples by outlierness. [Describe each algorithm briefly and how they 
-complement each other.]
+\subsection{DORA configurable pipeline} 
+Figure \ref{fig:dora} illustrates the architecture of DORA.
+\todo{Steven please add a couple sentences about the
+approach/design choices for this pipeline.} \todo{Describe the configuration
+file setup. Discuss definitions of data to fit and data to score.
+Maybe include example files in the appendix?}
+We designed DORA to be readily extensible to support additional data types or 
+formats, outlier detection algorithms, and results organization or 
+visualization methods by writing new modules that follow the DORA API.
 
+\paragraph{Data types.} We chose to initially implement data loaders for four
+data types that are commonly used by the machine learning and domain science
+communities: time series, feature vectors, images (grayscale or RGB), and 
+$N$-band rasters. $N$-band rasters are essentially images in which every pixel
+is associated with a location defined by a geographic coordinate reference 
+system (e.g., latitude/longitude in degrees). Most satellite data are 
+distributed as rasters; common formats include GeoTIFF, NetCDF, and 
+\todo{HDF}. A data loader for each data type locates the data by the path(s)
+defined in the configuration file and loads samples into a dictionary of numpy
+ arrays indexed by the sample id. This $data_dict$ (\todo{format as 
+ fixed-width}) is then passed to each of the ranking algorithms specified in 
+ the configuration file.
+
+\paragraph{Outlier ranking algorithms.} We implemented 7 algorithms for 
+scoring and 
+ranking samples by outlierness. We chose these algorithms to include a diverse
+set of approaches to scoring outliers since different algorithms may have 
+better performance for different datasets and use cases. 
+%We describe each
+%approach to scoring outliers and the associated methods below.
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{figures/dora-system-diagram-v5.pdf}
+    \caption{DORA pipeline architecture.}
+    \label{fig:dora}
+\end{figure}
+
+\subparagraph{Reconstruction error.}
+\todo{Describe PCA: Hannah.}
 The DEMUD algorithm~\citep{wagstaff:demud13} differs from other
@@
 
-\paragraph{Results organization} Eric, Umaa, Steven, Vinay
-\paragraph{Outlier visualization} Jake?
+\subparagraph{Distance.}
+\todo{Describe RX and Local RX: Hannah.}
+
+\subparagraph{Sparsity.}
+\todo{Describe isolation forest: Umaa.}
+
+\subparagraph{Likelihood.}
+\todo{Describe PAE: Hannah or Bryce (if it is ready to be included.)}
+\todo{Describe negative sampling: Steven.}
+
+\paragraph{Results organization.} 
+Each of the outlier ranking algorithms returns an array containing the sample
+index, outlier score, and selection index (index of each sample after sorting
+by outlier score). DORA provides results organization modules intended to help
+ users interpret and make decisions based on these outputs. The simplest 
+ module saves a CSV of the samples sorted by their outlier score 
+ (i.e., selection order). Clustering the top
+  $N$ outlier selections can enable users to investigate the different types of 
+ outliers that might be present in the dataset; this could be especially useful
+ for separating outliers caused by noise or data artifacts vs. scientifically 
+ interesting samples. We implemented the K-means and self-organizing maps 
+ (SOMs) algorithms for clustering the top-N outliers. For use cases in which an
+ evaluation dataset containing known outliers is available, we provide a module
+ to assess how well algorithm selections correlate with known outliers. This is
+ done by plotting the number of true outliers vs. number of selections made. 
+ This can also be represented as the precision at $N$ (\todo{add ref?}), i.e., 
+ the number of true outliers divided by the number of selections $N$. Finally,
+ we provide a module for plotting histograms of outlier scores to visualize the
+ distribution of scores in the dataset (which may be, e.g., multimodal or 
+ long-tailed).
+
+\paragraph{Results visualization.} 
+\todo{Jake: describe website if ready by deadline.}
+\todo{Eric: describe causal graphs if added by deadline.}
 
@@
 
-\subsubsection{Astrophysical objects in Dark Energy Survey}
-Umaa
+\subsubsection{Astrophysical objects in Dark Energy Survey.}
+\todo{Umaa}
 
@@
 
-\subsubsection{MNIST and FashionMNIST}
-We added this to have a traditional benchmark and use case demonstration.
+\subsubsection{MNIST and FashionMNIST.}
+We added this to demonstrate the capabilities of DORA with a traditional
+benchmark dataset. \todo{Bryce. TBD if we will end up including due to
+space and/or relevance.}
 
 \subsection{Results}
-Discuss the overall approach to the setup of the experiments. In each 
-subsection discuss specific setup details (including evaluation metrics) for 
-each dataset then the results from the experiments.
+In this section we discuss the experimental setup and results for each dataset.
+We designed the experiments to be relevant to the real-world scientific use 
+cases associated with each dataset domain.
 
-\subsubsection{Astrophysical objects in Dark Energy Survey}
-Umaa
+\subsubsection{Astrophysical objects in Dark Energy Survey.}
+\todo{Umaa}
 
@@
     \includegraphics[width=\linewidth]{figures/faw_combined_plot.pdf}
-    \caption{Plot of number of true outliers ranked in top $N$ selections for 
+    \caption{Number of true outliers ranked in top $N$ selections for 
     Earth science dataset.}
@@
 
-\subsubsection{MNIST and FashionMNIST}
-We added this to have a traditional benchmark and use case demonstration.
+\subsubsection{MNIST and FashionMNIST.}
+\todo{Bryce. TBD if we will end up including due to
+space and/or relevance.}
 
 \section{Discussion}
+\todo{ 
+Potential topics TBD based on results:
 \begin{itemize}
@@
 \item Challenges of evaluating results in outlier detection?
-\item Downstream use of results in science applications/Path to deployment
+\item Downstream use of results in science applications
 \end{itemize}
+}
+\paragraph{Path to deployment.} \todo{Hannah - describe pip-installable
+package, open-source for extension by the wider scientific community, etc.}
 
 \section{Conclusions}
-Summarize conclusions and plans for future work, including public 
-pip-installable package
-
-[Note: these datasets could also be used by other researchers as OOD
+\todo{
+Summarize conclusions and plans for future work
+[Note that these datasets could also be used by other researchers as OOD
  benchmark tasks?]
+}
 
@@
 We thank the UN Food and Agriculture Organization (FAO) and Dr. Ritvik Sahajpal
-(University of Maryland) for providing the fall army worm ground observation
-dataset.
+(University of Maryland/NASA Harvest) for providing the fall army worm ground
+ observation dataset.
 

### 2021-09-09 dora_vis/public/electron.js
DORA visualizer MVP

@@ -1,0 +1,38 @@
+const electron = require('electron');
+const app = electron.app;
+const BrowserWindow = electron.BrowserWindow;
+
+const path = require('path');
+const url = require('url');
+const isDev = require('electron-is-dev');
+
+let mainWindow;
+
+function createWindow() {
+  mainWindow = new BrowserWindow({
+    width: 1024,
+    height: 768,
+    webPreferences: {
+      nodeIntegration: true,
+      nodeIntegrationInWorker: true,
+      contextIsolation: false
+      //webSecurity: false
+    }
+  });
+  mainWindow.loadURL(isDev ? 'http://localhost:3000' : `file://${path.join(__dirname, '../build/index.html')}`);
+  mainWindow.on('closed', () => mainWindow = null);
+}
+
+app.on('ready', createWindow);
+
+app.on('window-all-closed', () => {
+  if (process.platform !== 'darwin') {
+    app.quit();
+  }
+});
+
+app.on('activate', () => {
+  if (mainWindow === null) {
+    createWindow();
+  }
+});

### 2021-09-09 dora_vis/src/App.js

@@ -1,0 +1,329 @@
+import React from 'react';
+//import ReactDOM from 'react-dom';
+import './App.css';
+import DataTable from './Table';
+
+const parse = window.require('csv-parse')
+const yaml = window.require('js-yaml');
+const fs = window.require('fs');
+const path = window.require('path');
+
+const DATALOADERS = ["image"];
+
+
+class NavBar extends React.Component {
+  render() {
+    return(
+      <nav className="navbar navbar-expand-lg navbar-light bg-light">
+        <div className="container-fluid">
+          <span className="navbar-brand mb-0 h1">DORA Visualizer</span>
+          <button className="navbar-toggler" type="button" data-bs-toggle="collapse" data-bs-target="#navbarNav" aria-controls="navbarNav" aria-expanded="false" aria-label="Toggle navigation">
+            <span className="navbar-toggler-icon"></span>
+          </button>
+          <div className="collapse navbar-collapse" id="navbarNav">
+            <ul className="navbar-nav mr-auto">
+              <li className="nav-item">
+                <a className="nav-link active" aria-current="page" href="#">Configure</a>
+              </li>
+              <li className="nav-item">
+                <a className="nav-link" href="#">Table View</a>
+              </li>
+              <li className="nav-time">
+                <a className="nav-link" href="#">Gallery View</a>
+              </li>
+            </ul>
+          </div>
+        </div>
+      </nav>
+    );
+  }
+}
+
+class ConfigLoader extends React.Component {
+  constructor(props) {
+    super(props);
+    this.handleConfigPathChange = this.handleConfigPathChange.bind(this);
+    this.handleConfigPathSubmit = this.handleConfigPathSubmit.bind(this);
+
+    this.state = {configFile: "",
+                  configPath: ""};
+  }
+
+  handleConfigPathChange(e) {
+    /* Handle config file input field change */
+    this.setState({
+      configPath: e.target.files[0].path,
+      configFile: e.target.value
+    });
+  }
+
+  handleConfigPathSubmit(e) {
+    /* Handle submission of file input field */
+    e.preventDefault();
+    this.props.loadConfig(this.state["configPath"]);
+  }
+
+  render() {
+    return(
+      <div>
+        <h1>Load DORA Config</h1>
+        <form>
+          <div className="form-group">
+            <label htmlFor="configPath">Specify path to the DORA Config</label>
[…]

### 2021-09-09 dora_vis/src/Table.js

@@ -1,0 +1,192 @@
+import React from 'react';
+import './App.css';
+
+import { useTable, useSortBy, usePagination} from 'react-table'
+
+const fs = window.require('fs');
+const path = window.require('path');
+
+
+function Table({ columns, data }) {
+  // Use the state and functions returned from useTable to build your UI
+  const {
+    getTableProps,
+    getTableBodyProps,
+    headerGroups,
+    prepareRow,
+    page, // Instead of using 'rows', we'll use page,
+    // which has only the rows for the active page
+
+    // The rest of these things are super handy, too ;)
+    canPreviousPage,
+    canNextPage,
+    pageOptions,
+    pageCount,
+    gotoPage,
+    nextPage,
+    previousPage,
+    setPageSize,
+    state: { pageIndex, pageSize, sortBy },
+  } = useTable(
+    {
+      columns,
+      data,
+      initialState: { pageIndex: 0 },
+    },
+    useSortBy,
+    usePagination
+  )
+
+  // Render the UI for your table
+  return (
+    <>
+      <table {...getTableProps()}>
+        <thead>
+          {headerGroups.map(headerGroup => (
+            <tr {...headerGroup.getHeaderGroupProps()}>
+              {headerGroup.headers.map(column => (
+                <th {...column.getHeaderProps(column.getSortByToggleProps())}>
+                {column.render('Header')}
+                {/* Add a sort direction indicator */}
+                <span>
+                  {column.isSorted
+                    ? column.isSortedDesc
+                      ? ' 🔽'
+                      : ' 🔼'
+                    : ''}
+                </span>
+              </th>
+              ))}
+            </tr>
+          ))}
+        </thead>
+        <tbody {...getTableBodyProps()}>
+          {page.map((row, i) => {
+            prepareRow(row)
+            return (
+              <tr {...row.getRowProps()}>
+                {row.cells.map(cell => {
+                  return <td {...cell.getCellProps()}>{cell.render('Cell')}</td>
+                })}
+              </tr>
+            )
+          })}
+        </tbody>
+      </table>
+      {/* 
+        Pagination can be built however you'd like. 
+        This is just a very basic UI implementation:
+      */}
+      <div className="pagination">
+        <button onClick={() => gotoPage(0)} disabled={!canPreviousPage}>
+          {'<<'}
+        </button>{' '}
+        <button onClick={() => previousPage()} disabled={!canPreviousPage}>
[…]

### 2021-09-09 dora_exp_pipeline/dora_exp.py
Add PAE ranking and image directory data loader (#13)

@@ -10,6 +10,7 @@
 
 import os
 import sys
+import logging
 from tqdm import tqdm
 from dora_exp_pipeline.dora_config import DoraConfig
 from dora_exp_pipeline.dora_data_loader import get_data_loader_by_name
@@ -22,6 +23,7 @@ from dora_exp_pipeline.rx_outlier_detection import RXOutlierDetection
 from dora_exp_pipeline.random_outlier_detection import RandomOutlierDetection
 from dora_exp_pipeline.negative_sampling_outlier_detection import \
     NegativeSamplingOutlierDetection
+from dora_exp_pipeline.pae_outlier_detection import PAEOutlierDetection
 from dora_exp_pipeline.util import LogUtil
 from dora_exp_pipeline.dora_feature import extract_feature
 from dora_exp_pipeline.dora_feature import z_score_normalize
@@ -57,6 +59,10 @@ def register_od_algs():
     negative_sampling_outlier_detection = NegativeSamplingOutlierDetection()
     register_od_alg(negative_sampling_outlier_detection)
 
+    # Register PAE outlier detection algorithm in the pool
+    pae_outlier_detection = PAEOutlierDetection()
+    register_od_alg(pae_outlier_detection)
+
 
 def start(config_file: str, out_dir: str, log_file=None, seed=1234):
     if not os.path.exists(config_file):
@@ -82,6 +88,12 @@ def start(config_file: str, out_dir: str, log_file=None, seed=1234):
         if logger:
             logger.text(f'Created out_dir: '
                         f'{os.path.abspath(config.out_dir)}')
+
+    # Configure tensorflow
+    os.environ['CUDA_VISIBLE_DEVICES'] = '0'
+    os.environ['TF_FORCE_GPU_ALLOW_GROWTH'] = 'True'
+    os.environ['TF_CPP_MIN_LOG_LEVEL'] = '3'
+    logging.getLogger("tensorflow").setLevel(logging.ERROR)
 
     # Register all ranking algorithms supported
     register_od_algs()

### 2021-09-09 setup.py

@@ -22,6 +22,9 @@ setup(
         'pandas==1.1.5',
         'rasterio==1.2.6',
         'tqdm==4.62.0',
+        'tensorflow==2.5.1',
+        'tensorflow-probability==0.13.0',
+        'tensorflow-addons==0.13.0'
         'sklearn-som==1.1.0'
     ],
     provide=[

### 2021-09-09 setup.py
Fix setup.py syntax error; update lint path

@@ -4,7 +4,7 @@ exec(open('dora_exp_pipeline/version.py').read())
 
 setup(
     name='dora_exp_pipeline',
-    version=__version__,
+    version=__version__,  # noqa: F821
     description='DORA Experiment Pipeline',
     url='https://github.com/nasaharvest/dora',
     author='DORA team',
@@ -24,7 +24,7 @@ setup(
         'tqdm==4.62.0',
         'tensorflow==2.5.1',
         'tensorflow-probability==0.13.0',
-        'tensorflow-addons==0.13.0'
+        'tensorflow-addons==0.13.0',
         'sklearn-som==1.1.0'
     ],
     provide=[

### 2021-09-09 publications/iaai21/dora_outlier_detection_iaai.tex
add negative sampling reference #42

@@
    Reed-Xiaoli (RX) detector,  
- Local RX, negative sampling, and probabilistic autoencoder. Given a set of
-  data samples, each algorithm assigns an outlier score to each sample. DORA
-   provides results organization and visualization modules to help users 
-   process the results, including sorting samples by outlier score, evaluating 
-   outlier recall for a set of known/labeled outliers, clustering groups of 
-   similar outliers together, and causal graphs. To demonstrate the 
-   cross-domain applicability of DORA, we conducted outlier detection
-    experiments for three real-world use cases: 1) cleaning noisy ground-truth 
-    datasets (Earth Science), 
-     2) flagging non-astrophysical objects in the Dark Energy Survey 
-     (astrophysics), and 3) novelty-guided onboard targeting of instruments
-      for Mars rovers (planetary science). [We showed that ...]
+ Local RX, negative sampling~\cite{sipple:neg-sampling20}, and probabilistic 
+ autoencoder. Given a set of data samples, each algorithm assigns an outlier 
+ score to each sample. DORA provides results organization and visualization 
+ modules to help users process the results, including sorting samples by outlier 
+ score, evaluating outlier recall for a set of known/labeled outliers, clustering 
+ groups of similar outliers together, and causal graphs. To demonstrate the 
+ cross-domain applicability of DORA, we conducted outlier detection experiments 
+ for three real-world use cases: 1) cleaning noisy ground-truth datasets (Earth 
+ Science), 2) flagging non-astrophysical objects in the Dark Energy Survey 
+ (astrophysics), and 3) novelty-guided onboard targeting of instruments for Mars 
+ rovers (planetary science). [We showed that ...]
 

### 2021-09-10 publications/iaai21/dora_outlier_detection_iaai.tex
add dora configurable pipeline and negative sampling paragraphs #42

@@
 \subsection{DORA configurable pipeline} 
-Figure \ref{fig:dora} illustrates the architecture of DORA.
-\todo{Steven please add a couple sentences about the
-approach/design choices for this pipeline.} \todo{Describe the configuration
-file setup. Discuss definitions of data to fit and data to score.
-Maybe include example files in the appendix?}
-We designed DORA to be readily extensible to support additional data types or 
-formats, outlier detection algorithms, and results organization or 
-visualization methods by writing new modules that follow the DORA API.
+Figure \ref{fig:dora} illustrates the architecture of DORA. It includes 
+data loading, feature extraction, outlier ranking, results organization, and 
+results visualizaiton modules. The DORA pipeline is implemented in Python 3.
+In order to improve the readability and execution speed of the code, we adopted
+object-oriented and functional programming practices.  We designed DORA to be 
+readily extensible to support additional data types or formats, outlier detection 
+algorithms, and results organization or visualization methods by writing new 
+modules that follow the DORA API. Experimental settings are controlled by 
+configuration files. In a configuration file, users can specify the input data, 
+feature extraction methods, normalization method, outlier ranking methods, and 
+results organization methods.
 
@@
 \todo{Describe PAE: Hannah or Bryce (if it is ready to be included.)}
-\todo{Describe negative sampling: Steven.}
+The negative sampling algorithm is implemented by converting the unsupervised 
+outlier ranking problem into a semi-supervised problem. Negative (anomalous) 
+examples are created by sampling from a expanded space defined by the minimum 
+and maximum values of each dimension of the positive (normal) examples. The 
+negative and positive examples are then used to train a random forest 
+classifier. We use the posterior probabilities of the random forest classifier 
+as outlier scores, which means that the observations with higher posterior 
+probabilities are more likely to be outliers.
 

### 2021-09-10 publications/iaai21/dora_outlier_detection_iaai.tex

@@
 
-\subsection{DORA configurable pipeline} 
+%\subsection{DORA configurable pipeline} 
 Figure \ref{fig:dora} illustrates the architecture of DORA. It includes 
@@
 defined in the configuration file and loads samples into a dictionary of numpy
- arrays indexed by the sample id. This $data_dict$ (\todo{format as 
- fixed-width}) is then passed to each of the ranking algorithms specified in 
+ arrays indexed by the sample id. This \texttt{data\_dict} is then passed to each of the ranking algorithms specified in 
  the configuration file.
@@
 \subparagraph{Reconstruction error.}
+Reconstruction-based methods aim to construct a model of a dataset by learning 
+a mapping between the data samples and a lower-dimensional representation that
+ minimizes the loss between the input and its reconstruction from the 
+ low-dimensional representation~\citep{kerner2020comparison}. The 
+ reconstruction error is used to compute the outlier score because samples
+  that are unlike the data used to fit the model will be more poorly 
+  reconstructed compared to inliers. 
 \todo{Describe PCA: Hannah.}

### 2021-09-10 dora_exp_pipeline/combined_plot_script.py
Normalize AUC by number of total selections. #53

@@ -82,13 +82,13 @@ def combine_plots(resultsdir, label_path, precision_at_n):
     num_outliers = sum([x for x in validationLabels.values()])
 
     if precision_at_n:
-        plt.axhline(y=1, label='Best line', linestyle='--', color='k')
+        plt.axhline(y=1, label='Oracle', linestyle='--', color='k')
         plt.title('Precision at N')
         plt.xlabel('Number of selections (N)')
         plt.ylabel('Precision at N')
     else:
         bestLine = list(range(1, num_outliers+1))
-        plt.plot(bestLine, bestLine, label='Best line', linestyle='--',
+        plt.plot(bestLine, bestLine, label='Oracle', linestyle='--',
                  color='k')
         plt.title('True Outliers vs. Algorithm Selections')
         plt.xlabel('Number of selections')
@@ -104,8 +104,8 @@ def combine_plots(resultsdir, label_path, precision_at_n):
         else:
             x, y = get_selections_curve(scores, validationLabels)
             index = x.index(y[-1])
-            area = np.trapz(y[:index+1], x[:index+1])
-            plt.plot(x, y, label="{} (AUC: {})".format(labels[i], area))
+            area = np.trapz(y[:index+1], x[:index+1])/len(scores)
+            plt.plot(x, y, label="{} (AUC: {:.2f})".format(labels[i], area))
 
     axes.set_xlim(1, len(scores))
     plt.legend()

### 2021-09-10 dora_exp_pipeline/dora_results_organization.py

@@ -111,11 +111,11 @@ class SaveComparisonPlot(ResultsOrganization):
 
         fig, axes = plt.subplots()
         index = x.index(y[-1])
-        area = np.trapz(y[:index+1], x[:index+1])
+        area = np.trapz(y[:index+1], x[:index+1])/len(labels)
 
-        plt.plot(x, y, label=outlier_alg_name)
-        plt.plot(x, x, label='Best Line')
-        plt.plot([], [], ' ', label=f'Area: {area}')
+        plt.plot(x, y, label="{} (AUC: {:.2f})".format(outlier_alg_name, area),
+                 linestyle='--', color='k')
+        plt.plot(x, x, label='Oracle')
         plt.title('Correct Outliers vs Selected Outliers')
         plt.xlabel('Number of Outliers Selected')
         plt.ylabel('Number of True Outliers')

### 2021-09-10 dora_exp_pipeline/combined_plot_script.py
Sort by selection index instead of score. #53

@@ -1,6 +1,4 @@
 import os
-# import sys
-# sys.path.append('.')
 import numpy as np
 import matplotlib.pyplot as plt
 import argparse
@@ -9,7 +7,7 @@ import argparse
 def alg_indexes(filename):
     with open(filename, 'r') as f:
         text = f.read().split("\n")[:-1]
-        scores = sorted(text, key=lambda x: float(x.split(", ")[-1]))
+        scores = sorted(text, key=lambda x: float(x.split(", ")[0]))
         scores = [int(i.split(", ")[1]) for i in scores]
 
     return scores

### 2021-09-10 dora_exp_pipeline/dora_results_organization.py

@@ -113,9 +113,8 @@ class SaveComparisonPlot(ResultsOrganization):
         index = x.index(y[-1])
         area = np.trapz(y[:index+1], x[:index+1])/len(labels)
 
-        plt.plot(x, y, label="{} (AUC: {:.2f})".format(outlier_alg_name, area),
-                 linestyle='--', color='k')
-        plt.plot(x, x, label='Oracle')
+        plt.plot(x, y, label="{} (AUC: {:.2f})".format(outlier_alg_name, area))
+        plt.plot(x, x, label='Oracle', linestyle='--', color='k')
         plt.title('Correct Outliers vs Selected Outliers')
         plt.xlabel('Number of Outliers Selected')
         plt.ylabel('Number of True Outliers')

### 2021-09-10 publications/iaai21/dora_outlier_detection_iaai.tex
Update results for Earth exp with corrected plots and PAE. #42

@@
 algorithm by plotting the number of true outliers detected out of the top
-$N$ selections. We reported the area under the curve (AUC) in the legend
+$N$ selections. We reported the area under the curve (AUC), normalized
+by the total number of selections, in the legend
 for each algorithm to give a quantitative comparison of the curves.
@@
 $N=n_{outlier}=37$. The negative sampling algorithm had the best performance
-in both metrics. RX had the lowest precision at $N=n_{outlier}$ while PCA had
-the lowest AUC of all algorithms except random.
+in both metrics. DEMUD, PCA, and PAE tied for the lowest precision at 
+$N=n_{outlier}$ while DEMUD and PCA tied for the lowest AUC.
 
@@
     \hline
-    RX & 0.30 \\
     Random & 0.32 \\
-    DEMUD & 0.32 \\
-    PCA & 0.35 \\
+    DEMUD & 0.41 \\
+    PCA & 0.41 \\
+    PAE & 0.41 \\
+    RX & 0.43 \\
     Isolation Forest & 0.46 \\
-    \textbf{Negative Sampling} & \textbf{0.54} \\
+    \textbf{Negative Sampling} & \textbf{0.49} \\
     \hline

### 2021-09-10 dora_exp_pipeline/dora_results_organization.py
Use dts_sels to index into labels. (#23)

@@ -98,14 +98,13 @@ class SaveComparisonPlot(ResultsOrganization):
 
         # Outliers will be 1s and inliers will be 0s.
         labels = self._get_validation_labels(validation_dir)
-        scores = np.argsort(dts_scores)[::-1]
 
-        x = list(range(1, len(scores)+1))
+        x = list(range(1, len(dts_sels)+1))
         y = []
         numOutliers = 0
 
-        for i in range(len(scores)):
-            if(labels[scores[i]] == 1):
+        for s in dts_sels:
+            if (labels[s] == 1):
                 numOutliers += 1
             y.append(numOutliers)
 

### 2021-09-13 dora_exp_pipeline/combined_plot_script.py
Update comparison plots to report MDR instead of AUC. #53

@@ -101,9 +101,8 @@ def combine_plots(resultsdir, label_path, precision_at_n):
             plt.plot(x, y, label="{}".format(labels[i]))
         else:
             x, y = get_selections_curve(scores, validationLabels)
-            index = x.index(y[-1])
-            area = np.trapz(y[:index+1], x[:index+1])/len(scores)
-            plt.plot(x, y, label="{} (AUC: {:.2f})".format(labels[i], area))
+            area = sum(y)/sum([i for i in range(len(scores))])
+            plt.plot(x, y, label="{} (MDR: {:.2f})".format(labels[i], area))
 
     axes.set_xlim(1, len(scores))
     plt.legend()

### 2021-09-13 dora_exp_pipeline/dora_results_organization.py

@@ -104,16 +104,15 @@ class SaveComparisonPlot(ResultsOrganization):
         y = []
         numOutliers = 0
 
-        for s in dts_sels:
-            if (labels[s] == 1):
+        for i in range(len(dts_sels)):
+            if(labels[dts_sels[i]] == 1):
                 numOutliers += 1
             y.append(numOutliers)
 
         fig, axes = plt.subplots()
-        index = x.index(y[-1])
-        area = np.trapz(y[:index+1], x[:index+1])/len(labels)
+        area = sum(y)/sum([i for i in range(len(labels))])
 
-        plt.plot(x, y, label="{} (AUC: {:.2f})".format(outlier_alg_name, area))
+        plt.plot(x, y, label="{} (MDR: {:.2f})".format(outlier_alg_name, area))
         plt.plot(x, x, label='Oracle', linestyle='--', color='k')
         plt.title('Correct Outliers vs Selected Outliers')
         plt.xlabel('Number of Outliers Selected')

### 2021-09-13 dora_exp_pipeline/combined_plot_script.py
Remove unused numpy import. #53

@@ -1,5 +1,4 @@
 import os
-import numpy as np
 import matplotlib.pyplot as plt
 import argparse
 

### 2021-09-14 publications/iaai21/dora_outlier_detection_iaai.tex

@@
  (astrophysics), and 3) novelty-guided onboard targeting of instruments for Mars 
- rovers (planetary science). [We showed that ...]
+ rovers (planetary science). \todo{[We showed that ...]}
 
@@
 scoring and 
-ranking samples by outlierness. We chose these algorithms to include a diverse
+ranking samples by outlierness. All algorithms are unsupervised.
+We chose these algorithms to include a diverse
 set of approaches to scoring outliers since different algorithms may have 
@@
   reconstructed compared to inliers. 
-\todo{Describe PCA: Hannah.}
+Principal component analysis (PCA) has been used for outlier detection by
+scoring samples using the reconstruction error (here, the L2 norm)
+between inputs and their inverse
+ transformation from the principal subspace~\citep{kerner2020comparison}.
 The DEMUD algorithm~\citep{wagstaff:demud13} differs from other
@@
 \subparagraph{Distance.}
-\todo{Describe RX and Local RX: Hannah.}
+Distance-based methods score outliers based on their distance from some
+``background'' which is usually defined by a sample distribution. The 
+Reed-Xiaoli (RX) detector is commonly used for anomaly detection in
+multispectral and hyperspectral remote sensing; RX uses the Mahalanobis
+ distance between a sample and a background dataset mean and covariance
+ to score samples~\citep{reed1990adaptive}. The local variant of RX (Local
+ RX or LRX) can be used for image or raster data types and scores each pixel in
+  an image with respect to a window ``ring'' of pixels surrounding 
+  it~\citep{molero2013analysis}. 
+  LRX requires two parameters that define the ring: an outer window which 
+  defines the size of the window surrounding the pixel and an inner window
+   which defines the size of the window around the target pixel to exclude from
+   the background distribution.
 
@@
 \subparagraph{Likelihood.}
-\todo{Describe PAE: Hannah or Bryce (if it is ready to be included.)}
+\todo{Describe PAE: Hannah or Bryce}
 The negative sampling algorithm is implemented by converting the unsupervised 
-outlier ranking problem into a semi-supervised problem. Negative (anomalous) 
-examples are created by sampling from a expanded space defined by the minimum 
+outlier ranking problem into a semi-supervised 
+problem~\citep{sipple:neg-sampling20}. Negative (anomalous) 
+examples are created by sampling from an expanded space defined by the minimum 
 and maximum values of each dimension of the positive (normal) examples. The 
@@
  done by plotting the number of true outliers vs. number of selections made. 
- This can also be represented as the precision at $N$ (\todo{add ref?}), i.e., 
- the number of true outliers divided by the number of selections $N$. Finally,
+ This can also be represented as the precision at $N$, i.e., 
+ the number of true outliers divided by the number of selections $N$~\citep{campos2016evaluation}. Finally,
  we provide a module for plotting histograms of outlier scores to visualize the

### 2021-09-14 publications/iaai21/dora_outlier_detection_iaai.tex
Add precision-at-k results for planetary rover. (#60,#42)
Add selection figure for planetary rover results. (#60,#42)

@@
 \begin{table}
-  \caption{Precision at $N=n_{outlier}$ for Earth science dataset. }
+  \caption{Precision at $N=n_{outlier}$ for three applications; the
+  best result for each data set is in bold.}
   \label{tab:faw_results}
   \centering
-  \begin{tabular}{cc}
+  \begin{tabular}{l|ccc}
     \hline
-    Algorithm & Precision at $N=n_{outlier}$ \\
+    Algorithm & Earth & Planetary & Astrophysics \\
     \hline
-    Random & 0.32 \\
-    DEMUD & 0.41 \\
-    PCA & 0.41 \\
-    PAE & 0.41 \\
-    RX & 0.43 \\
-    Isolation Forest & 0.46 \\
-    \textbf{Negative Sampling} & \textbf{0.49} \\
+    PCA               & 0.41 & 0.44 \\
+    DEMUD             & 0.41 & 0.44 \\
+    \hline
+    RX                & 0.43 & ---- \\
+    LRX               & ---- & 0.33 \\
+    \hline
+    Isolation Forest  & 0.46 & \textbf{0.56} \\
+    \hline
+    PAE               & 0.41 & 0.22 \\
+    Neg. Sampling & \textbf{0.49} & 0.33 \\
+    \hline
+    Random            & 0.32 & 0.11 \\
     \hline
@@
     \label{fig:faw_results}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{figures/planetary_combined_plot.pdf}
+    \caption{Number of true outliers ranked in top $N$ selections for 
+    planetary rover dataset.}
+    \label{fig:planetary_results}
 \end{figure}

### 2021-09-14 publications/iaai21/dora_outlier_detection_iaai.tex
Update details describing planetary rover data set. (#42)

@@
 investigation~\cite{kerner2020comparison,wagstaff:rover-novelty20}.
-We collected \todo{XXX} images from the Navigation camera (Navcam) 
-onboard the Mars Science Laboratory (MSL) rover
+We collected \num{72} images from the Navigation camera (Navcam) 
+on the Mars Science Laboratory (MSL) rover
 and employed the Rockster~\cite{burl:rockster16}
 method currently used by onboard rover software to identify candidate rock
-targets with an area of at least 100 pixels, yielding \num{6005}
+targets with an area of at least 100 pixels, yielding \num{1050}
 targets.  We cropped out a \num{64}$\times$\num{64} pixel image
 centered on each target.
-%
+
+We simulated the operational setting in which the rover has observed
+targets up through mission day (sol) $s$ and the goal is to rank all
+subsequent targets (after sol $s$) to inform which recent targets
+merit further study.  We organized the Mars target images
+chronologically and partitioned them into the corresponding ``prior''
+set $D_s$ (sols \num{1343} to \num{1693}, $n=992$) and ``assessment''
+set $D_{s+}$ (sols \num{1694} to \num{1703}, $n=58)$.
+
 We collaborated with an MSL science planning team member to
-independently review the targets and identify those that were
-considered novel by the mission ($n_{novel} = 108$).  Our goal for
+independently review the targets in $D_{s+}$ and identify those that were
+considered novel by the mission ($n_{outlier} = 9$).  Our goal for
 this application is to assess how well the selections made by each
@@
 \subsubsection{Targets in Mars rover images.}
-We simulated the operational setting in which the rover has observed
-targets up through mission day (sol) $s$ and the goal is to rank all
-subsequent targets (after sol $s$) to inform which recent targets
-merit further study.  We organized the Mars target images
-chronologically and partitioned them into the corresponding ``prior''
-set $D_s$ and ``assessment'' set $D_{s+}$. We initialized each
-algorithm with $D_s$ and requested a ranking of targets in $D_{s+}$.
+% We initialized each algorithm with $D_s$ and requested a ranking of
+% targets in $D_{s+}$. 
 \todo{Add table showing number (or percentage?) of novel targets in

### 2021-09-14 publications/iaai21/dora_outlier_detection_iaai.tex

@@
 subsequent targets (after sol $s$) to inform which recent targets
-merit further study.  We organized the Mars target images
-chronologically and partitioned them into the corresponding ``prior''
-set $D_s$ (sols \num{1343} to \num{1693}, $n=992$) and ``assessment''
-set $D_{s+}$ (sols \num{1694} to \num{1703}, $n=58)$.
+merit further study.  Our rover image data covers sols \num{1343}
+to \num{1703}.  We partitioned the images chronologically to assess
+outlier detection in the 10 most recent sols, using ``prior'' set
+$D_{1343-1693}$ ($n=992$) and ``assessment'' set $D_{1694-1703}$
+($n=58)$ for evaluation.
 
 We collaborated with an MSL science planning team member to
-independently review the targets in $D_{s+}$ and identify those that were
+independently review the targets in $D_{1694-1703}$ and identify those
+that were 
 considered novel by the mission ($n_{outlier} = 9$).  Our goal for
@@
 % targets in $D_{s+}$.
-\todo{Add table showing number (or percentage?) of novel targets in
-the top $N$ selections, where $N$ = known number of novel targets in
-$D_{s+}$.  And/or, report average rank of novel targets if alg ranks all
-$|D_{s+}|$ targets. If time/room permits, do this for different values
-of $s$.}
+\todo{Consider merging all three data set discussions since they are
+in one table? with their own figures? The point here isn't really to
+compare algs or data sets but instead to show that we CAN compare
+them, using DORA.}
+%
+Table~\ref{tab:faw_results} compares the performance of each ranking 
+algorithm by reporting the precision obtained when the number of
+selections is the same as the total number of outliers.  We could not
+employ RX in this case since the data dimensionality ($64 \times 64 =
+4096$) exceeded the data set size.  We found that
+the Isolation Forest achieved the highest precision (best outlier
+detection) when allowed to select only \num{9} images.
+Figure~\ref{fig:planetary_results} shows the complete (cumulative)
+outlier detection performance for each algorithm when ranking
+all \num{58} target images in $D_{1694-1703}$. \todo{Explain MDR the first
+time it appears. }
 
@@
 in both metrics. DEMUD, PCA, and PAE tied for the lowest precision at 
-$N=n_{outlier}$ while DEMUD and PCA tied for the lowest AUC.
+$N=n_{outlier}$ while DEMUD and PCA tied for the lowest
+AUC. \todo{For completeness, Explain why no LRX here.}
 
@@
     \hline
-    Algorithm & Earth & Planetary & Astrophysics \\
+    Algorithm & Astrophysics & Earth & Planetary \\
     \hline
-    PCA               & 0.41 & 0.44 \\
-    DEMUD             & 0.41 & 0.44 \\
+    PCA               & & 0.41 & 0.44 \\
+    DEMUD             & & 0.41 & 0.44 \\
     \hline
-    RX                & 0.43 & ---- \\
-    LRX               & ---- & 0.33 \\
+    RX                & & 0.43 & ---- \\
+    LRX               & & ---- & 0.33 \\
     \hline
-    Isolation Forest  & 0.46 & \textbf{0.56} \\
+    Isolation Forest  & & 0.46 & \textbf{0.56} \\
     \hline
-    PAE               & 0.41 & 0.22 \\
-    Neg. Sampling & \textbf{0.49} & 0.33 \\
+    PAE               & & 0.41 & 0.22 \\
+    Neg. Sampling & & \textbf{0.49} & 0.33 \\
     \hline
-    Random            & 0.32 & 0.11 \\
+    Random            & & 0.32 & 0.11 \\
     \hline
@@
 \end{table}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{figures/planetary_combined_plot.pdf}
+    \caption{Number of true outliers ranked in top $N$ selections for 
+    planetary rover dataset.}
+    \label{fig:planetary_results}
+\end{figure}
 
@@
     \label{fig:faw_results}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{figures/planetary_combined_plot.pdf}
-    \caption{Number of true outliers ranked in top $N$ selections for 
-    planetary rover dataset.}
-    \label{fig:planetary_results}
 \end{figure}

### 2021-09-14 dora_vis/public/electron.js
Working navbar and viewstate handling

@@ -16,7 +16,6 @@ function createWindow() {
       nodeIntegration: true,
       nodeIntegrationInWorker: true,
       contextIsolation: false
-      //webSecurity: false
     }
   });
   mainWindow.loadURL(isDev ? 'http://localhost:3000' : `file://${path.join(__dirname, '../build/index.html')}`);

### 2021-09-14 dora_vis/src/App.js

@@ -12,7 +12,52 @@ const DATALOADERS = ["image"];
 
 
 class NavBar extends React.Component {
+  constructor(props) {
+    super(props);
+
+    this.handleNavClick = this.handleNavClick.bind(this);
+  }
+
+  handleNavClick(e) {
+    e.preventDefault();
+    switch(e.target.innerText) {
+      case "Configure":
+        this.props.changeView("loadConfig");
+        break;
+      case "Aggregate View":
+        this.props.changeView("aggTable");
+        break;
+      case "Table View":
+        this.props.changeView("dataTable");
+        break;
+      case "Gallery View":
+        this.props.changeView("gallery");
+        break;
+      default:
+        break;
+    }
+  }
+
   render() {
+    let configureActive = "";
+    let aggregateActive = "";
+    let tableActive = "";
+    let galleryActive = "";
+    if (this.props.currView === "loadConfig" || this.props.currView === "parseConfig") {
+      configureActive = "active";
+      aggregateActive = tableActive = galleryActive = "";
+    } else if (this.props.currView === "aggTable") {
+      aggregateActive = "active";
+      configureActive = tableActive = galleryActive = "";
+    } else if (this.props.currView === "dataTable") {
+      tableActive = "active";
+      configureActive = aggregateActive = galleryActive = "";
+    } else if (this.props.currView === "gallery"){
+      galleryActive = "active";
+      configureActive = aggregateActive = tableActive = "";
+    } else {
+      configureActive = aggregateActive = tableActive = galleryActive = "";
+    }
     return(
       <nav className="navbar navbar-expand-lg navbar-light bg-light">
         <div className="container-fluid">
@@ -23,13 +68,16 @@ class NavBar extends React.Component {
           <div className="collapse navbar-collapse" id="navbarNav">
             <ul className="navbar-nav mr-auto">
               <li className="nav-item">
-                <a className="nav-link active" aria-current="page" href="#">Configure</a>
+                <a className={"nav-link " + configureActive} href="#" onClick={this.handleNavClick}>Configure</a>
               </li>
               <li className="nav-item">
-                <a className="nav-link" href="#">Table View</a>
+                <a className={"nav-link " + aggregateActive} href="#" onClick={this.handleNavClick}>Aggregate View</a>
+              </li>
+              <li className="nav-item">
[…]

### 2021-09-14 dora_vis/src/Table.js

@@ -137,11 +137,13 @@ class DataTable extends React.Component {
   }
 
   componentDidMount() {
-    this.props.loadData(this.props.configData["outlier_detection"], Object.keys(this.props.configData["outlier_detection"])[0]);
-    this.setState({
-      methods: Object.keys(this.props.configData["outlier_detection"]),
-      currMethod: Object.keys(this.props.configData["outlier_detection"])[0]
-    });
+    if (this.props.configData != null) {
+      this.props.loadData(this.props.configData["outlier_detection"], Object.keys(this.props.configData["outlier_detection"])[0]);
+      this.setState({
+        methods: Object.keys(this.props.configData["outlier_detection"]),
+        currMethod: Object.keys(this.props.configData["outlier_detection"])[0]
+      });
+    }
   }
 
   switchMethod(e) {
@@ -176,13 +178,22 @@ class DataTable extends React.Component {
       }
     ];
 
-    return (
-    <div>
-      <h1>{this.state["currMethod"]}</h1>
+    let datapass = null;
+    if (this.props.data == null) {
+      datapass = <h1> No DORA configuration loaded </h1>;
+    } else {
+      datapass = <>
       <Table columns={columns} data={this.props.data}/>
       <select onChange={this.switchMethod}>
         {this.state["methods"].map((method) => <option value={method}>{method}</option>)}
       </select>
+      </>;
+    }
+
+    return (
+    <div className="container-fluid">
+      <h1>{this.state["currMethod"]}</h1>
+      {datapass}
     </div>
     );
     

### 2021-09-14 dora_vis/src/Table.js
Moved filename to image hovertext

@@ -162,13 +162,9 @@ class DataTable extends React.Component {
         accessor: "id"
       },
       {
-        Header: "Filename",
-        accessor: "fileName"
-      },
-      {
         Header: "Image",
         Cell: (row) => {
-          return <div><img height={100} src={"data:image/png;base64,"+row.row.original.imageData}/></div>;
+          return <div><img height={100} src={"data:image/png;base64,"+row.row.original.imageData} title={row.row.original.fileName}/></div>;
         },
         id: "image"
       },

### 2021-09-14 dora_vis/src/Table.js
Clicking image copies filename to clipboard

@@ -164,7 +164,13 @@ class DataTable extends React.Component {
       {
         Header: "Image",
         Cell: (row) => {
-          return <div><img height={100} src={"data:image/png;base64,"+row.row.original.imageData} title={row.row.original.fileName}/></div>;
+          return (
+          <div>
+            <a href="javascript:undefined;" onClick={() => {navigator.clipboard.writeText(row.row.original.fileName);alert(row.row.original.fileName+ " copied to clipboard.");}}>
+              <img height={100} src={"data:image/png;base64,"+row.row.original.imageData} title={row.row.original.fileName}/>
+            </a>
+          </div>
+          );
         },
         id: "image"
       },

### 2021-09-15 publications/iaai21/dora_outlier_detection_iaai.tex
Add isolation forest description. #42

@@
 distributed as rasters; common formats include GeoTIFF, NetCDF, and 
-\todo{HDF}. A data loader for each data type locates the data by the path(s)
+HDF. A data loader for each data type locates the data by the path(s)
 defined in the configuration file and loads samples into a dictionary of numpy
@@
 \subparagraph{Sparsity.}
-\todo{Describe isolation forest: Umaa.}
+Sparsity-based methods score outliers based on how isolated or sparse they
+are in the feature space. Isolation forests are a common sparsity-based method
+that constructs many random binary trees from a 
+dataset~\cite{liu2008isolation}. The outlier score for
+a sample is quantified as the average distance from the root to the item’s 
+leaf. Shorter distances are indicative of outliers because fewer splits were 
+required to isolate the sample.
 
 \subparagraph{Likelihood.}
-\todo{Describe PAE: Hannah or Bryce}
 The negative sampling algorithm is implemented by converting the unsupervised 
@@
 as outlier scores, which means that the observations with higher posterior 
-probabilities are more likely to be outliers.
+probabilities are more likely to be outliers. \todo{Describe PAE: Hannah or Bryce}
 

### 2021-09-15 dora_vis/src/Table.js
Table formatting and styling

@@ -1,5 +1,5 @@
 import React from 'react';
-import './App.css';
+import './Table.css';
 
 import { useTable, useSortBy, usePagination} from 'react-table'
 
@@ -40,20 +40,22 @@ function Table({ columns, data }) {
   // Render the UI for your table
   return (
     <>
-      <table {...getTableProps()}>
+      <table className="table table-striped" {...getTableProps()}>
         <thead>
           {headerGroups.map(headerGroup => (
-            <tr {...headerGroup.getHeaderGroupProps()}>
+            <tr className="d-flex" {...headerGroup.getHeaderGroupProps()}>
               {headerGroup.headers.map(column => (
-                <th {...column.getHeaderProps(column.getSortByToggleProps())}>
+                <th {...column.getHeaderProps({...column.getSortByToggleProps(),...{className: column.headerClassName}})}>
                 {column.render('Header')}
                 {/* Add a sort direction indicator */}
                 <span>
-                  {column.isSorted
-                    ? column.isSortedDesc
-                      ? ' 🔽'
-                      : ' 🔼'
-                    : ''}
+                  {column.canSort
+                    ? column.isSorted
+                      ? column.isSortedDesc
+                        ? ' 🔽'
+                        : ' 🔼'
+                      : ' ⏺️'
+                    : '  '}
                 </span>
               </th>
               ))}
@@ -64,9 +66,9 @@ function Table({ columns, data }) {
           {page.map((row, i) => {
             prepareRow(row)
             return (
-              <tr {...row.getRowProps()}>
+              <tr className="d-flex" {...row.getRowProps()}>
                 {row.cells.map(cell => {
-                  return <td {...cell.getCellProps()}>{cell.render('Cell')}</td>
+                  return <td {...cell.getCellProps({className: cell.column.className})}>{cell.render('Cell')}</td>
                 })}
               </tr>
             )
@@ -78,16 +80,16 @@ function Table({ columns, data }) {
         This is just a very basic UI implementation:
       */}
       <div className="pagination">
-        <button onClick={() => gotoPage(0)} disabled={!canPreviousPage}>
+        <button type="button" className="btn btn-outline-dark btn-sm" onClick={() => gotoPage(0)} disabled={!canPreviousPage}>
           {'<<'}
         </button>{' '}
-        <button onClick={() => previousPage()} disabled={!canPreviousPage}>
[…]

### 2021-09-15 dora_vis/src/App.js
Move method selection to top

@@ -117,10 +117,12 @@ class ConfigLoader extends React.Component {
       <div className="container-fluid">
         <h1>Load DORA Config</h1>
         <form>
-          <div className="form-group">
-            <label htmlFor="configPath">Specify path to the DORA Config</label>
-            <input type="file" className="form-control" id="configPath" onChange={this.handleConfigPathChange} value={this.state["configFile"]} required />
-            <button type="submit" className="btn btn-primary" onClick={this.handleConfigPathSubmit}>Submit</button>
+          <div className="col-md-6">
+            <div className="form-group">
+              <label htmlFor="configPath">Specify path to the DORA Config</label>
+              <input type="file" className="form-control" id="configPath" onChange={this.handleConfigPathChange} value={this.state["configFile"]} required />
+              <button type="submit" className="btn btn-primary" onClick={this.handleConfigPathSubmit}>Submit</button>
+            </div>
           </div>
         </form>
       </div>
@@ -233,10 +235,12 @@ class ConfigParser extends React.Component {
       componentDataRoot = <label>{this.state["dataRootPath"]} is a valid data root directory.</label>;
     } else {
       componentDataRoot = 
+      <div className="col-md-6">
       <div className="form-group">
         <label htmlFor="dataRootField">Can't locate {this.props.configData["data_to_score"]}, Specify the root directory of input data</label>
         <input type="file" className="form-control" id="dataRootField" onChange={this.handleDataRootChange} value={this.state["dataRoot"]} required directory="" webkitdirectory=""/>
         <button type="submit" className="btn btn-primary" onClick={this.handleDataRootSubmit}>Submit</button>
+      </div>
       </div>;
     }
 
@@ -246,10 +250,12 @@ class ConfigParser extends React.Component {
       componentOutDir = <label>{this.state["outDirPath"]} is a valid output directory.</label>;
     } else {
       componentOutDir = 
+      <div className="col-md-6">
       <div className="form-group">
         <label htmlFor="outDirField">Can't locate {this.props.configData["out_dir"]}, Specify the root directory of DORA results.</label>
         <input type="file" className="form-control" id="outDirField" onChange={this.handleOutDirChange} value={this.state["outDir"]} required directory="" webkitdirectory=""/>
         <button type="submit" className="btn btn-primary" onClick={this.handleOutDirSubmit}>Submit</button>
+      </div>
       </div>;
     }
 

### 2021-09-15 dora_vis/src/Table.js

@@ -201,18 +201,22 @@ class DataTable extends React.Component {
     if (this.props.data == null) {
       datapass = <h1> No DORA configuration loaded </h1>;
     } else {
-      datapass = <>
+      datapass = 
+      <>
+      <div className="col-md-4 methodSelect">
+        <label for="methodSel">Results from method:</label>
+        <select className="form-select form-select-lg" id="methodSel" onChange={this.switchMethod}>
+          {this.state["methods"].map((method) => <option value={method}>{method}</option>)}
+        </select>
+      </div>
       <Table columns={columns} data={this.props.data}/>
-      <select onChange={this.switchMethod}>
-        {this.state["methods"].map((method) => <option value={method}>{method}</option>)}
-      </select>
       </>;
     }
 
     return (
     <div className="container">
-      <h1>{this.state["currMethod"]}</h1>
       {datapass}
+      <br/>
     </div>
     );
     

### 2021-09-15 dora_vis/src/Table.js
Fixed table sorting

@@ -159,13 +159,15 @@ class DataTable extends React.Component {
         Header: "Rank",
         accessor: "rank",
         className: "rankCell",
-        headerClassName: "rankHeader"
+        headerClassName: "rankHeader",
+        sortType: "number"
       },
       {
         Header: "ID",
         accessor: "id",
         className: "idCell",
-        headerClassName: "idHeader"
+        headerClassName: "idHeader",
+        sortType: "number"
       },
       {
         Header: "Image",
@@ -193,7 +195,8 @@ class DataTable extends React.Component {
         },
         accessor: "score",
         className: "scoreCell",
-        headerClassName: "scoreHeader"
+        headerClassName: "scoreHeader",
+        sortType: "number"
       }
     ];
 

### 2021-09-15 publications/iaai21/dora_outlier_detection_iaai.tex
Update PAE results and figure for planetary rover (#13,#60).
Add notes to Hannah in todos. (#42)

@@
 % targets in $D_{s+}$.
-\todo{Consider merging all three data set discussions since they are
+\todo{Hannah: Consider merging all three data set discussions since they are
 in one table? with their own figures? The point here isn't really to
@@
 outlier detection performance for each algorithm when ranking
-all \num{58} target images in $D_{1694-1703}$. \todo{Explain MDR the first
-time it appears. }
+all \num{58} target images in $D_{1694-1703}$. \todo{Hannah: Explain
+MDR the first time it appears. }
 
@@
 $N=n_{outlier}$ while DEMUD and PCA tied for the lowest
-AUC. \todo{For completeness, Explain why no LRX here.}
+AUC. \todo{Hannah: Update AUC -> MDR and for completeness, explain why
+no LRX here.}
 
@@
     \hline
-    PAE               & & 0.41 & 0.22 \\
+    PAE               & & 0.41 & 0.44 \\
     Neg. Sampling & & \textbf{0.49} & 0.33 \\

### 2021-09-15 dora_vis/src/App.js
Aggregate Table View

@@ -1,7 +1,7 @@
 import React from 'react';
 //import ReactDOM from 'react-dom';
 import './App.css';
-import DataTable from './Table';
+import { DataTable, AggTable } from './Table';
 
 const parse = window.require('csv-parse')
 const yaml = window.require('js-yaml');
@@ -288,16 +288,18 @@ class App extends React.Component {
     this.loadConfig = this.loadConfig.bind(this);
     this.setConfig = this.setConfig.bind(this);
     this.loadData = this.loadData.bind(this);
+    this.loadAggData = this.loadAggData.bind(this);
     this.changeView = this.changeView.bind(this);
 
     this.state = {
-      viewset: ["loadConfig", "parseConfig", "dataTable"],
+      viewset: ["loadConfig", "parseConfig", "dataTable", "aggTable"],
       view: "loadConfig",
       configData: null,
       dataLoader: null,
       dataRoot: null,
       outDir: null,
-      data: null
+      data: null,
+      aggData: null
     };
   }
 
@@ -345,8 +347,7 @@ class App extends React.Component {
     const methodCSVName = "selections-" + methodName + ".csv";
     const methodCSVPath = path.join(this.state["outDir"], methodDirName, methodCSVName);
 
-    const data = [];
-    const dataObj = {};
+    let data = [];
 
     fs.createReadStream(methodCSVPath)
       .pipe(parse({delimiter: ', '}))
@@ -369,6 +370,55 @@ class App extends React.Component {
 
   }
 
+  loadAggData() {
+    // Get list of available methods
+    const methods = Object.keys(this.state["configData"]["outlier_detection"]);
+    let data = [];
+
+    // For each method available
+    for (let methodName of methods) {
+      const currData = [];
+
+      // Define and find CSV
+      const methodOptions = [methodName];
+      if (this.state["configData"]["outlier_detection"][methodName] != null) {
+        for (const [key, value] of Object.entries(this.state["configData"]["outlier_detection"][methodName])) {
+          methodOptions.push(key + '=' + value);
+        }
+      }
+      const methodDirName = methodOptions.join('-');
+      const methodCSVName = "selections-" + methodName + ".csv";
+      const methodCSVPath = path.join(this.state["outDir"], methodDirName, methodCSVName);
+      
+      // Read imagepaths from CSV
+      fs.createReadStream(methodCSVPath)
+        .pipe(parse({delimiter:', '}))
+        .on('data', csvrow => {
+          currData.push(csvrow)
+        })
+        .on('end', () => {
[…]

### 2021-09-15 dora_vis/src/Table.js

@@ -127,6 +127,60 @@ function Table({ columns, data }) {
   )
 }
 
+class AggTable extends React.Component {
+  constructor(props) {
+    super(props);
+  }
+
+  componentDidMount() {
+    this.props.loadAggData();
+  }
+
+  render() {
+    const columns = [{
+      Header: "Rank",
+      accessor: "rank",
+      className: "rankCell",
+      headerClassName: "rankHeader",
+      sortType: "number"
+    }]
+
+    for (let methodName of Object.keys(this.props.configData["outlier_detection"])) {
+      columns.push({
+        Header: methodName,
+        className: "imageCell",
+        headerClassName: "imageHeader",
+        Cell: (row) => {
+          return(
+          <div>
+            <a className="imageLink" onClick={() => {navigator.clipboard.writeText(row.row.original[methodName+"Name"]);alert(row.row.original[methodName+"Name"]+ " copied to clipboard.");}}>
+              <img src={"data:image/png;base64,"+row.row.original[methodName]} title={row.row.original[methodName+"Name"]}/>
+            </a>
+          </div>
+          );
+        }
+      });
+    }
+
+    let datapass = null;
+    if (this.props.data == null) {
+      datapass = <h1> No DORA configuration loaded </h1>;
+    } else {
+      datapass = <>
+        <h1>Aggregate Results</h1>
+        <Table columns={columns} data={this.props.data}/>
+      </>;
+    }
+
+    return (
+    <div className="container">
+      {datapass}
+      <br/>
+    </div>
+    );
+  }
+}
+
 class DataTable extends React.Component {
   constructor(props) {
     super(props);
@@ -207,7 +261,7 @@ class DataTable extends React.Component {
       datapass = 
       <>
       <div className="col-md-4 methodSelect">
-        <label for="methodSel">Results from method:</label>
+        <label>Results from method:</label>
         <select className="form-select form-select-lg" id="methodSel" onChange={this.switchMethod}>
           {this.state["methods"].map((method) => <option value={method}>{method}</option>)}
         </select>
@@ -226,4 +280,4 @@ class DataTable extends React.Component {
   }
 }
 
-export default DataTable;
+export { DataTable, AggTable };

### 2021-09-15 dora_vis/src/App.js
Headers for configuration step

@@ -270,8 +270,13 @@ class ConfigParser extends React.Component {
       <div className="container-fluid">
         <h1>Parse DORA Config</h1>
         <form>
+          <h2> Dataloader Check </h2>
           {componentDataLoader}
+          <br/>
+          <h2> Data Root Check </h2>
           {componentDataRoot}
+          <br/>
+          <h2> Results Dir Check </h2>
           {componentOutDir}
           <br/>
           {componentConfigButton}

### 2021-09-15 dora_vis/src/Table.js
Clipboard copy fix

@@ -5,6 +5,7 @@ import { useTable, useSortBy, usePagination} from 'react-table'
 
 const fs = window.require('fs');
 const path = window.require('path');
+const {clipboard} = window.require('electron');
 
 
 function Table({ columns, data }) {
@@ -153,7 +154,7 @@ class AggTable extends React.Component {
         Cell: (row) => {
           return(
           <div>
-            <a className="imageLink" onClick={() => {navigator.clipboard.writeText(row.row.original[methodName+"Name"]);alert(row.row.original[methodName+"Name"]+ " copied to clipboard.");}}>
+            <a className="imageLink" onClick={() => {clipboard.writeText(row.row.original[methodName+"Name"]);alert(row.row.original[methodName+"Name"]+ " copied to clipboard.");}}>
               <img src={"data:image/png;base64,"+row.row.original[methodName]} title={row.row.original[methodName+"Name"]}/>
             </a>
           </div>
@@ -228,7 +229,7 @@ class DataTable extends React.Component {
         Cell: (row) => {
           return (
           <div>
-            <a className="imageLink" onClick={() => {navigator.clipboard.writeText(row.row.original.fileName);alert(row.row.original.fileName+ " copied to clipboard.");}}>
+            <a className="imageLink" onClick={() => {clipboard.writeText(row.row.original.fileName);alert(row.row.original.fileName+ " copied to clipboard.");}}>
               <img src={"data:image/png;base64,"+row.row.original.imageData} title={row.row.original.fileName}/>
             </a>
           </div>

### 2021-09-16 publications/iaai21/dora_outlier_detection_iaai.tex
Add description of PAE. #42

--- 2021-09-15 publications/iaai21/dora_outlier_detection_iaai.tex
@@
 
-\subparagraph{Likelihood.}
+\subparagraph{Probability.}
 The negative sampling algorithm is implemented by converting the unsupervised 
@@
 as outlier scores, which means that the observations with higher posterior 
-probabilities are more likely to be outliers. \todo{Describe PAE: Hannah or Bryce}
+probabilities are more likely to be outliers. 
+The probabilistic autoencoder is a generative model consisting of an 
+autoencoder trained to reconstruct input data which is interpreted
+probabilistically after training using a normalizing flow on the autoencoder
+latent space~\cite{bohm2020probabilistic}. 
+Samples are scored as outliers using the log likelihood in the latent
+distribution, the autoencoder reconstruction error, or a combination of
+the two.
 
@@
   visual inspection of the satellite images.
- \todo{If we have time, I want to also evaluate the downstream classifier
- performance before and after removing top N outliers with DORA.}
+% descoped for lack of time and space
+% \todo{If we have time, I want to also evaluate the downstream classifier
+% performance before and after removing top N outliers with DORA.}
 
@@
 We added this to demonstrate the capabilities of DORA with a traditional
-benchmark dataset. \todo{Bryce. TBD if we will end up including due to
-space and/or relevance.}
+benchmark dataset. \todo{Bryce.}
 
@@
   \centering
-  \begin{tabular}{l|ccc}
+  \begin{tabular}{l|cccc}
     \hline
-    Algorithm & Astrophysics & Earth & Planetary \\
+    Algorithm & Astro. & Earth & Planetary & FMNIST \\
     \hline
@@
     \hline
-    Isolation Forest  & & 0.46 & \textbf{0.56} \\
+    Iso. Forest  & & 0.46 & \textbf{0.56} \\
     \hline

### 2021-09-16 dora_exp_pipeline/combined_plot_script.py
Change 'true' to 'known' in plot axis labels. #53

@@ -87,9 +87,9 @@ def combine_plots(resultsdir, label_path, precision_at_n):
         bestLine = list(range(1, num_outliers+1))
         plt.plot(bestLine, bestLine, label='Oracle', linestyle='--',
                  color='k')
-        plt.title('True Outliers vs. Algorithm Selections')
+        plt.title('Known Outliers vs. Algorithm Selections')
         plt.xlabel('Number of selections')
-        plt.ylabel('Number of true outliers')
+        plt.ylabel('Number of known outliers')
 
     for i in range(len(filenames)):
         scores = alg_indexes(filenames[i])

### 2021-09-16 publications/iaai21/dora_outlier_detection_iaai.tex
Define MDR and explain why no LRX. #42

@@
 algorithm by plotting the number of true outliers detected out of the top
-$N$ selections. We reported the area under the curve (AUC), normalized
-by the total number of selections, in the legend
-for each algorithm to give a quantitative comparison of the curves.
+$N$ selections. We reported the Mean Discovery Rate (MDR) in the legend
+for each algorithm to give a quantitative comparison of the curves. We
+defined MDR as:
+\begin{equation}
+MDR = \frac{\sum^{N_s}_{i=1} n_i}{\sum^{N_s}_{i=1} s_i}
+\end{equation}
+where $i \in [1, N_s]$ is the selection index,
+ $N_s$ is the total number of selections, 
+ $s_i$ is the number of selections made up to index $i$, 
+ and $n_i$ is the number of known outliers (true positives)
+among $s_i$ selections.
 Table \ref{tab:faw_results} gives the precision at $N$ where
@@
 $N=n_{outlier}$ while DEMUD and PCA tied for the lowest
-AUC. \todo{For completeness, Explain why no LRX here.}
+MDR. We did not evaluate LRX for this time series dataset because LRX can only 
+be applied to gridded image or raster data types.
 

### 2021-09-16 publications/iaai21/dora_outlier_detection_iaai.tex

@@
 distribution, the autoencoder reconstruction error, or a combination of
-the two.
+both.
 

### 2021-09-16 publications/iaai21/dora_outlier_detection_iaai.tex
Add to todo for reorganizing data/results description. #42
Reorganize results section and add table for high-level dataset description. #42

@@
  done by plotting the number of true outliers vs. number of selections made. 
- This can also be represented as the precision at $N$, i.e., 
- the number of true outliers divided by the number of selections $N$~\citep{campos2016evaluation}. Finally,
+  Finally,
  we provide a module for plotting histograms of outlier scores to visualize the
@@
 
+\begin{table}
+  \caption{Number of samples in the training and test sets for each dataset.}
+  \label{tab:datasets}
+  \centering
+  \begin{tabular}{l|c|cc}
+    \hline
+    Dataset & Training & \multicolumn{2}{c}{Test}\\
+     & Unlabeled &  Outliers &  Inliers \\
+
+    \hline
+    Astrophysics &   &   &   \\
+    Earth & 6,757  &  37 & 76  \\
+    Planetary &   &   &   \\
+    (F)MNIST &   &   &   \\
+    \hline
+  \end{tabular}
+\end{table}
+
 \subsection{Results}
-In this section we discuss the experimental setup and results for each dataset.
-We designed the experiments to be relevant to the real-world scientific use 
-cases associated with each dataset domain.
-
-\subsubsection{Astrophysical objects in Dark Energy Survey.}
-\todo{Umaa}
-
-\subsubsection{Targets in Mars rover images.}
-% We initialized each algorithm with $D_s$ and requested a ranking of
-% targets in $D_{s+}$.
-\todo{Hannah: Consider merging all three data set discussions since they are
-in one table? with their own figures? The point here isn't really to
-compare algs or data sets but instead to show that we CAN compare
-them, using DORA.}
-%
-Table~\ref{tab:faw_results} compares the performance of each ranking 
-algorithm by reporting the precision obtained when the number of
-selections is the same as the total number of outliers.  We could not
-employ RX in this case since the data dimensionality ($64 \times 64 =
-4096$) exceeded the data set size.  We found that
-the Isolation Forest achieved the highest precision (best outlier
-detection) when allowed to select only \num{9} images.
-Figure~\ref{fig:planetary_results} shows the complete (cumulative)
-outlier detection performance for each algorithm when ranking
-all \num{58} target images in $D_{1694-1703}$. \todo{Hannah: Explain
-MDR the first time it appears. }
-
-\subsubsection{Satellite time series for ground-truth observations.}
-We used the full dataset of 6,757 samples to fit the ranking algorithms and
-used the independently-labeled set of 113 points in Kenya for evaluation.
-Table \ref{fig:faw_results} compares the performance of each ranking 
-algorithm by plotting the number of true outliers detected out of the top
-$N$ selections. We reported the Mean Discovery Rate (MDR) in the legend
-for each algorithm to give a quantitative comparison of the curves. We
-defined MDR as:
+The experimental setup for each dataset was to fit or train a model for
+each ranking algorithm using a larger unlabeled dataset and then apply 
+the models to compute the outlier scores for a smaller test dataset for which
+labels of known outliers were available (Table \ref{tab:datasets}). 
+For each test set, we created a plot of the number of known/true outliers 
+detected out of the top $N$ selections. We also 
+reported the Mean Discovery Rate (MDR) in the legend
+for each algorithm to give a quantitative comparison of the curves that could
+be comparable across the datasets. We defined MDR as:
 \begin{equation}
@@
 among $s_i$ selections.
-Table \ref{tab:faw_results} gives the precision at $N$ where
-$N=n_{outlier}=37$. The negative sampling algorithm had the best performance
+We also reported the precision at $N=n_{outlier}$ for each test set where
+$n_{outlier}$ is the number of known outliers, i.e.,  the precision obtained 
+when the number of selections is the same as the total number of outliers. 
+Precision at $N$ is the number of true outliers divided by the number of 
+selections $N$~\citep{campos2016evaluation}.
+Table~\ref{tab:faw_results} compares the precision at $N=n_{outlier}$ for
+each dataset and ranking algorithm. 
+%\todo{Hannah: Consider merging all three data set discussions since they are
+%in one table? with their own figures? The point here isn't really to
+%compare algs or data sets but instead to show that we CAN compare
+%them, using DORA.}
+% We initialized each algorithm with $D_s$ and requested a ranking of
+% targets in $D_{s+}$.
+%
+
+For the planetary dataset, we found that
+the Isolation Forest achieved the highest precision (best outlier
+detection) when allowed to select only \num{9} images.
+Figure~\ref{fig:planetary_results} shows the complete (cumulative)
+outlier detection performance for each algorithm when ranking
+all \num{58} target images in $D_{1694-1703}$. We could not
+employ RX since the data dimensionality ($64 \times 64 =
+4096$) exceeded the data set size.
+
+For the Earth dataset, the negative sampling algorithm had the best performance
 in both metrics. DEMUD, PCA, and PAE tied for the lowest precision at 
 $N=n_{outlier}$ while DEMUD and PCA tied for the lowest
-MDR. We did not evaluate LRX for this time series dataset because LRX can only
-be applied to gridded image or raster data types.
+MDR (Figure \ref{fig:faw_results}). We did not evaluate LRX for this time 
+series dataset because LRX can only be applied to gridded image or raster 
+data types.
+
+\todo{Umaa: add results for astro use case.}
+
+\todo{Bryce: add results for (F)MNIST use case.}
 
@@
 \end{figure}
-
-%\subsubsection{Volcanic thermal anomalies}
-
-\subsubsection{MNIST and FashionMNIST.}
-\todo{Bryce. TBD if we will end up including due to
-space and/or relevance.}
 

### 2021-09-16 dora_exp_pipeline/dora_feature.py
PAE fixes for planetary rover data (#65)
- add pixel normalization to image laoder
- add pae config option disable flow
- re-run planetary rover experiment

Co-authored-by: Kiri Wagstaff <kiri.l.wagstaff@jpl.nasa.gov>

@@ -132,6 +132,9 @@ class FlattenedPixelValuesExtractor(FeatureExtractor):
             else:
                 ret_features[ind, :] = data.flatten()
 
+        if kwargs.get('normalize_pixels'):
+            ret_features = ret_features / 255.
+
         return ret_features
 
 

### 2021-09-16 dora_exp_pipeline/dora_results_organization.py
Merge with master

@@ -114,9 +114,9 @@ class SaveComparisonPlot(ResultsOrganization):
 
         plt.plot(x, y, label="{} (MDR: {:.2f})".format(outlier_alg_name, area))
         plt.plot(x, x, label='Oracle', linestyle='--', color='k')
-        plt.title('Correct Outliers vs Selected Outliers')
+        plt.title('Known Outliers vs. Selected Outliers')
         plt.xlabel('Number of Outliers Selected')
-        plt.ylabel('Number of True Outliers')
+        plt.ylabel('Number of Known Outliers')
         plt.legend()
         axes.set_xlim(1, x[-1])
         axes.set_ylim(1, y[-1])

### 2021-09-17 publications/iaai21/dora_outlier_detection_iaai.tex
Add results and figure for fmnist experiment
Add text for fmnist dataset and results sections

@@
  (astrophysics, planetary science, and Earth science). We also included a 
- benchmark dataset that uses MNIST and FashionMNIST. Here we describe
+ benchmark dataset that uses MNIST and Fashion-MNIST. Here we describe
  the details of each dataset.
@@
 
-\subsubsection{MNIST and FashionMNIST.}
-We added this to demonstrate the capabilities of DORA with a traditional
-benchmark dataset. \todo{Bryce.}
+\subsubsection{MNIST and Fashion-MNIST.}
+We included MNIST and F-MNIST to demonstrate the capabilities of DORA with a
+traditional benchmark dataset. In this experiment, we used 60,000 images from 
+F-MNIST as the training set and a combined set of 2,000 images from MNIST and 
+F-MNIST as the evaluation set. Well-discriminating algorithms should assign 
+higher outlier scores to samples from MNIST than F-MNIST.
 
@@
     Planetary &   &   &   \\
-    (F)MNIST &   &   &   \\
+    F-MNIST & 60,000 & 1,000 & 1,000  \\
     \hline
@@
 
-\todo{Bryce: add results for (F)MNIST use case.}
+For the MNIST and F-MNIST dataset, PCA and DEMUD tied for the highest precision,
+while DEMUD, PCA, and PAE tied for the highest MDR. Negative sampling had the 
+lowest performance in both metrics.
 
@@
     \hline
-    Algorithm & Astro. & Earth & Planetary & FMNIST \\
+    Algorithm & Astro. & Earth & Planetary & F-MNIST \\
     \hline
-    PCA               & & 0.41 & 0.44 \\
-    DEMUD             & & 0.41 & 0.44 \\
+    PCA               & & 0.41 & 0.44 & \textbf{0.84} \\
+    DEMUD             & & 0.41 & 0.44 & \textbf{0.84} \\
     \hline
-    RX                & & 0.43 & ---- \\
-    LRX               & & ---- & 0.33 \\
+    RX                & & 0.43 & ---- & 0.82 \\
+    LRX               & & ---- & 0.33 & 0.56 \\
     \hline
-    Iso. Forest  & & 0.46 & \textbf{0.56} \\
+    Iso. Forest  & & 0.46 & \textbf{0.56} & 0.74 \\
     \hline
-    PAE               & & 0.41 & 0.44 \\
-    Neg. Sampling & & \textbf{0.49} & 0.33 \\
+    PAE               & & 0.41 & 0.44 & 0.83 \\
+    Neg. Sampling & & \textbf{0.49} & 0.33 & 0.43 \\
     \hline
-    Random            & & 0.32 & 0.11 \\
+    Random            & & 0.32 & 0.11 & 0.49 \\
     \hline
@@
     \label{fig:faw_results}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{figures/fmnist_combined_plot.pdf}
+    \caption{Number of true outliers ranked in top $N$ selections for 
+    F-MNIST dataset.}
+    \label{fig:fmnist_results}
 \end{figure}

### 2021-09-17 dora_exp_pipeline/combined_plot_script.py
Added theoretical random to comparison plot. Closes #63.

@@ -1,6 +1,7 @@
 import os
 import matplotlib.pyplot as plt
 import argparse
+import math, statistics
 
 
 def alg_indexes(filename):
@@ -70,6 +71,30 @@ def get_precision_curve(scores, validationLabels):
     return x, y
 
 
+def random_sel_EV(scores, labels, n):
+    n_scores = len(scores)
+    n_outliers = sum(labels.values())
+    return sum([math.comb(n_outliers, i) * math.comb(n_scores-n_outliers, n-i) * i for i in range(n+1)]) / math.comb(n_scores, n)
+
+
+def get_random_selections_curve(scores, validationLabels):
+    x = list(range(1, len(scores)+1))
+    y = []
+    for i in range(len(scores)):
+        y.append(random_sel_EV(scores, validationLabels, i))
+    
+    return x, y
+
+
+def get_random_precision_curve(scores, validationLabels):
+    x = list(range(1, len(scores)+1))
+    y = []
+    for i in range(len(scores)):
+        y.append(random_sel_EV(scores, validationLabels, i) / (i+1))
+
+    return x, y
+
+
 def combine_plots(resultsdir, label_path, precision_at_n):
     filenames, labels = filenames_and_labels(resultsdir)
 
@@ -102,6 +127,14 @@ def combine_plots(resultsdir, label_path, precision_at_n):
             x, y = get_selections_curve(scores, validationLabels)
             area = sum(y)/sum([i for i in range(len(scores))])
             plt.plot(x, y, label="{} (MDR: {:.2f})".format(labels[i], area))
+    
+    if precision_at_n:
+        x, y = get_random_precision_curve(scores, validationLabels)
+        plt.plot(x, y, label="{}".format("Theoretical Random"), linestyle='--')
+    else:
+        x, y = get_random_selections_curve(scores, validationLabels)
+        area = sum(y)/sum([i for i in range(len(scores))])
+        plt.plot(x, y, label="{} (MDR: {:.2f})".format("Theoretical Random", area), linestyle='--')
 
     axes.set_xlim(1, len(scores))
     plt.legend()

### 2021-09-17 dora_exp_pipeline/combined_plot_script.py
Satisfying the lint gods

@@ -1,7 +1,7 @@
 import os
 import matplotlib.pyplot as plt
 import argparse
-import math, statistics
+import math
 
 
 def alg_indexes(filename):
@@ -74,7 +74,9 @@ def get_precision_curve(scores, validationLabels):
 def random_sel_EV(scores, labels, n):
     n_scores = len(scores)
     n_outliers = sum(labels.values())
-    return sum([math.comb(n_outliers, i) * math.comb(n_scores-n_outliers, n-i) * i for i in range(n+1)]) / math.comb(n_scores, n)
+    return sum([math.comb(n_outliers, i) * \
+                math.comb(n_scores-n_outliers, n-i) * i \
+                for i in range(n+1)]) / math.comb(n_scores, n)
 
 
 def get_random_selections_curve(scores, validationLabels):
@@ -82,7 +84,7 @@ def get_random_selections_curve(scores, validationLabels):
     y = []
     for i in range(len(scores)):
         y.append(random_sel_EV(scores, validationLabels, i))
-    
+
     return x, y
 
 
@@ -127,14 +129,16 @@ def combine_plots(resultsdir, label_path, precision_at_n):
             x, y = get_selections_curve(scores, validationLabels)
             area = sum(y)/sum([i for i in range(len(scores))])
             plt.plot(x, y, label="{} (MDR: {:.2f})".format(labels[i], area))
-    
+
     if precision_at_n:
         x, y = get_random_precision_curve(scores, validationLabels)
         plt.plot(x, y, label="{}".format("Theoretical Random"), linestyle='--')
     else:
         x, y = get_random_selections_curve(scores, validationLabels)
         area = sum(y)/sum([i for i in range(len(scores))])
-        plt.plot(x, y, label="{} (MDR: {:.2f})".format("Theoretical Random", area), linestyle='--')
+        plt.plot(x, y, 
+                 label="{} (MDR: {:.2f})".format("Theoretical Random", area), 
+                 linestyle='--')
 
     axes.set_xlim(1, len(scores))
     plt.legend()

### 2021-09-17 dora_exp_pipeline/combined_plot_script.py
More linting

@@ -74,8 +74,8 @@ def get_precision_curve(scores, validationLabels):
 def random_sel_EV(scores, labels, n):
     n_scores = len(scores)
     n_outliers = sum(labels.values())
-    return sum([math.comb(n_outliers, i) * \
-                math.comb(n_scores-n_outliers, n-i) * i \
+    return sum([math.comb(n_outliers, i) *
+                math.comb(n_scores-n_outliers, n-i) * i
                 for i in range(n+1)]) / math.comb(n_scores, n)
 
 
@@ -136,8 +136,8 @@ def combine_plots(resultsdir, label_path, precision_at_n):
     else:
         x, y = get_random_selections_curve(scores, validationLabels)
         area = sum(y)/sum([i for i in range(len(scores))])
-        plt.plot(x, y, 
-                 label="{} (MDR: {:.2f})".format("Theoretical Random", area), 
+        plt.plot(x, y,
+                 label="{} (MDR: {:.2f})".format("Theoretical Random", area),
                  linestyle='--')
 
     axes.set_xlim(1, len(scores))

### 2021-09-17 dora_vis/src/Table.js
Changed table page qty options

@@ -117,7 +117,7 @@ function Table({ columns, data }) {
             setPageSize(Number(e.target.value))
           }}
         >
-          {[10, 20, 30, 40, 50].map(pageSize => (
+          {[5, 10, 20, 50, 100].map(pageSize => (
             <option key={pageSize} value={pageSize}>
               Show {pageSize}
             </option>

### 2021-09-17 publications/iaai21/dora_outlier_detection_iaai.tex
Minor updates to text, formatting, todos. #42

@@
 
-\section{Experiments}
+\section{Datasets}
 
-\subsection{Datasets}
 We constructed three datasets to evaluate the utility of DORA and the
@@
 We included MNIST and F-MNIST to demonstrate the capabilities of DORA with a
-traditional benchmark dataset. In this experiment, we used 60,000 images from 
+traditional benchmark dataset. We used 60,000 images from 
 F-MNIST as the training set and a combined set of 2,000 images from MNIST and 
@@
     Earth & 6,757 & 37 & 76  \\
-    Planetary &   &   &   \\
+    Planetary & 992 &  9 & 49 \\
     F-MNIST & 60,000 & 1,000 & 1,000  \\
@@
 
-\subsection{Results}
+\section{Results}
 The experimental setup for each dataset was to fit or train a model for
@@
 each dataset and ranking algorithm. 
+\todo{Jake: add description of theoretical random baseline.}
 %\todo{Hannah: Consider merging all three data set discussions since they are
@@
 
-For the MNIST and F-MNIST dataset, PCA and DEMUD tied for the highest precision,
+For the MNIST and F-MNIST dataset, PCA and DEMUD tied for the highest
+ precision at $N=n_{outlier}$
 while DEMUD, PCA, and PAE tied for the highest MDR. Negative sampling had the 
@@
 \begin{table}
-  \caption{Precision at $N=n_{outlier}$ for three applications; the
-  best result for each data set is in bold.}
+  \caption{Precision at $N=n_{outlier}$ for three applications and F-MNIST
+   benchmark; the best result for each data set is in bold.}
   \label{tab:faw_results}
@@
 \section{Discussion}
-\todo{ 
-Potential topics TBD based on results:
-\begin{itemize}
-\item Discuss results from experiments
-\item Challenges of evaluating results in outlier detection?
-\item Downstream use of results in science applications
-\end{itemize}
-}
+
+\paragraph{Algorithm performance.} \todo{Discuss overall results from
+ experiments and that we left hyperparameter tuning and extensive
+ evaluation for future work.}
+
+\paragraph{Evaluation in outlier detection.} \todo{Discuss why evaluation
+in outlier detection applications is hard since we often don't have complete
+labeled data for testing and labels can be subjective.}
+
 \paragraph{Path to deployment.} \todo{Hannah - describe pip-installable

### 2021-09-17 dora_exp_pipeline/combined_plot_script.py
Specify colors to use for each algorithm so they are consistent across datasets. #53

@@ -2,6 +2,16 @@ import os
 import matplotlib.pyplot as plt
 import argparse
 
+alg_colors = {
+                'demud': 'red',
+                'pca': 'purple',
+                'iforest': 'pink',
+                'pae': 'blue',
+                'negative sampling': 'brown',
+                'random': 'gray',
+                'rx': 'green',
+                'lrx': 'orange'
+             }
 
 def alg_indexes(filename):
     with open(filename, 'r') as f:
@@ -97,11 +107,13 @@ def combine_plots(resultsdir, label_path, precision_at_n):
         if precision_at_n:
             print(labels[i])
             x, y = get_precision_curve(scores, validationLabels)
-            plt.plot(x, y, label="{}".format(labels[i]))
+            plt.plot(x, y, label="{}".format(labels[i]),
+                     color=alg_colors[labels[i]])
         else:
             x, y = get_selections_curve(scores, validationLabels)
             area = sum(y)/sum([i for i in range(len(scores))])
-            plt.plot(x, y, label="{} (MDR: {:.2f})".format(labels[i], area))
+            plt.plot(x, y, label="{} (MDR: {:.2f})".format(labels[i], area),
+                     color=alg_colors[labels[i]])
 
     axes.set_xlim(1, len(scores))
     plt.legend()

### 2021-09-17 publications/iaai21/dora_outlier_detection_iaai.tex
Resolve dataset summary table todo. #42

@@
  (astrophysics, planetary science, and Earth science). We also included a 
- benchmark dataset that uses MNIST and Fashion-MNIST. Here we describe
- the details of each dataset.
-\todo{Maybe add
- a table with number of samples, number of reference known outliers (if 
- available), etc, if room? }
- \todo{Can we make any of these datasets publicly available with the paper?}
+ benchmark dataset that uses MNIST and Fashion-MNIST. Table \ref{tab:datasets}
+ summarizes the number of unlabeled samples used for training and labeled
+ samples used for evaluation for each dataset. We describe each in detail
+ below.
 
@@
  benchmark tasks?]
+ \todo{Can we make any of these datasets publicly available with the paper?}
 }

### 2021-09-17 dora_exp_pipeline/combined_plot_script.py
Use fixed colors for plotting algorithms. #63

@@ -3,6 +3,16 @@ import matplotlib.pyplot as plt
 import argparse
 import math
 
+alg_colors = {
+                'demud': 'red',
+                'pca': 'purple',
+                'iforest': 'pink',
+                'pae': 'blue',
+                'negative sampling': 'brown',
+                'random': 'gray',
+                'rx': 'green',
+                'lrx': 'orange'
+             }
 
 def alg_indexes(filename):
     with open(filename, 'r') as f:
@@ -124,11 +134,13 @@ def combine_plots(resultsdir, label_path, precision_at_n):
         if precision_at_n:
             print(labels[i])
             x, y = get_precision_curve(scores, validationLabels)
-            plt.plot(x, y, label="{}".format(labels[i]))
+            plt.plot(x, y, label="{}".format(labels[i]),
+                     color=alg_colors[labels[i]])
         else:
             x, y = get_selections_curve(scores, validationLabels)
             area = sum(y)/sum([i for i in range(len(scores))])
-            plt.plot(x, y, label="{} (MDR: {:.2f})".format(labels[i], area))
+            plt.plot(x, y, label="{} (MDR: {:.2f})".format(labels[i], area),
+                     color=alg_colors[labels[i]])
 
     if precision_at_n:
         x, y = get_random_precision_curve(scores, validationLabels)

### 2021-09-17 dora_exp_pipeline/combined_plot_script.py
Replace math.comb with custom comb for py3.6

@@ -1,7 +1,6 @@
 import os
 import matplotlib.pyplot as plt
 import argparse
-import math
 
 alg_colors = {
                 'demud': 'red',
@@ -13,6 +12,19 @@ alg_colors = {
                 'rx': 'green',
                 'lrx': 'orange'
              }
+
+
+def comb(n: int, k: int) -> int:
+    if k < 0 or k > n:
+        return 0
+    if k == 0 or k == n:
+        return 1
+    k = min(k, n - k)  # Take advantage of symmetry
+    c = 1
+    for i in range(k):
+        c = c * (n - i) / (i + 1)
+    return c
+
 
 def alg_indexes(filename):
     with open(filename, 'r') as f:
@@ -84,9 +96,9 @@ def get_precision_curve(scores, validationLabels):
 def random_sel_EV(scores, labels, n):
     n_scores = len(scores)
     n_outliers = sum(labels.values())
-    return sum([math.comb(n_outliers, i) *
-                math.comb(n_scores-n_outliers, n-i) * i
-                for i in range(n+1)]) / math.comb(n_scores, n)
+    return sum([comb(n_outliers, i) *
+                comb(n_scores-n_outliers, n-i) * i
+                for i in range(n+1)]) / comb(n_scores, n)
 
 
 def get_random_selections_curve(scores, validationLabels):

### 2021-09-18 publications/iaai21/dora_outlier_detection_iaai.tex
Fill in related work section. #42

@@
 \section{Related Work}
+Methods for outlier detection have been 
+surveyed extensively in prior work
+~\cite{pimentel2014review,chandola2009anomaly,markou2003novelty,
+markou2003novelty2}.
+Methods can be differentiated primarily based on how they define outlierness.
+%
+Reconstruction-based methods aim to construct a model of a dataset by learning 
+a mapping between the data samples and a lower-dimensional representation that
+ minimizes the loss between the input and its reconstruction from the 
+ low-dimensional representation~\citep{kerner2020comparison}. The 
+ reconstruction error is used to compute the outlier score because samples
+  that are unlike the data used to fit the model will be more poorly 
+  reconstructed compared to inliers. Reconstruction-based methods
+  include PCA~\cite{jablonski2015principal}, 
+  autoencoders~\cite{richter2017safe,park2018multimodal}, and
+  generative adversarial networks~\cite{akcay2018ganomaly}.
+%
+ Distance-based methods score outliers based on their distance from some
+``background'' which can be defined in a variety of ways. For example,
+the Reed-Xiaoli (RX) detector computes the 
+Mahalanobis distance between each sample and the background dataset
+defined by its mean and covariance matrix~\cite{reed1990adaptive}.
+%
+Sparsity-based methods, such as the isolation forest~\cite{liu2008isolation}
+and local outlier factor\cite{breunig2000lof},
+ score outliers based on how isolated or sparse they
+are in a given feature space. 
+%
+Probability distribution and density based methods estimate the underlying 
+distribution or probability density of a dataset and score samples based on
+their likelihood under the learned distribution. Examples include the probabilistic
+autoencoder, which scores samples based on the log likelihood under the learned
+latent space distribution~\cite{bohm2020probabilistic}, Gaussian mixture 
+Models, and kernel density estimators~\cite{chandola2009anomaly}. 
+%
+Other methods treat detection of outliers as a supervised classification 
+problem, usually with only one class constituted by the inliers or 
+known normal samples.
+Such methods include one-class support vector 
+machines~\cite{scholkopf1999support},
+support vector data description~\cite{pimentel2014review},
+and negative sampling~\cite{sipple:neg-sampling20}.
+
 In astrophysics, 
 outlier detection methods have been employed for identifying astrophysical
-     objects with unique characteristics (Xiong et al., 2010; 
-     Storey-Fisher et al., 2020; Hayat et al., 2020) as well as 
-     data/modeling artifacts~\citep{wagstaff:des-anom20} (Lochner et al., 2020) in 
-     astronomical surveys. Examples of outlier detection applications in 
+objects with unique characteristics~\cite{xiong2010anomaly,
+storey2020anomaly,hayat2021self} as well as data or modeling 
+artifacts in astronomical surveys~\citep{wagstaff:des-anom20,
+lochner2021astronomaly}. 
+Examples of outlier detection applications in 
 Earth science include detecting anomalous objects or materials
-     (Chang and Chiang, 2002; Zhou et al., 2016), data artifacts/noise 
-     (Liu et al., 2017; Alvera-Azcárate et al., 2012), change (Touati et al.,
-      2020; Zhou et al., 2016),  and ocean extremes (Prochaska et al.). 
+~\cite{chang2002anomaly,zhou2016novel}, data artifacts or noise 
+~\cite{liu2017unsupervised,alvera2012outlier}, change~\cite{zhou2016novel,
+touati2020anomaly}, and ocean extremes~\cite{prochaska2021deep}. 
 Planetary science applications of outlier detection have mostly focused 
 on prioritizing samples with novel geologic or geochemical features for
-       follow-up targeting or analysis (Kerner et al., 2020a; Kerner et al., 
-       2020b) \citep{wagstaff:demud13}.
-
-\begin{itemize}
-\item Methods for outlier detection
-\item Applications of outlier detection related to our focus domains 
-(e.g., \citet{kerner2020comparison})
-\item Outlier detection toolkits: PyOD Python
-package~\citep{zhao:pyod19}, \url{https://github.com/yzhao062/pyod} 
-\item Tools for applying outlier detection or other methods across
-domains?
-\end{itemize}
+follow-up targeting or analysis~\cite{kerner2020analysis,kerner2020comparison,
+wagstaff:demud13}.
+%
+These examples show the promise of applying outlier detection methods in
+a variety of real-world science use cases. However, the effort required to 
+apply and evaluate the variety of available algorithms is non-trivial and can 
+be daunting for non-ML experts, thus impeding the uptake of outlier
+detection methods in science applications.  
+There is a need for tools that make it easier for domain scientists to apply
+outlier detection methods as well as compare results across datasets.
+While there have been some efforts to develop tools for facilitating the
+application of outlier detection methods~\cite{zhao2019pyod}, they cover
+limited data formats and algorithms. The aim of DORA is to fill the need 
+for tools that faciliate application, evaluation, and interpretation of outlier
+detection methods.
 
@@
 \subparagraph{Reconstruction error.}
-Reconstruction-based methods aim to construct a model of a dataset by learning 
-a mapping between the data samples and a lower-dimensional representation that
- minimizes the loss between the input and its reconstruction from the 
- low-dimensional representation~\citep{kerner2020comparison}. The 
- reconstruction error is used to compute the outlier score because samples
-  that are unlike the data used to fit the model will be more poorly 
-  reconstructed compared to inliers. 
 Principal component analysis (PCA) has been used for outlier detection by
@@
 \subparagraph{Distance.}
-Distance-based methods score outliers based on their distance from some
-``background'' which is usually defined by a sample distribution. The 
+ The 
 Reed-Xiaoli (RX) detector is commonly used for anomaly detection in
@@
 \subparagraph{Sparsity.}
-Sparsity-based methods score outliers based on how isolated or sparse they
-are in the feature space. Isolation forests are a common sparsity-based method
+Isolation forests are a common sparsity-based method
 that constructs many random binary trees from a 

### 2021-09-18 dora_exp_pipeline/combined_plot_script.py
Make theoretical random line the random color. #63

@@ -156,13 +156,14 @@ def combine_plots(resultsdir, label_path, precision_at_n):
 
     if precision_at_n:
         x, y = get_random_precision_curve(scores, validationLabels)
-        plt.plot(x, y, label="{}".format("Theoretical Random"), linestyle='--')
+        plt.plot(x, y, label="{}".format("Theoretical Random"), linestyle='--',
+                 color=alg_colors['random'])
     else:
         x, y = get_random_selections_curve(scores, validationLabels)
         area = sum(y)/sum([i for i in range(len(scores))])
         plt.plot(x, y,
                  label="{} (MDR: {:.2f})".format("Theoretical Random", area),
-                 linestyle='--')
+                 linestyle='--', color=alg_colors['random'])
 
     axes.set_xlim(1, len(scores))
     plt.legend()

### 2021-09-18 dora_exp_pipeline/combined_plot_script.py
Added printout of precision at N

@@ -114,7 +114,12 @@ def get_random_precision_curve(scores, validationLabels):
     x = list(range(1, len(scores)+1))
     y = []
     for i in range(len(scores)):
-        y.append(random_sel_EV(scores, validationLabels, i) / (i+1))
+        numOutliers = random_sel_EV(scores, validationLabels, i)
+        y.append(numOutliers / (i+1))
+
+        if (i+1) == sum(validationLabels.values()):
+            print('Precision at N=%d: %f' %
+                (sum(validationLabels.values()), float(numOutliers)/(i+1)))
 
     return x, y
 
@@ -155,6 +160,7 @@ def combine_plots(resultsdir, label_path, precision_at_n):
                      color=alg_colors[labels[i]])
 
     if precision_at_n:
+        print("random_theoretical")
         x, y = get_random_precision_curve(scores, validationLabels)
         plt.plot(x, y, label="{}".format("Theoretical Random"), linestyle='--')
     else:

### 2021-09-18 dora_exp_pipeline/combined_plot_script.py
pylint tabbing

@@ -119,7 +119,7 @@ def get_random_precision_curve(scores, validationLabels):
 
         if (i+1) == sum(validationLabels.values()):
             print('Precision at N=%d: %f' %
-                (sum(validationLabels.values()), float(numOutliers)/(i+1)))
+                  (sum(validationLabels.values()), float(numOutliers)/(i+1)))
 
     return x, y
 

### 2021-09-18 publications/iaai21/dora_outlier_detection_iaai.tex

@@
 \paragraph{Results visualization.}
-\todo{Jake: describe website if ready by deadline.}
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.75\linewidth]{figures/doravis.png}
+  \caption{A screenshot of the DORA visualizer displaying results from the
+  planetary science dataset.}
+  \label{fig:doravis}
+\end{figure}
+
+We developed a desktop application to easily visualize DORA results with the
+Electron application framework and React frontend library. The application loads
+the DORA configuration file to locate the dataset and result CSVs. Then,
+it displays the ranked samples and their scores in a table sorted by their
+selection order. This allows for fast and easy comparison of the results of
+different methods. Figure \ref{fig:doravis} shows a screenshot of the
+``Aggregate Table'' view, which displays all results from different algorithms
+side-by-side.
+
 \todo{Eric: describe causal graphs if added by deadline.}
@@
 Table~\ref{tab:faw_results} compares the precision at $N=n_{outlier}$ for
-each dataset and ranking algorithm. 
-\todo{Jake: add description of theoretical random baseline.}
+each dataset and ranking algorithm. \todo{Random desc by Jake, please edit}
+Metrics for random selection are calculated
+using the expected value of $n_i$ for $i$ random selections:
+\begin{equation}
+\mathbf{E}[n_i, i \in[1, N_s]] = \frac{\sum^i_{j=0} {n_{outlier}\choose j} {D-n_{outlier}\choose i-j} j}{{D\choose i}}
+\end{equation}
+
 %\todo{Hannah: Consider merging all three data set discussions since they are

### 2021-09-18 publications/iaai21/dora_outlier_detection_iaai.tex
Add description and figure for web interface. #42

--- 2021-09-18 publications/iaai21/dora_outlier_detection_iaai.tex
@@
 
-\paragraph{Results organization.} 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.9\linewidth]{figures/website.png}
+    \caption{Example view of algorithm ranking results in DORA web interface 
+    for planetary dataset.}
+    \label{fig:website}
+\end{figure}
+
+\paragraph{Results interpretation.} 
 Each of the outlier ranking algorithms returns an array containing the sample
 index, outlier score, and selection index (index of each sample after sorting
-by outlier score). DORA provides results organization modules intended to help
- users interpret and make decisions based on these outputs. The simplest 
+by outlier score). DORA provides organization and visualization modules 
+intended to help users interpret and make decisions based on these outputs. 
+The simplest 
  module saves a CSV of the samples sorted by their outlier score 
@@
  done by plotting the number of true outliers vs. number of selections made. 
-  Finally,
- we provide a module for plotting histograms of outlier scores to visualize the
+We provide a module for plotting histograms of outlier scores to visualize the
  distribution of scores in the dataset (which may be, e.g., multimodal or 
  long-tailed).
-
-\paragraph{Results visualization.} 
-\todo{Jake: describe website if ready by deadline.}
-\todo{Eric: describe causal graphs if added by deadline.}
+In addition, we are developing a web interface which will be packaged as an
+executable desktop app to further facilitate
+viewing and probing of the results. Figure \ref{fig:website} shows an example 
+view of the rankings across algorithms for the planetary dataset.
 
@@
 
-\subsubsection{Astrophysical objects in Dark Energy Survey.}
+\subsubsection{Astrophysics: objects in Dark Energy Survey.}
 \todo{Umaa}
 
-\subsubsection{Targets in Mars rover images.}
+\subsubsection{Planetary: targets in Mars rover images.}
 Mars exploration is fundamentally an exercise in discovery with the
@@
 
-\subsubsection{Satellite time series for ground-truth observations.}
+\subsubsection{Earth: satellite time series for ground-truth observations.}
 Many Earth science applications using satellite Earth observation (EO) data
@@
 
-\subsubsection{MNIST and Fashion-MNIST.}
+\subsubsection{Benchmark: MNIST and Fashion-MNIST.}
 We included MNIST and F-MNIST to demonstrate the capabilities of DORA with a
@@
  \todo{Can we make any of these datasets publicly available with the paper?}
+ \todo{Future work: other algorithms, datasets, causal graphs.}
 }

### 2021-09-18 publications/iaai21/dora_outlier_detection_iaai.tex

@@
  long-tailed).
-In addition, we are developing a web interface which will be packaged as an
-executable desktop app to further facilitate
-viewing and probing of the results. Figure \ref{fig:website} shows an example 
-view of the rankings across algorithms for the planetary dataset.
+
+\begin{figure}
+  \centering
+  \includegraphics[width=0.75\linewidth]{figures/doravis.png}
+  \caption{A screenshot of the DORA visualizer displaying results from the
+  planetary science dataset.}
+  \label{fig:doravis}
+\end{figure}
+
+We developed a desktop application to easily visualize DORA results with the
+Electron application framework and React frontend library. The application loads
+the DORA configuration file to locate the dataset and result CSVs. Then,
+it displays the ranked samples and their scores in a table sorted by their
+selection order. This allows for fast and easy comparison of the results of
+different methods. Figure \ref{fig:doravis} shows a screenshot of the
+``Aggregate Table'' view, which displays all results from different algorithms
+side-by-side.
 

### 2021-09-18 dora_exp_pipeline/combined_plot_script.py
Create update fmnist plot

@@ -13,6 +13,7 @@ alg_colors = {
                 'lrx': 'orange'
              }
 
+
 def alg_indexes(filename):
     with open(filename, 'r') as f:
         text = f.read().split("\n")[:-1]

### 2021-09-18 publications/iaai21/dora_outlier_detection_iaai.tex
Remove duplicate figure. #42

@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=0.9\linewidth]{figures/website.png}
-    \caption{Example view of algorithm ranking results in DORA web interface
-    for planetary dataset.}
-    \label{fig:website}
-\end{figure}
-
 \paragraph{Results interpretation.}
@@
 Table~\ref{tab:faw_results} compares the precision at $N=n_{outlier}$ for
-each dataset and ranking algorithm. \todo{Random desc by Jake, please edit}
-Metrics for random selection are calculated
-using the expected value of $n_i$ for $i$ random selections:
+each dataset and ranking algorithm. 
+We calculated a random selection baseline which we refer to as ``Theoretical
+Random'' using the expected value of $n_i$ for $i$ random selections:
 \begin{equation}

### 2021-09-18 dora_exp_pipeline/combined_plot_script.py
Simp. EV func, updated paper, added amsmath (#63)

@@ -14,18 +14,6 @@ alg_colors = {
              }
 
 
-def comb(n: int, k: int) -> int:
-    if k < 0 or k > n:
-        return 0
-    if k == 0 or k == n:
-        return 1
-    k = min(k, n - k)  # Take advantage of symmetry
-    c = 1
-    for i in range(k):
-        c = c * (n - i) / (i + 1)
-    return c
-
-
 def alg_indexes(filename):
     with open(filename, 'r') as f:
         text = f.read().split("\n")[:-1]
@@ -96,9 +84,14 @@ def get_precision_curve(scores, validationLabels):
 def random_sel_EV(scores, labels, n):
     n_scores = len(scores)
     n_outliers = sum(labels.values())
-    return sum([comb(n_outliers, i) *
-                comb(n_scores-n_outliers, n-i) * i
-                for i in range(n+1)]) / comb(n_scores, n)
+
+    # Full formulation
+    #return sum([comb(n_outliers, i) *
+    #            comb(n_scores-n_outliers, n-i) * i
+    #            for i in range(n+1)]) / comb(n_scores, n)
+
+    # Simplified formulation
+    return (n_outliers * n) / n_scores
 
 
 def get_random_selections_curve(scores, validationLabels):

### 2021-09-18 publications/iaai21/dora_outlier_detection_iaai.tex
Update all results plots with same colors and theoretical random. #42
adding description of DES and data used plus associated references

@@
 \usepackage{siunitx}
+\usepackage{amsmath}
 \usepackage[dvipsnames]{xcolor}
@@
 
+Astronomical data sets are large and growing. Large modern optical
+imaging surveys are producing catalogs for of order $10^8$ stars
+and galaxies, with dozens or hundreds of distinct measured features
+for each entry. Discovery science becomes difficult at this data
+volume: the scale is too large for expert human inspection, and separating real
+astrophysical anomalies from non-astrophysical sources like detector artifacts or
+satellite trails is a challenging problem for current
+methods.
+
+The Dark Energy Survey (DES), an ongoing imaging survey of
+$5000~{\rm deg}^2$ of the southern sky conducted from the Cerro-Tololo
+Inter-American Observatory in Chile \cite{DESY1}. The galaxy catalogs produced from
+DES imaging have provided some of the strongest constraints to date on
+the physical properties of dark energy and the accelerated expansion
+of the Universe. The first version of this catalog, released in June 2018, incorporated
+only cuts on signal-to-noise, resolution, masks against known detector
+anomalies and data quality indicators, and the automated data quality
+flags produced during processing. In December 2019, the full catalog
+was released after 18 months of extensive manual vetting. Therefore,
+we were able to use the second version of the catalog as a validation
+set for anomaly detection on the first version.
+
+We compared all methods on a data set that consists of 100K galaxy objects
+observed by the Dark Energy Survey (DES) sampled from the initial June 2018 release.
+25,339 objects from this 100K set that did not appear in the later Dec 2019
+release were likely eliminated during the manual vetting process and were labeled as outliers.
+Note, that the remaining 74,661 objects may also contain outliers, but we assume them to be
+inliers in this experiment.
+
+We used publicly-available photometry from the $g-$, $r-$, $i-$ and $z-band$ DES exposures.  We
+transformed the photometry into luptitudes\footnote{A
+  ``Luptitude''~\cite{lupton99} is an arcsinh-scaled flux, with
+  properties quantitatively equal to
+  traditional astronomical magnitudes for bright sources, but which
+  gracefully handles non-detections and negative fluxes.}.  The inputs features used
+  were the $r$-band luptitude, colors computed as band differences between
+$g-r$, $i-r$, and $z-r$, and associated observational errors, for a total of eight features.
+%
+
 \subsubsection{Planetary: targets in Mars rover images.}
@@
 Random'' using the expected value of $n_i$ for $i$ random selections:
-\begin{equation}
-\mathbf{E}[n_i, i \in[1, N_s]] = \frac{\sum^i_{j=0} {n_{outlier}\choose j} {D-n_{outlier}\choose i-j} j}{{D\choose i}}
-\end{equation}
+\begin{align}
+\mathbf{E}[n_i, i \in[1, N_s]] &= \frac{\sum^i_{j=0} {n_{outlier}\choose j} {D-n_{outlier}\choose i-j} j}{{D\choose i}} \\
+&= \frac{n_{outlier}i}{D}
+\end{align}
 
@@
     \hline
-    Random            & & 0.32 & 0.11 & 0.49 \\
+    Random            & & 0.32 & 0.14 & 0.49 \\
     \hline

### 2021-09-18 publications/iaai21/dora_outlier_detection_iaai.tex
removed my TODO flag; removed italics around z-band to just around z-
added des precision numbers to tab:faw_results
added DES combined plot

@@
 \subsubsection{Astrophysics: objects in Dark Energy Survey.}
-\todo{Umaa}
 
@@
 
-We used publicly-available photometry from the $g-$, $r-$, $i-$ and $z-band$ DES exposures.  We
+We used publicly-available photometry from the $g-$, $r-$, $i-$ and $z-$band DES exposures.  We
 transformed the photometry into luptitudes\footnote{A
@@
     \hline
-    PCA               & & 0.41 & 0.44 & \textbf{0.84} \\
-    DEMUD             & & 0.41 & 0.44 & \textbf{0.84} \\
+    PCA               & \textbf{0.42} & 0.41 & 0.44 & \textbf{0.84} \\
+    DEMUD             & ---- & 0.41 & 0.44 & \textbf{0.84} \\
     \hline
-    RX                & & 0.43 & ---- & 0.82 \\
-    LRX               & & ---- & 0.33 & 0.56 \\
+    RX                & 0.40 & 0.43 & ---- & 0.82 \\
+    LRX               & ----& ---- & 0.33 & 0.56 \\
     \hline
-    Iso. Forest  & & 0.46 & \textbf{0.56} & 0.74 \\
+    Iso. Forest  & 0.34 & 0.46 & \textbf{0.56} & 0.74 \\
     \hline
-    PAE               & & 0.41 & 0.44 & 0.83 \\
-    Neg. Sampling & & \textbf{0.49} & 0.33 & 0.43 \\
+    PAE               & 0.35 & 0.41 & 0.44 & 0.83 \\
+    Neg. Sampling & 0.32 & \textbf{0.49} & 0.33 & 0.43 \\
     \hline
-    Random            & & 0.32 & 0.14 & 0.49 \\
+    Random            & 0.26 & 0.32 & 0.14 & 0.49 \\
     \hline
@@
 \end{table}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{figures/des_combined_plot.pdf}
+    \caption{Number of true outliers ranked in top $N$ selections for 
+    the astrophysics (DES) dataset.}
+    \label{fig:des_results}
+\end{figure}
 

### 2021-09-18 publications/iaai21/dora_outlier_detection_iaai.tex
added DES info to tab:datasets
added small paragraph for DES results

@@
     \hline
-    Astrophysics &   &   &   \\
+    Astrophysics & 100,000   &  25,339 &  74,661 \\
     Earth & 6,757 & 37 & 76  \\
@@
 
-\todo{Umaa: add results for astro use case.}
+For the astrophysics dataset (Figure~\ref{fig:des_results}, DEMUD was omitted due to computational time 
+and LRX was omitted as it applies only to image data.  Of the remaining methods, 
+PCA achieved the highest precision, following by RX.  Negative sampling 
+performs well initially before its performance drops off.  The PAE finds the most outliers 
+overall.
 

### 2021-09-18 publications/iaai21/dora_outlier_detection_iaai.tex
for DES data tables, changed num inliers to ----
Add authors. #42

@@
 \pdfinfo{
-/Title (AAAI Press Formatting Instructions for Authors Using LaTeX -- A Guide)
+/Title (Domain-agnostic Outlier Ranking Algorithms (DORA)---A Configurable Pipeline for Facilitating Outlier Detection in Scientific Datasets)
 /Author (AAAI Press Staff, Pater Patel Schneider, Sunil Issar, J. Scott Penberthy, George Ferguson, Hans Guesgen, Francisco Cruz, Marc Pujol-Gonzalez)
@@
 \iffalse
-\title{AAAI Press Formatting Instructions \\for Authors Using \LaTeX{} --- A Guide }
+\title{Domain-agnostic Outlier Ranking Algorithms (DORA)---A Configurable Pipeline for Facilitating Outlier Detection in Scientific Datasets}
 \author{
@@
     % Authors
-    First Author Name,\textsuperscript{\rm 1}
-    Second Author Name, \textsuperscript{\rm 2}
-    Third Author Name \textsuperscript{\rm 1} \\
+    Hannah Kerner,\textsuperscript{\rm 1}
+    Umaa Rebbapragada, \textsuperscript{\rm 2}
+    Kiri Wagstaff, \textsuperscript{\rm 2} 
+    Steven Lu, \textsuperscript{\rm 2} 
+    Bryce Dubayah, \textsuperscript{\rm 1} \\
+    Eric Huff, \textsuperscript{\rm 2} 
+    Jake Lee, \textsuperscript{\rm 2} 
+    Vinay Raman, \textsuperscript{\rm 3}
+    Sakshum Kulshrestha \textsuperscript{\rm 1}  \\
 }
@@
     % Affiliations
-    \textsuperscript{\rm 1} Affiliation 1 \\
-    \textsuperscript{\rm 2} Affiliation 2 \\
-    firstAuthor@affiliation1.com, secondAuthor@affilation2.com, thirdAuthor@affiliation1.com
+    \textsuperscript{\rm 1} University of Maryland, College Park \\
+    \textsuperscript{\rm 2} Jet Propulsion Laboratory, 
+    									   California Institute of Technology \\
+    \textsuperscript{\rm 3} Montgomery Blair High School \\
+    \{hkerner\}@umd.edu, \{umaa.d.rebbapragada,
+    										kiri.l.wagstaff,
+    										you.lu,
+    										eric.m.huff\}@jpl.nasa.gov
 }
@@
     \hline
-    Astrophysics & 100,000   &  25,339 &  74,661 \\
+    Astrophysics & 100,000   &  25,339 &  ---- \\
     Earth & 6,757 & 37 & 76  \\

### 2021-09-18 publications/iaai21/dora_outlier_detection_iaai.tex
Add discussion of overall algorithm performance. #42

@@
 
-For the astrophysics dataset (Figure~\ref{fig:des_results}, DEMUD was omitted due to computational time 
+For the astrophysics dataset (Figure~\ref{fig:des_results}), DEMUD was omitted due to computational time 
 and LRX was omitted as it applies only to image data.  Of the remaining methods, 
@@
 
-\paragraph{Algorithm performance.} \todo{Discuss overall results from
- experiments and that we left hyperparameter tuning and extensive
- evaluation for future work.}
+\paragraph{Algorithm performance.} There was no one algorithm that 
+consistently had the best performance across all four datasets. PCA had
+the best performance overall for the astrophysics and F-MNIST datasets, 
+while negative sampling and isolation forest had the best performance for
+the Earth science and planetary datasets respectively. [This shows why
+it is best for us to include a variety of algorithms because the best 
+algorithm will vary depending on the dataset.] The purpose of this study
+was to demonstrate how DORA could be used to facilitate comprehensive
+outlier detection experiments and compare results across datasets from 
+different domains. Thus we chose hyperparameters based on common 
+practice and sense and leave hyperparameter tuning, which would likely 
+improve results for each dataset but would require
+creation of additional validation sets, for future work.
 

### 2021-09-18 publications/iaai21/dora_outlier_detection_iaai.tex
Add discussion of overall algorithm performance. #42

--- 2021-09-18 publications/iaai21/dora_outlier_detection_iaai.tex
@@
 
-\paragraph{Algorithm performance.} \todo{Discuss overall results from
- experiments and that we left hyperparameter tuning and extensive
- evaluation for future work.}
+\paragraph{Algorithm performance.} There was no one algorithm that 
+ had the best performance across all four datasets. PCA had
+the best performance for the astrophysics and F-MNIST datasets, 
+while negative sampling and isolation forest was best for
+the Earth science and planetary datasets respectively. This illustrates
+the importance of including a diverse set of algorithms and tools for
+easily inter-comparing them in DORA, since the best
+algorithm will vary for different datasets. The purpose of this study
+was to demonstrate how DORA could be used to facilitate 
+outlier detection experiments and compare results across datasets from 
+different domains. Thus we chose hyperparameters based on common 
+practice and sense and left hyperparameter tuning, which would likely 
+improve results for each dataset but would require
+creation of additional validation sets, for future work.
 

### 2021-09-18 publications/iaai21/dora_outlier_detection_iaai.tex
Add discussion of evaluation metrics and experiments. #42

@@
 
-\paragraph{Evaluation in outlier detection.}  \todo{Discuss why evaluation
-in outlier detection applications is hard since we often don't have complete
-labeled data for testing and labels can be subjective.}
+\paragraph{Evaluation in outlier detection.} Prior work has emphasized
+the difficulty of creating standardized metrics for outlier detection that 
+represents how models will perform in real world settings while also enabling
+intercomparison between datasets~\cite{campos2016evaluation}. We chose two
+complementary metrics with this in mind: precision at $N=n_{outliers}$, which
+measures the fraction of selections that are known outliers when the number of
+selections is equivalent to the number of outliers, and
+ Mean Discovery Rate, which measures the fraction of selections that are known outliers on average. 
+Designing experiments to evaluate outlier detection methods for real-world
+use cases is also difficult because it is difficult, or sometimes impossible, 
+to obtain labeled samples of outliers, inliers, or both for evaluation. 
+In addition, labels are often subjective or uncertain, 
+especially in the case of scientific datasets. For example, a dataset of known
+outliers was available for the astrophysics dataset from human annotation
+in prior work, but the remainder of samples in the dataset used for evaluation
+were not known to be inliers or outliers. This can result in evaluation metrics
+ that are deceptively low because unlabeled samples that might actually be
+ outliers (as was found to be common in prior work~\cite{wagstaff:des-anom20})
+ are counted as false positives.
 

### 2021-09-19 publications/iaai21/dora_outlier_detection_iaai.tex
Reorder table 1 by size. #42
Update random results for astro and fmnist. #42
Trim length by making text more concise. #42

@@
      (astrophysics), and 3) novelty-guided onboard targeting of instruments
-      for Mars rovers (planetary science). [We showed that ...]
+      for Mars rovers (planetary science). \todo{[We showed that ...]}
 \end{abstract}
@@
 Methods for outlier detection have been 
-surveyed extensively in prior work
+surveyed extensively and can be differentiated primarily based on how they 
+score outliers
 ~\cite{pimentel2014review,chandola2009anomaly,markou2003novelty,
 markou2003novelty2}.
-Methods can be differentiated primarily based on how they define outlierness.
 %
-Reconstruction-based methods aim to construct a model of a dataset by learning 
-a mapping between the data samples and a lower-dimensional representation that
+Reconstruction-based methods construct a model of a dataset by learning 
+a mapping between the input data and a lower-dimensional representation that
  minimizes the loss between the input and its reconstruction from the 
  low-dimensional representation~\citep{kerner2020comparison}. The 
- reconstruction error is used to compute the outlier score because samples
+ reconstruction error is used as the outlier score because samples
   that are unlike the data used to fit the model will be more poorly 
@@
 %
- Distance-based methods score outliers based on their distance from some
+ Distance-based methods score outliers based on their distance from a
 ``background'' which can be defined in a variety of ways. For example,
@@
 %
-Sparsity-based methods, such as the isolation forest~\cite{liu2008isolation}
-and local outlier factor\cite{breunig2000lof},
- score outliers based on how isolated or sparse they
+Sparsity-based methods such as isolation forest~\cite{liu2008isolation}
+and local outlier factor~\cite{breunig2000lof}
+ score outliers based on how isolated or sparse samples 
 are in a given feature space. 
@@
 distribution or probability density of a dataset and score samples based on
-their likelihood under the learned distribution. Examples include the probabilistic
+their likelihood. Examples include the probabilistic
 autoencoder, which scores samples based on the log likelihood under the learned
@@
 %
-Other methods treat detection of outliers as a supervised classification 
-problem, usually with only one class constituted by the inliers or 
+Other methods formulate outlier detection as supervised classification, 
+usually with only one class constituted by the inliers or 
 known normal samples.
@@
 
-\paragraph{Data types.} We chose to initially implement data loaders for four
+\paragraph{Data types.} We chose to implement data loaders for four
 data types that are commonly used by the machine learning and domain science
 communities: time series, feature vectors, images (grayscale or RGB), and 
-$N$-band rasters. $N$-band rasters are essentially images in which every pixel
-is associated with a location defined by a geographic coordinate reference 
-system (e.g., latitude/longitude in degrees). Most satellite data are 
-distributed as rasters; common formats include GeoTIFF, NetCDF, and 
-HDF. A data loader for each data type locates the data by the path(s)
+$N$-band rasters. $N$-band rasters are images or grids in which every pixel
+is associated with a location (e.g., latitude/longitude in degrees); 
+most satellite data are distributed as rasters.
+A data loader for each data type locates the data by the path(s)
 defined in the configuration file and loads samples into a dictionary of numpy
- arrays indexed by the sample id. This \texttt{data\_dict} is then passed to each of the ranking algorithms specified in 
- the configuration file.
+ arrays indexed by the sample id. This \texttt{data\_dict} is then passed to each of the ranking algorithms.
 
-\paragraph{Outlier ranking algorithms.} We implemented 7 algorithms for 
-scoring and 
-ranking samples by outlierness. All algorithms are unsupervised.
+\paragraph{Outlier ranking algorithms.} We implemented 7 unsupervised 
+algorithms for scoring and 
+ranking samples by outlierness. 
 We chose these algorithms to include a diverse
-set of approaches to scoring outliers since different algorithms may have 
-better performance for different datasets and use cases. 
+set of approaches to scoring outliers since different algorithms may  
+perform better for different datasets and use cases. 
 %We describe each
@@
 
-We constructed three datasets to evaluate the utility of DORA and the
-performance of each ranking algorithm for a variety of scientific domains
+We constructed three datasets to evaluate the utility of DORA and 
+algorithm performance for a variety of scientific domains
  (astrophysics, planetary science, and Earth science). We also included a 
  benchmark dataset that uses MNIST and Fashion-MNIST. Table \ref{tab:datasets}
- summarizes the number of unlabeled samples used for training and labeled
- samples used for evaluation for each dataset. We describe each in detail
- below.
+ summarizes the number of unlabeled samples used for training evaluation for
+ each dataset. We describe each dataset in detail below.
 
 \subsubsection{Astrophysics: objects in Dark Energy Survey.}
-
 Astronomical data sets are large and growing. Large modern optical
-imaging surveys are producing catalogs for of order $10^8$ stars
+imaging surveys are producing catalogs of order $10^8$ stars
 and galaxies, with dozens or hundreds of distinct measured features
@@
 astrophysical anomalies from non-astrophysical sources like detector artifacts or
-satellite trails is a challenging problem for current
-methods.
+satellite trails is a challenging problem for current methods.
 
-The Dark Energy Survey (DES), an ongoing imaging survey of
-$5000~{\rm deg}^2$ of the southern sky conducted from the Cerro-Tololo
-Inter-American Observatory in Chile \cite{DESY1}. The galaxy catalogs produced from
-DES imaging have provided some of the strongest constraints to date on
-the physical properties of dark energy and the accelerated expansion
-of the Universe. The first version of this catalog, released in June 2018, incorporated
-only cuts on signal-to-noise, resolution, masks against known detector
+The Dark Energy Survey (DES) is an ongoing imaging survey of
+$5000~{\rm deg}^2$ of the southern sky from the Cerro-Tololo
+Inter-American Observatory in Chile \cite{DESY1}. The resulting galaxy catalogs produced have provided some of the strongest constraints to date on
+the physical properties of dark energy and accelerated expansion
+of the Universe. The first version of this catalog, released June 2018, incorporated
+only cuts on signal-to-noise and resolution, masks against known detector
 anomalies and data quality indicators, and the automated data quality
-flags produced during processing. In December 2019, the full catalog
-was released after 18 months of extensive manual vetting. Therefore,
-we were able to use the second version of the catalog as a validation
-set for anomaly detection on the first version.
+flags produced during processing to filter outliers. 
+In December 2019, the full catalog
+was released after 18 months of extensive manual vetting. We used the samples 
+that were removed in the second version of the catalog as a set of known
+outliers for evaluating anomaly detection methods on the first version.
 
-We compared all methods on a data set that consists of 100K galaxy objects
+We compared all methods on a dataset of 100K galaxy objects
 observed by the Dark Energy Survey (DES) sampled from the initial June 2018 release.
-25,339 objects from this 100K set that did not appear in the later Dec 2019
-release were likely eliminated during the manual vetting process and were labeled as outliers.
-Note, that the remaining 74,661 objects may also contain outliers, but we assume them to be
-inliers in this experiment.
-
-We used publicly-available photometry from the $g-$, $r-$, $i-$ and $z-$band DES exposures.  We
+We labeled the 25,339 objects from this 100K set that did not appear in the later December 2019
+release, thus were likely eliminated during the manual vetting process, 
+as outliers.
+While the remaining 74,661 objects may also contain outliers, we assume them to be inliers in this experiment.
+We used publicly-available photometry from the $g-$, $r-$, $i-$ and $z-$band DES exposures.
+We
 transformed the photometry into luptitudes\footnote{A
@@
   traditional astronomical magnitudes for bright sources, but which
-  gracefully handles non-detections and negative fluxes.}.  The inputs features used
+  gracefully handles non-detections and negative fluxes.}.  The input features 
   were the $r$-band luptitude, colors computed as band differences between
@@
 $D_{1343-1693}$ ($n=992$) and ``assessment'' set $D_{1694-1703}$
-($n=58)$ for evaluation.
-
-We collaborated with an MSL science planning team member to
+($n=58)$ for evaluation. We collaborated with an MSL science team member to
 independently review the targets in $D_{1694-1703}$ and identify those
-that were 
 considered novel by the mission ($n_{outlier} = 9$).  Our goal for
@@
 
-\subsubsection{Earth: satellite time series for ground-truth observations.}
+\subsubsection{Earth: satellite time series for ground observations.}
 Many Earth science applications using satellite Earth observation (EO) data
- require ground-truth observations for understanding and modeling ground-
+ require ground-truth observations for identifying and modeling ground-
  identified objects in the satellite observations. These ground observations
@@
  learning models. For example, a model trained to classify crop types in
- satellite observations requires geo-referenced labels of crop type annotated
- in the field~\cite{tseng2021learning}. A widespread challenge of using ground-
- annotated datasets is that there are often points with erroneous location or 
+ satellite observations requires ground-annotated labels of crop type~\cite{tseng2021learning}. A widespread challenge for ground-
+ annotated labels is that there are often points with erroneous location or 
  label information (e.g., due to GPS location error or human entry error) that
- need to be cleaned before the dataset can be used for machine learning or
- other downstream use cases. Automatically detecting these outliers could save 
+ need to be cleaned before the labels can be used for machine learning or
+ other downstream uses. Automatically detecting these outliers could save 
  substantial time required for cleaning datasets and improve the performance of
  downstream analyses that rely on high-quality datasets.
- %
- We used a dataset of ground
- observations of fall army worm (FAW) in maize crops collected by the UN Food
+
+ We used a dataset of ground annotations of maize crops collected by the UN Food
  and Agriculture Organization (FAO). This dataset includes 6,757 samples
@@
  such as water, buildings, or forests. We constructed an evaluation set 
- consisting of all samples in Kenya ($n=113$) and manually annotated whether 
- each sample was considered an inlier (i.e., located in a crop field) or not 
- using high-resolution Planet satellite images in Collect Earth 
- Online~\cite{planetdata} ($n_{inlier}=76$, $n_{outlier}=37$). We used Google
+ of all samples in Kenya ($n=113$) and manually annotated whether 
+ each sample was located in a crop field (inlier) or not (outlier)
+ using high-resolution satellite images in Collect Earth 
+ Online ($n_{inlier}=76$, $n_{outlier}=37$). We used Google
  Earth Engine to export the Sentinel-1 synthetic aperture radar (SAR) monthly
- time series for each sample location, using data from year each sample was 
+ time series for each sample location, using data from the year each sample was 
  collected. We used SAR data because it is sensitive to ground texture and
   penetrates clouds, which is important for the often-cloudy region covered
-  by the dataset. Our goal for this application is to assess how well the 
+  by the dataset. Our goal for this application was to assess how well the 
   selections made by each algorithm correlate with outliers determined by
@@
 \subsubsection{Benchmark: MNIST and Fashion-MNIST.}
-We included MNIST and F-MNIST to demonstrate the capabilities of DORA with a
+We used MNIST and F-MNIST to demonstrate the capabilities of DORA with a
 traditional benchmark dataset. We used 60,000 images from 
-F-MNIST as the training set and a combined set of 2,000 images from MNIST and 
-F-MNIST as the evaluation set. Well-discriminating algorithms should assign 
-higher outlier scores to samples from MNIST than F-MNIST.
+F-MNIST as the training set and a test set of 1,000 images from MNIST (outliers)
+ and 1,000 from F-MNIST (inliers). 
+ %Well-discriminating algorithms should assign 
+%higher outlier scores to samples from MNIST than F-MNIST.
 
 \begin{table}
-  \caption{Number of samples in the training and test sets for each dataset.}
+  \caption{Number of samples in the training and test sets for each dataset,
+  ordered by dataset size.}
   \label{tab:datasets}
@@
     \hline
-    Astrophysics & 100,000   &  25,339 &  ---- \\
+    Planetary & 992 &  9 & 49 \\
     Earth & 6,757 & 37 & 76  \\
-    Planetary & 992 &  9 & 49 \\
     F-MNIST & 60,000 & 1,000 & 1,000  \\
+    Astrophysics & 100,000   &  25,339 &  74,661 \\
     \hline
@@
     \hline
-    Random            & 0.26 & 0.32 & 0.14 & 0.49 \\
+    Random            & 0.25 & 0.32 & 0.14 & 0.50 \\
     \hline

### 2021-09-19 publications/iaai21/dora_outlier_detection_iaai.tex
Trim text, fill in abstract/intro todos. #42

@@
 \begin{abstract}
-Automatic detection of out-of-distribution (OOD) samples or features is
-universally needed when working with scientific datasets. For example,
-OOD detection methods can be used for cleaning datasets or flagging
-novel samples in order to guide instrument acquisition or scientific analysis.
-We present Domain-agnostic Outlier Ranking Algorithms (DORA) to provide
-a configurable pipeline that makes it easy for scientists to apply and evaluate
+Automatic detection of outliers is
+universally needed when working with scientific datasets, e.g., for cleaning 
+datasets or flagging novel samples to guide instrument acquisition or
+ scientific analysis.
+We present Domain-agnostic Outlier Ranking Algorithms (DORA),
+a configurable pipeline that facilitates application and evaluation of
 outlier detection methods in a variety of domains. DORA allows users to 
-configure outlier detection experiments by specifying the location of their 
+configure experiments by specifying the location of their 
 dataset(s), the input data type, feature extraction methods, and which 
@@
 or feature vector input data types and outlier detection methods that include
- Isolation Forest, Discovery via Eigenbasis Modeling of Uninteresting Data 
- (DEMUD), principal component analysis (PCA), Reed-Xiaoli (RX) detector, 
- Local RX, negative sampling, and probabilistic autoencoder. Given a set of
-  data samples, each algorithm assigns an outlier score to each sample. DORA
-   provides results organization and visualization modules to help users 
+ Isolation Forest, DEMUD, PCA, RX detector, 
+ Local RX, negative sampling, and probabilistic autoencoder. Each algorithm assigns an outlier score to each data sample. DORA
+   provides results interpretation modules to help users 
    process the results, including sorting samples by outlier score, evaluating 
-   outlier recall for a set of known/labeled outliers, clustering groups of 
-   similar outliers together, and causal graphs. To demonstrate the 
-   cross-domain applicability of DORA, we conducted outlier detection
-    experiments for three real-world use cases: 1) cleaning noisy ground-truth 
-    datasets (Earth Science), 
-     2) flagging non-astrophysical objects in the Dark Energy Survey 
-     (astrophysics), and 3) novelty-guided onboard targeting of instruments
-      for Mars rovers (planetary science). \todo{[We showed that ...]}
+   the fraction of known outliers in $n$ selections, clustering groups of 
+   similar outliers together, and web visualization. We demonstrated how DORA
+   facilitates application, evaluation, and interpretation of outlier detection
+   methods by performing
+    experiments for three real-world datasets from Earth science, planetary
+  science, and astrophysics, as well as one benchmark dataset 
+  (MNIST/Fashion-MNIST). We found that no single algorithm
+      performed best across all datasets, underscoring the need for a tool 
+      that enables comparison of multiple algorithms.
 \end{abstract}
@@
  would benefit from a tool that made it easy for them to apply popular outlier
- detection algorithms to their science datasets. We created DORA (Domain-
- agnostic Outlier Ranking Algorithms) to provide a tool for applying outlier 
+ detection algorithms to their science datasets. We created DORA 
+ (Domain-agnostic Outlier Ranking Algorithms) 
+ to provide a tool for applying outlier 
  detection algorithms to a variety of scientific data sets with minimal coding
@@
  Local RX, negative sampling~\cite{sipple:neg-sampling20}, and probabilistic 
- autoencoder. Given a set of data samples, each algorithm assigns an outlier 
- score to each sample. DORA provides results organization and visualization 
+ autoencoder (PAE). 
+ Each algorithm assigns an outlier 
+ score to each sample in a given dataset. 
+ DORA provides results organization and visualization 
  modules to help users process the results, including sorting samples by outlier 
  score, evaluating outlier recall for a set of known/labeled outliers, clustering 
- groups of similar outliers together, and causal graphs. To demonstrate the 
- cross-domain applicability of DORA, we conducted outlier detection experiments 
- for three real-world use cases: 1) cleaning noisy ground-truth datasets (Earth 
- Science), 2) flagging non-astrophysical objects in the Dark Energy Survey 
- (astrophysics), and 3) novelty-guided onboard targeting of instruments for Mars 
- rovers (planetary science). \todo{[We showed that ...]}
+ groups of similar outliers together, and web visualization. 
+ We demonstrated how DORA
+   facilitates application, evaluation, and interpretation of outlier detection
+   methods by performing
+    experiments for three real-world datasets from Earth science, planetary
+  science, and astrophysics, as well as one benchmark dataset 
+  (MNIST/Fashion-MNIST). We found that no single algorithm
+      performed best across all datasets, underscoring the need for a tool 
+      that enables comparison of multiple algorithms.
 
@@
 \subsubsection{Benchmark: MNIST and Fashion-MNIST.}
-We used MNIST and F-MNIST to demonstrate the capabilities of DORA with a
+We used MNIST and F-MNIST to demonstrate DORA with a
 traditional benchmark dataset. We used 60,000 images from 
-F-MNIST as the training set and a test set of 1,000 images from MNIST (outliers)
- and 1,000 from F-MNIST (inliers). 
+F-MNIST as the training set and a test set of 1,000 images each from MNIST (outliers) and F-MNIST (inliers). 
  %Well-discriminating algorithms should assign 

### 2021-09-19 publications/iaai21/dora_outlier_detection_iaai.tex
Write conclusion. #42

@@
 In astrophysics, 
-outlier detection methods have been employed for identifying astrophysical
+outlier detection methods have been used to identify astrophysical
 objects with unique characteristics~\cite{xiong2010anomaly,
@@
 lochner2021astronomaly}. 
-Examples of outlier detection applications in 
+Example outlier detection applications in 
 Earth science include detecting anomalous objects or materials
@@
 touati2020anomaly}, and ocean extremes~\cite{prochaska2021deep}. 
-Planetary science applications of outlier detection have mostly focused 
+Planetary science applications have mostly focused 
 on prioritizing samples with novel geologic or geochemical features for
@@
 %
-These examples show the promise of applying outlier detection methods in
+These examples show the benefit of applying outlier detection methods in
 a variety of real-world science use cases. However, the effort required to 
-apply and evaluate the variety of available algorithms is non-trivial and can 
+apply and evaluate the many available algorithms is non-trivial and can 
 be daunting for non-ML experts, thus impeding the uptake of outlier
@@
   \centering
-  \includegraphics[width=0.75\linewidth]{figures/doravis.png}
+  \includegraphics[width=0.9\linewidth]{figures/doravis.png}
   \caption{A screenshot of the DORA visualizer displaying results from the
@@
 \section{Datasets}
+\begin{table}
+  \caption{Number of samples in the training and test sets for each dataset,
+  ordered by dataset size.}
+  \label{tab:datasets}
+  \centering
+  \begin{tabular}{l|c|cc}
+    \hline
+    Dataset & Training & \multicolumn{2}{c}{Test}\\
+     & Unlabeled &  Outliers &  Inliers \\
+
+    \hline
+    Planetary & 992 &  9 & 49 \\
+    Earth & 6,757 & 37 & 76  \\
+    F-MNIST & 60,000 & 1,000 & 1,000  \\
+    Astrophysics & 100,000   &  25,339 &  74,661 \\
+    \hline
+  \end{tabular}
+\end{table}
 
@@
 \begin{table}
-  \caption{Number of samples in the training and test sets for each dataset,
-  ordered by dataset size.}
-  \label{tab:datasets}
+  \caption{Precision at $N=n_{outlier}$ for three applications and F-MNIST
+   benchmark; the best result for each data set is in bold.}
+  \label{tab:faw_results}
   \centering
-  \begin{tabular}{l|c|cc}
+  \begin{tabular}{l|cccc}
     \hline
-    Dataset & Training & \multicolumn{2}{c}{Test}\\
-     & Unlabeled &  Outliers &  Inliers \\
-
+    Algorithm & Astro. & Earth & Planetary & F-MNIST \\
     \hline
-    Planetary & 992 &  9 & 49 \\
-    Earth & 6,757 & 37 & 76  \\
-    F-MNIST & 60,000 & 1,000 & 1,000  \\
-    Astrophysics & 100,000   &  25,339 &  74,661 \\
+    PCA               & \textbf{0.42} & 0.41 & 0.44 & \textbf{0.84} \\
+    DEMUD             & ---- & 0.41 & 0.44 & \textbf{0.84} \\
+    \hline
+    RX                & 0.40 & 0.43 & ---- & 0.82 \\
+    LRX               & ----& ---- & 0.33 & 0.56 \\
+    \hline
+    Iso. Forest  & 0.34 & 0.46 & \textbf{0.56} & 0.74 \\
+    \hline
+    PAE               & 0.35 & 0.41 & 0.44 & 0.83 \\
+    Neg. Sampling & 0.32 & \textbf{0.49} & 0.33 & 0.43 \\
+    \hline
+    Random            & 0.25 & 0.32 & 0.14 & 0.50 \\
     \hline
@@
 \section{Results}
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{figures/planetary_combined_plot.pdf}
+    \caption{Number of true outliers ranked in top $N$ selections for 
+    planetary rover dataset.}
+    \label{fig:planetary_results}
+\end{figure}
+\begin{figure}
+    \centering
+    \includegraphics[width=\linewidth]{figures/faw_combined_plot.pdf}
+    \caption{Number of true outliers ranked in top $N$ selections for 
+    Earth science dataset.}
+    \label{fig:faw_results}
+\end{figure}
 The experimental setup for each dataset was to fit or train a model for
@@
 
-\begin{table}
-  \caption{Precision at $N=n_{outlier}$ for three applications and F-MNIST
-   benchmark; the best result for each data set is in bold.}
-  \label{tab:faw_results}
-  \centering
-  \begin{tabular}{l|cccc}
-    \hline
-    Algorithm & Astro. & Earth & Planetary & F-MNIST \\
-    \hline
-    PCA               & \textbf{0.42} & 0.41 & 0.44 & \textbf{0.84} \\
-    DEMUD             & ---- & 0.41 & 0.44 & \textbf{0.84} \\
-    \hline
-    RX                & 0.40 & 0.43 & ---- & 0.82 \\
-    LRX               & ----& ---- & 0.33 & 0.56 \\
-    \hline
-    Iso. Forest  & 0.34 & 0.46 & \textbf{0.56} & 0.74 \\
-    \hline
-    PAE               & 0.35 & 0.41 & 0.44 & 0.83 \\
-    Neg. Sampling & 0.32 & \textbf{0.49} & 0.33 & 0.43 \\
-    \hline
-    Random            & 0.25 & 0.32 & 0.14 & 0.50 \\
-    \hline
-  \end{tabular}
-\end{table}
-
 \begin{figure}
@@
     \label{fig:des_results}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{figures/planetary_combined_plot.pdf}
-    \caption{Number of true outliers ranked in top $N$ selections for 
-    planetary rover dataset.}
-    \label{fig:planetary_results}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{figures/faw_combined_plot.pdf}
-    \caption{Number of true outliers ranked in top $N$ selections for 
-    Earth science dataset.}
-    \label{fig:faw_results}
 \end{figure}
@@
 
-\paragraph{Algorithm performance.} There was no one algorithm that 
+\paragraph{Algorithm performance.} 
+There was no one algorithm that 
  had the best performance across all four datasets. PCA had
@@
 
-\paragraph{Evaluation in outlier detection.} Prior work has emphasized
+\paragraph{Evaluation in outlier detection.} 
+Prior work has emphasized
 the difficulty of creating standardized metrics for outlier detection that 
@@
 
-\paragraph{Path to deployment.} \todo{Hannah - describe pip-installable
-package, open-source for extension by the wider scientific community, etc.}
+\paragraph{Path to deployment.} 
+Our goal is for DORA to enable increased application and benefit of
+outlier detection methods in real-world scientific use cases. We have 
+designed DORA to make it as easy as possible for scientists to apply
+algorithms and to compare and interpret their results; users need only
+to specify the specifics of their data (e.g., path, data type) in a 
+configuration file to start running experiments and seeing results. We will
+make DORA publicly available as a pip-installable package to make it
+easy to integrate into existing scientific workflows. 
+The source code and example datasets used in this study will also be 
+open-sourced, enabling DORA to be improved and expanded by the 
+machine learning and domain science communities. 
+In addition, DORA will be infused into the scientific workflows for the
+three use cases we demonstrated results for in this study.
+%the use cases from this will also lead to infusion in real scenarios
 
-\section{Conclusions}
-\todo{
-Summarize conclusions and plans for future work
-[Note that these datasets could also be used by other researchers as OOD
- benchmark tasks?]
- \todo{Can we make any of these datasets publicly available with the paper?}
- \todo{Future work: other algorithms, datasets, causal graphs.}
-}
+\section{Conclusion}
+The ability to automatically find outliers in large datasets is critical
+for a variety of scientific and real-world use cases. 
+We presented Domain-agnostic Outlier Ranking Algorithms (DORA),
+a configurable pipeline that facilitates application and evaluation of
+outlier detection methods in a variety of domains. DORA minimizes the
+coding and ML expertise required for domain scientists since users
+need only to specify their experiment details in a configuration file
+to get results from all available algorithms. This is particularly important
+because the experiments for three cross-domain science datasets in this
+study showed that no one algorithm performs best for all datasets.
+DORA will be publicly accessible as a python package to make it
+easy to integrate into existing scientific workflows. 
+The will be
+open-sourced to enable continued improvement and expansion of DORA
+to serve the needs of the science community. The datasets used in this
+study will also be public and can serve as real-world benchmarks 
+for future outlier detection methods. In future work, we will
+continue to improve DORA based on the experience of deploying it in
+the workflows of the domain scientists associated with the datasets in
+this study and add additional interpretation modules 
+including causal inference graphs.
 

### 2021-09-19 publications/iaai21/dora_outlier_detection_iaai.tex
Trim text to 6 pages. #42

@@
 application of outlier detection methods~\cite{zhao2019pyod}, they cover
-limited data formats and algorithms. The aim of DORA is to fill the need 
+limited data formats and algorithms. DORA aims to fill the need 
 for tools that faciliate application, evaluation, and interpretation of outlier
@@
 \section{Methods}
-
 %\subsection{DORA configurable pipeline} 
-Figure \ref{fig:dora} illustrates the architecture of DORA. It includes 
-data loading, feature extraction, outlier ranking, results organization, and 
-results visualizaiton modules. The DORA pipeline is implemented in Python 3.
+Figure \ref{fig:dora} illustrates the architecture of DORA including 
+data loading, feature extraction, outlier ranking, and results organization
+and visualization modules. 
 In order to improve the readability and execution speed of the code, we adopted
@@
 modules that follow the DORA API. Experimental settings are controlled by 
-configuration files. In a configuration file, users can specify the input data, 
+configuration files in which users can specify the input data, 
 feature extraction methods, normalization method, outlier ranking methods, and 
 results organization methods. 
+DORA is implemented in Python 3.
 
@@
  transformation from the principal subspace~\citep{kerner2020comparison}.
-The DEMUD algorithm~\citep{wagstaff:demud13} differs from other
+ DEMUD~\citep{wagstaff:demud13} differs from other
 outlier ranking methods: instead of independently scoring all
@@
 item.  DEMUD's goal is to identify diverse outliers and avoid
-redundant selections.  Once an outlier is encountered, repeated
+redundant selections.  Once an outlier is found, repeated
 occurrences of that outlier are deprioritized.  Methods that score
-outliers independently maximize coverage of outliers, while DEMUD
-maximizes the fast discovery of distinct outlier types.
+samples independently maximize coverage of outliers, while DEMUD
+maximizes fast discovery of distinct outlier types.
 
@@
 Reed-Xiaoli (RX) detector is commonly used for anomaly detection in
-multispectral and hyperspectral remote sensing; RX uses the Mahalanobis
- distance between a sample and a background dataset mean and covariance
- to score samples~\citep{reed1990adaptive}. The local variant of RX (Local
- RX or LRX) can be used for image or raster data types and scores each pixel in
+multispectral and hyperspectral remote sensing. RX scores samples using
+ the Mahalanobis
+ distance between a sample and a background mean and covariance
+ ~\citep{reed1990adaptive}. The local variant of RX (Local
+ RX or LRX) can be used for image or raster data and scores each pixel in
   an image with respect to a window ``ring'' of pixels surrounding 
   it~\citep{molero2013analysis}. 
-  LRX requires two parameters that define the ring: an outer window which 
-  defines the size of the window surrounding the pixel and an inner window
-   which defines the size of the window around the target pixel to exclude from
-   the background distribution.
+  LRX requires two parameters to define the size of the outer window 
+  surrounding the pixel and the inner window
+   around the target pixel to exclude from the background distribution.
 
@@
 Each of the outlier ranking algorithms returns an array containing the sample
-index, outlier score, and selection index (index of each sample after sorting
+index, outlier score, and selection index (index after sorting
 by outlier score). DORA provides organization and visualization modules
@@
  to assess how well algorithm selections correlate with known outliers. This is
- done by plotting the number of true outliers vs. number of selections made. 
+ done by plotting the number of known outliers vs. number of selections made. 
 We provide a module for plotting histograms of outlier scores to visualize the
@@
  long-tailed). 
+We developed a desktop application to easily visualize DORA results with the
+Electron application framework and React frontend library. 
+This enables fast and easy comparison of the results from
+different methods. Figure \ref{fig:doravis} shows a screenshot of the
+``Aggregate Table'' view, which displays all algorithm results 
+side-by-side.
+
+%We developed a desktop application to easily visualize DORA results with the
+%Electron application framework and React frontend library. The application loads
+%the DORA configuration file to locate the dataset and result CSVs. Then,
+%it displays the ranked samples and their scores in a table sorted by their
+%selection order. This allows for fast and easy comparison of the results of
+%different methods. Figure \ref{fig:doravis} shows a screenshot of the
+%``Aggregate Table'' view, which displays all results from different algorithms
+%side-by-side.
 
@@
 
-We developed a desktop application to easily visualize DORA results with the
-Electron application framework and React frontend library. The application loads
-the DORA configuration file to locate the dataset and result CSVs. Then,
-it displays the ranked samples and their scores in a table sorted by their
-selection order. This allows for fast and easy comparison of the results of
-different methods. Figure \ref{fig:doravis} shows a screenshot of the
-``Aggregate Table'' view, which displays all results from different algorithms
-side-by-side.
+\section{Datasets}
+We constructed three datasets to evaluate the utility of DORA and 
+algorithm performance for a variety of scientific domains
+ (astrophysics, planetary science, and Earth science). We also included a 
+ benchmark dataset that uses MNIST and Fashion-MNIST. Table \ref{tab:datasets}
+ summarizes the number of unlabeled samples used for training evaluation for
+ each dataset. We describe each dataset in detail below.
 
-\section{Datasets}
  \begin{table}
@@
 \end{table}
-
-We constructed three datasets to evaluate the utility of DORA and 
-algorithm performance for a variety of scientific domains
- (astrophysics, planetary science, and Earth science). We also included a 
- benchmark dataset that uses MNIST and Fashion-MNIST. Table \ref{tab:datasets}
- summarizes the number of unlabeled samples used for training evaluation for
- each dataset. We describe each dataset in detail below.
 
@@
 Mars exploration is fundamentally an exercise in discovery with the
-goal of increasing our understanding of the history, evolution,
-composition, and currently active processes on Mars.  Outliers
+goal of increasing our understanding of Mars's history, evolution,
+composition, and currently active processes.  Outliers
 identified in Mars observations can inspire new discoveries and inform
@@
 on the Mars Science Laboratory (MSL) rover
-and employed the Rockster~\cite{burl:rockster16}
-method currently used by onboard rover software to identify candidate rock
+and employed Rockster~\cite{burl:rockster16}
+(currently used by onboard rover software) to identify candidate rock
 targets with an area of at least 100 pixels, yielding \num{1050}
@@
 \begin{table}
-  \caption{Precision at $N=n_{outlier}$ for three applications and F-MNIST
-   benchmark; the best result for each data set is in bold.}
+  \caption{Precision at $N=n_{outlier}$ for four datasets; the best result for each data set is in bold.}
   \label{tab:faw_results}
@@
 \section{Results}
-\begin{figure}
+\begin{figure}[h]
     \centering
     \includegraphics[width=\linewidth]{figures/planetary_combined_plot.pdf}
-    \caption{Number of true outliers ranked in top $N$ selections for 
+    \caption{Number of known outliers ranked in top $N$ selections for 
     planetary rover dataset.}
@@
 \end{figure}
-\begin{figure}
+
+\begin{figure}[h]
     \centering
     \includegraphics[width=\linewidth]{figures/faw_combined_plot.pdf}
-    \caption{Number of true outliers ranked in top $N$ selections for 
+    \caption{Number of known outliers ranked in top $N$ selections for 
     Earth science dataset.}
@@
 \end{figure}
+
+\begin{figure}[h]
+    \centering
+    \includegraphics[width=\linewidth]{figures/des_combined_plot.pdf}
+    \caption{Number of known outliers ranked in top $N$ selections for 
+    the astrophysics (DES) dataset.}
+    \label{fig:des_results}
+\end{figure}
+
 The experimental setup for each dataset was to fit or train a model for
@@
 labels of known outliers were available (Table \ref{tab:datasets}). 
-For each test set, we created a plot of the number of known/true outliers 
+For each test set, we created a plot of the number of known outliers 
 detected out of the top $N$ selections. We also 
@@
 when the number of selections is the same as the total number of outliers. 
-Precision at $N$ is the number of true outliers divided by the number of 
+Precision at $N$ is the number of known outliers divided by the number of 
 selections $N$~\citep{campos2016evaluation}.
@@
 
-For the Earth dataset, the negative sampling algorithm had the best performance
+For the Earth dataset, negative sampling had the best performance
 in both metrics. DEMUD, PCA, and PAE tied for the lowest precision at 
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{figures/des_combined_plot.pdf}
-    \caption{Number of true outliers ranked in top $N$ selections for 
-    the astrophysics (DES) dataset.}
-    \label{fig:des_results}
-\end{figure}
+\section{Discussion}
 
@@
     \includegraphics[width=\linewidth]{figures/fmnist_combined_plot.pdf}
-    \caption{Number of true outliers ranked in top $N$ selections for 
+    \caption{Number of known outliers ranked in top $N$ selections for 
     F-MNIST dataset.}
@@
 
-\section{Discussion}
-
 \paragraph{Algorithm performance.} 
-There was no one algorithm that 
+No one algorithm  
  had the best performance across all four datasets. PCA had
@@
 outlier detection experiments and compare results across datasets from 
-different domains. Thus we chose hyperparameters based on common 
-practice and sense and left hyperparameter tuning, which would likely 
-improve results for each dataset but would require
-creation of additional validation sets, for future work.
+different domains. Thus we did not perform hyperparameter tuning which
+could improve results for each dataset; we leave this for future work.
 
@@
 We thank the UN Food and Agriculture Organization (FAO) and Dr. Ritvik Sahajpal
-(University of Maryland/NASA Harvest) for providing the fall army worm ground
+(University of Maryland/NASA Harvest) for providing the maize ground
  observation dataset.

### 2021-09-19 publications/iaai21/dora_outlier_detection_iaai.tex
Comment out acknowledgements due to page limits, will add to camera ready later. #42

@@
     \textsuperscript{\rm 3} Montgomery Blair High School \\
-    \{hkerner\}@umd.edu, \{umaa.d.rebbapragada,
+    hkerner@umd.edu, \{umaa.d.rebbapragada,
     										kiri.l.wagstaff,
     										you.lu,
-    										eric.m.huff\}@jpl.nasa.gov
+    										eric.m.huff,
+    										jake.h.lee\}@jpl.nasa.gov
 }
@@
 
+%\section{Acknowledgments}
+%We thank the NASA Center Innovation Fund and the NASA Science Mission 
+%Directorate (under the NASA Cross-Divisional AI Use Case Demonstration grants)
+% for supporting this work.
+%
+%We thank the UN Food and Agriculture Organization (FAO) and Dr. Ritvik Sahajpal
+%(University of Maryland/NASA Harvest) for providing the maize ground
+% observation dataset.
+% 
+% We thank Raymond Francis for analyzing the Navcam rover images to
+%identify targets considered to be ``novel'' for evaluating outlier
+%detection in the planetary rover scenario.
+%
+%Part of this research was carried out at the Jet Propulsion
+%Laboratory, California Institute of Technology, under a contract with
+%the National Aeronautics and Space Administration.
+
 \bibliography{dora_references}
-
-\section{Acknowledgments}
-We thank the NASA Center Innovation Fund and the NASA Science Mission 
-Directorate (under the NASA Cross-Divisional AI Use Case Demonstration grants)
- for supporting this work.
-
-\todo{Acknowledge DES?}
-We thank the UN Food and Agriculture Organization (FAO) and Dr. Ritvik Sahajpal
-(University of Maryland/NASA Harvest) for providing the maize ground
- observation dataset.
-
-We thank Raymond Francis for analyzing the Navcam rover images to
-identify targets considered to be ``novel'' for evaluating outlier
-detection in the planetary rover scenario.
-
-Part of this research was carried out at the Jet Propulsion
-Laboratory, California Institute of Technology, under a contract with
-the National Aeronautics and Space Administration.
 

### 2021-10-17 dora_exp_pipeline/dora_data_loader.py
Unify JPL and UMD virtual computing environments (#44)
- Rely on data id instead of row index to match results to labels file, addressing an issue where os.listdir provides inconsistent ordering across machines
- Use data id from labels.csv instead of row number for comparison plots, allowing experiments to be run on a subset of the test data
- Ensure all data ids are strings

Co-authored-by: Kiri Wagstaff <kiri.l.wagstaff@jpl.nasa.gov>

@@ -177,7 +177,7 @@ class RasterPixelLoader(DataLoader):
                 img = np.reshape(img, [img.shape[0]*img.shape[1],
                                        img.shape[2]])
                 # set the ID to the index of the pixel
-                data_dict['id'] = range(img.shape[0])
+                data_dict['id'] = [str(i) for i in range(img.shape[0])]
                 data_dict['data'] = list(img)
         else:
             raise RuntimeError(f'File extension not supported. '
@@ -256,7 +256,7 @@ class FeatureVectorLoader(DataLoader):
         if dir_path.endswith('.h5'):
             # Load the .h5
             df = pd.read_hdf(dir_path)
-            data_dict['id'] = df.index
+            data_dict['id'] = df.index.astype(str)
             data_dict['data'] = df.values
 
         else:
@@ -295,7 +295,7 @@ class TimeSeriesLoader(DataLoader):
                 for row in csv_reader:
                     # Assumes the first column is the ID
                     # and all other columns are time steps
-                    data_dict['id'].append(row[0])
+                    data_dict['id'].append(str(int(row[0])))
                     data_dict['data'].append(np.array(row[1:]))
         else:
             raise RuntimeError(f'File extension not supported. '

### 2021-10-17 dora_exp_pipeline/dora_data_loader.py
Change raster loader to not flatten patches; change flattened pixel value extractor to scale based on min max #58

@@ -230,7 +230,7 @@ class RasterPatchLoader(DataLoader):
                             # append the patch coordinate as the id
                             data_dict['id'].append('%d-%d' % (i, j))
                             # append the patch data
-                            data_dict['data'].append(patch.flatten())
+                            data_dict['data'].append(patch)
                             j += stride
                     i += stride
         else:

### 2021-10-17 dora_exp_pipeline/dora_feature.py

@@ -133,7 +133,10 @@ class FlattenedPixelValuesExtractor(FeatureExtractor):
                 ret_features[ind, :] = data.flatten()
 
         if kwargs.get('normalize_pixels'):
-            ret_features = ret_features / 255.
+            min_val = np.min(ret_features)
+            max_val = np.max(ret_features)
+            scaled = np.subtract(ret_features, min_val) / (max_val - min_val)
+            ret_features = scaled
 
         return ret_features
 

### 2021-10-18 dora_exp_pipeline/dora_results_organization.py
remove used class #68

@@ -222,16 +222,6 @@ som_cluster = SOMCluster()
 register_org_method(som_cluster)
 
 
-class NodeBlock:
-    """
-    members is a list of node names
-    order is a single integer. Low-->high is left-->right
-    """
-    def __init__(self, members=None, order=None):
-        self.members = members
-        self.order = order
-
-
 def generate_causal_graphs(data_to_fit, data_to_cluster, cluster_groups,
                            out_dir, logger, seed):
     causal_tags = ['feature-%d' % col for col in range(len(data_to_cluster[0]))]

### 2021-10-18 dora_exp_pipeline/dora_results_organization.py
(1) add sparsity parameter; (2) improve graph visualization #68

@@ -149,7 +149,7 @@ class KmeansCluster(ResultsOrganization):
 
     def _run(self, data_ids, dts_scores, dts_sels, data_to_fit, data_to_score,
              outlier_alg_name, out_dir, logger, seed, top_n, n_clusters,
-             causal_graph):
+             causal_graph, sparsity):
         if not os.path.exists(out_dir):
             os.mkdir(out_dir)
             if logger:
@@ -177,7 +177,7 @@ class KmeansCluster(ResultsOrganization):
 
         if causal_graph:
             generate_causal_graphs(data_to_fit, data_to_cluster, groups,
-                                   out_dir, logger, seed)
+                                   out_dir, logger, seed, sparsity)
 
 
 kmeans_cluster = KmeansCluster()
@@ -190,7 +190,7 @@ class SOMCluster(ResultsOrganization):
 
     def _run(self, data_ids, dts_scores, dts_sels, data_to_fit, data_to_score,
              outlier_alg_name, out_dir, logger, seed, top_n, n_clusters,
-             causal_graph):
+             causal_graph, sparsity):
         if not os.path.exists(out_dir):
             os.mkdir(out_dir)
             if logger:
@@ -203,7 +203,8 @@ class SOMCluster(ResultsOrganization):
             data_to_cluster.append(data_to_score[dts_ind])
         data_to_cluster = np.array(data_to_cluster, dtype=float)
 
-        som = SOM(m=n_clusters, n=1, dim=len(data_to_cluster[0]))
+        som = SOM(m=n_clusters, n=1, dim=len(data_to_cluster[0]),
+                  random_state=seed)
         som.fit(data_to_cluster)
         groups = som.predict(data_to_cluster)
 
@@ -215,7 +216,7 @@ class SOMCluster(ResultsOrganization):
 
         if causal_graph:
             generate_causal_graphs(data_to_fit, data_to_cluster, groups,
-                                   out_dir, logger, seed)
+                                   out_dir, logger, seed, sparsity)
 
 
 som_cluster = SOMCluster()
@@ -223,8 +224,8 @@ register_org_method(som_cluster)
 
 
 def generate_causal_graphs(data_to_fit, data_to_cluster, cluster_groups,
-                           out_dir, logger, seed):
-    causal_tags = ['feature-%d' % col for col in range(len(data_to_cluster[0]))]
+                           out_dir, logger, seed, sparsity):
[…]

### 2021-10-19 publications/iaai21/dora_outlier_detection_iaai.tex

@@
   OOD detection systems is to identify samples that deviate from the majority
-   of samples in a dataset in an unsupervised manner (Pimentel et al., 2014). 
+   of samples in a dataset in an unsupervised manner~\cite{pimentel2014review}. 
    In machine learning, these methods are commonly used for identifying 
-   mislabeled or otherwise invalid samples in a dataset 
-   (e.g., Liang et al., 2018; Böhm et al., 2020).
-
-   When working with science datasets, OOD detection can be used for 
+   mislabeled or otherwise invalid samples in a dataset. When working with science datasets, OOD detection can be used for 
    cleaning datasets, e.g., flagging ground-truth labels with GPS or human
@@
    supervised classification algorithms.
+
   Despite wide differences in applications, data types, and dimensionality,
@@
   include PCA~\cite{jablonski2015principal}, 
-  autoencoders~\cite{richter2017safe,park2018multimodal}, and
+  autoencoders~\cite{richter2017safe}, and
   generative adversarial networks~\cite{akcay2018ganomaly}.
@@
 Probability distribution and density based methods estimate the underlying 
-distribution or probability density of a dataset and score samples based on
-their likelihood. Examples include the probabilistic
-autoencoder, which scores samples based on the log likelihood under the learned
+distribution or probability density of a dataset and score samples using
+ likelihood. Examples include the probabilistic
+autoencoder, which scores samples based on the log likelihood under the 
 latent space distribution~\cite{bohm2020probabilistic}, Gaussian mixture 
@@
 Other methods formulate outlier detection as supervised classification, 
-usually with only one class constituted by the inliers or 
+usually with only one class constituted by 
 known normal samples.
 Such methods include one-class support vector 
-machines~\cite{scholkopf1999support},
-support vector data description~\cite{pimentel2014review},
+machines~\cite{scholkopf1999support}
 and negative sampling~\cite{sipple:neg-sampling20}.
@@
 outlier detection methods have been used to identify astrophysical
-objects with unique characteristics~\cite{xiong2010anomaly,
-storey2020anomaly,hayat2021self} as well as data or modeling 
-artifacts in astronomical surveys~\citep{wagstaff:des-anom20,
-lochner2021astronomaly}. 
+objects with unique characteristics~\cite{hayat2021self} 
+as well as data or modeling 
+artifacts in astronomical surveys~\citep{wagstaff:des-anom20}. 
 Example outlier detection applications in 
 Earth science include detecting anomalous objects or materials
-~\cite{chang2002anomaly,zhou2016novel}, data artifacts or noise 
-~\cite{liu2017unsupervised,alvera2012outlier}, change~\cite{zhou2016novel,
-touati2020anomaly}, and ocean extremes~\cite{prochaska2021deep}. 
+~\cite{zhou2016novel}, data artifacts or noise 
+~\cite{liu2017unsupervised}, change~\cite{touati2020anomaly}, 
+and ocean extremes~\cite{prochaska2021deep}. 
 Planetary science applications have mostly focused 
 on prioritizing samples with novel geologic or geochemical features for
-follow-up targeting or analysis~\cite{kerner2020analysis,kerner2020comparison,
+follow-up targeting or analysis~\cite{kerner2020comparison,
 wagstaff:demud13}.
@@
   \centering
-  \includegraphics[width=0.9\linewidth]{figures/doravis.png}
+  \includegraphics[width=0.83\linewidth]{figures/doravis.png}
   \caption{A screenshot of the DORA visualizer displaying results from the
@@
 $5000~{\rm deg}^2$ of the southern sky from the Cerro-Tololo
-Inter-American Observatory in Chile \cite{DESY1}. The resulting galaxy catalogs produced have provided some of the strongest constraints to date on
+Inter-American Observatory in Chile~\cite{zuntz2018dark}. The resulting galaxy catalogs produced have provided some of the strongest constraints to date on
 the physical properties of dark energy and accelerated expansion
@@
  learning models. For example, a model trained to classify crop types in
- satellite observations requires ground-annotated labels of crop type~\cite{tseng2021learning}. A widespread challenge for ground-
+ satellite observations requires ground-annotated labels of crop type. A widespread challenge for ground-
  annotated labels is that there are often points with erroneous location or 
@@
  using high-resolution satellite images in Collect Earth 
- Online ($n_{inlier}=76$, $n_{outlier}=37$). We used Google
- Earth Engine to export the Sentinel-1 synthetic aperture radar (SAR) monthly
- time series for each sample location, using data from the year each sample was 
+ Online ($n_{inlier}=76$, $n_{outlier}=37$). We used the Sentinel-1 
+ synthetic aperture radar (SAR) monthly median
+ time series for each sample location from the year the sample was 
  collected. We used SAR data because it is sensitive to ground texture and
@@
 reported the Mean Discovery Rate (MDR) in the legend
-for each algorithm to give a quantitative comparison of the curves that could
-be comparable across the datasets. We defined MDR as:
+for each algorithm to give a quantitative comparison across the datasets. We defined MDR as:
 \begin{equation}

### 2021-10-19 test/test_earth_time_series.py
Add fit_single_tree parameter to filenames. #7

@@ -31,7 +31,7 @@ class TestEarthTimeSeries(TestCase):
 
         # Check results for each algorithm
         for file_base in ['demud-k=3/selections-demud.csv',
-                          'iforest-n_trees=100/selections-iforest.csv',
+                          'iforest-n_trees=100-fit_single_trees=False/selections-iforest.csv',
                           'negative_sampling-percent_increase=20/selections-negative_sampling.csv',
                           'pca-k=3/selections-pca.csv',
                           'rx/selections-rx.csv',

### 2021-10-19 test/test_earth_volcanoes_patches.py

@@ -31,7 +31,7 @@ class TestEarthVolcanoesPatches(TestCase):
 
         # Check results for each algorithm
         for file_base in ['demud-k=3/selections-demud.csv',
-                          'iforest-n_trees=100/selections-iforest.csv',
+                          'iforest-n_trees=100-fit_single_trees=False/selections-iforest.csv',
                           'lrx-inner_window=7-outer_window=11-bands=1/selections-lrx.csv',
                           'negative_sampling-percent_increase=20/selections-negative_sampling.csv',
                           'pca-k=3/selections-pca.csv',

### 2021-10-19 test/test_earth_volcanoes_pixels.py

@@ -31,7 +31,7 @@ class TestEarthVolcanoesPixels(TestCase):
 
         # Check results for each algorithm
         for file_base in ['demud-k=1/selections-demud.csv',
-                          'iforest-n_trees=100/selections-iforest.csv',
+                          'iforest-n_trees=100-fit_single_trees=False/selections-iforest.csv',
                           'negative_sampling-percent_increase=20/selections-negative_sampling.csv',
                           'pca-k=1/selections-pca.csv',
                           'rx/selections-rx.csv',

### 2021-10-19 test/test_planetary.py

@@ -31,7 +31,7 @@ class TestPlanetary(TestCase):
 
         # Check results for each algorithm
         for file_base in ['demud-k=5/selections-demud.csv',
-                          'iforest-n_trees=100/selections-iforest.csv',
+                          'iforest-n_trees=100-fit_single_trees=False/selections-iforest.csv',
                           'lrx-inner_window=3-outer_window=9-bands=1/selections-lrx.csv',
                           'negative_sampling-percent_increase=20/selections-negative_sampling.csv',
                           'pca-k=5/selections-pca.csv',

### 2021-10-19 test/test_earth_time_series.py
Fix lint errors. #14

@@ -13,6 +13,7 @@ import pytest
 from test_utils import check_results
 from dora_exp_pipeline.dora_exp import start
 
+
 @pytest.mark.functional
 class TestEarthTimeSeries(TestCase):
 
@@ -31,8 +32,10 @@ class TestEarthTimeSeries(TestCase):
 
         # Check results for each algorithm
         for file_base in ['demud-k=3/selections-demud.csv',
-                          'iforest-n_trees=100-fit_single_trees=False/selections-iforest.csv',
-                          'negative_sampling-percent_increase=20/selections-negative_sampling.csv',
+                          'iforest-n_trees=100-fit_single_trees=False/'
+                          'selections-iforest.csv',
+                          'negative_sampling-percent_increase=20/'
+                          'selections-negative_sampling.csv',
                           'pca-k=3/selections-pca.csv',
                           'rx/selections-rx.csv',
                           'random/selections-random.csv'

### 2021-10-19 test/test_earth_volcanoes_patches.py

@@ -13,6 +13,7 @@ import pytest
 from test_utils import check_results
 from dora_exp_pipeline.dora_exp import start
 
+
 @pytest.mark.functional
 class TestEarthVolcanoesPatches(TestCase):
 
@@ -31,14 +32,18 @@ class TestEarthVolcanoesPatches(TestCase):
 
         # Check results for each algorithm
         for file_base in ['demud-k=3/selections-demud.csv',
-                          'iforest-n_trees=100-fit_single_trees=False/selections-iforest.csv',
-                          'lrx-inner_window=7-outer_window=11-bands=1/selections-lrx.csv',
-                          'negative_sampling-percent_increase=20/selections-negative_sampling.csv',
+                          'iforest-n_trees=100-fit_single_trees=False/'
+                          'selections-iforest.csv',
+                          'lrx-inner_window=7-outer_window=11-bands=1/'
+                          'selections-lrx.csv',
+                          'negative_sampling-percent_increase=20/'
+                          'selections-negative_sampling.csv',
                           'pca-k=3/selections-pca.csv',
                           'rx/selections-rx.csv',
                           'random/selections-random.csv'
                           ]:
-            correct_file = 'test/ref-results/earth_volcanoes_patches/%s' % file_base
+            correct_file = 'test/ref-results/earth_volcanoes_patches' \
+                            '/%s' % file_base
             output_file = '%s/%s' % (self.outdir, file_base)
 
             # Check results against correct output

### 2021-10-19 test/test_earth_volcanoes_pixels.py

@@ -13,6 +13,7 @@ import pytest
 from test_utils import check_results
 from dora_exp_pipeline.dora_exp import start
 
+
 @pytest.mark.functional
 class TestEarthVolcanoesPixels(TestCase):
 
@@ -31,13 +32,16 @@ class TestEarthVolcanoesPixels(TestCase):
 
         # Check results for each algorithm
         for file_base in ['demud-k=1/selections-demud.csv',
-                          'iforest-n_trees=100-fit_single_trees=False/selections-iforest.csv',
-                          'negative_sampling-percent_increase=20/selections-negative_sampling.csv',
+                          'iforest-n_trees=100-fit_single_trees=False/'
+                          'selections-iforest.csv',
+                          'negative_sampling-percent_increase=20/'
+                          'selections-negative_sampling.csv',
                           'pca-k=1/selections-pca.csv',
                           'rx/selections-rx.csv',
                           'random/selections-random.csv'
                           ]:
-            correct_file = 'test/ref-results/earth_volcanoes_pixels/%s' % file_base
+            correct_file = 'test/ref-results/earth_volcanoes_pixels' \
+                            '/%s' % file_base
             output_file = '%s/%s' % (self.outdir, file_base)
 
             # Check results against correct output

### 2021-10-19 test/test_planetary.py

@@ -13,6 +13,7 @@ import pytest
 from test_utils import check_results
 from dora_exp_pipeline.dora_exp import start
 
+
 @pytest.mark.functional
 class TestPlanetary(TestCase):
 
@@ -31,13 +32,17 @@ class TestPlanetary(TestCase):
 
         # Check results for each algorithm
         for file_base in ['demud-k=5/selections-demud.csv',
-                          'iforest-n_trees=100-fit_single_trees=False/selections-iforest.csv',
-                          'lrx-inner_window=3-outer_window=9-bands=1/selections-lrx.csv',
-                          'negative_sampling-percent_increase=20/selections-negative_sampling.csv',
+                          'iforest-n_trees=100-fit_single_trees=False/'
+                          'selections-iforest.csv',
+                          'lrx-inner_window=3-outer_window=9-bands=1/'
+                          'selections-lrx.csv',
+                          'negative_sampling-percent_increase=20/'
+                          'selections-negative_sampling.csv',
                           'pca-k=5/selections-pca.csv',
                           'random/selections-random.csv'
                           ]:
-            correct_file = 'test/ref-results/planetary_rover/functional/%s' % file_base
+            correct_file = 'test/ref-results/planetary_rover/functional' \
+                            '/%s' % file_base
             output_file = '%s/%s' % (self.outdir, file_base)
 
             # Check results against correct output

### 2021-10-19 test/test_utils.py

@@ -1,15 +1,17 @@
 # Helper functions for pytests
 import numpy as np
 
+
 def read_selections(f):
 
     sels = []
 
     with open(f, 'r') as inf:
-        for l in inf:
-            if l[0] == '#':
+        for line in inf:
+            if line[0] == '#':
                 continue
-            sels += [l.split(',')[2].split('.')[0]] # name excluding extension
+            # Name excluding extension
+            sels += [line.split(',')[2].split('.')[0]]
 
     return sels
 
@@ -20,11 +22,11 @@ def read_selections_and_scores(f):
     scores = []
 
     with open(f, 'r') as inf:
-        for l in inf:
-            if l[0] == '#':
+        for line in inf:
+            if line[0] == '#':
                 continue
             # Get the image filename and numeric score
-            vals = l.split(',')[2:4]
+            vals = line.split(',')[2:4]
             sels += [vals[0].split('.')[0]]  # exclude extension
             scores += [float(vals[1])]  # convert score to float
 

### 2021-10-20 dora_vis/src/App.js
catalog dataloader datatable

@@ -6,8 +6,14 @@ const parse = window.require('csv-parse');
 const yaml = window.require('js-yaml');
 const fs = window.require('fs');
 const path = window.require('path');
+const hdf5 = window.require('jsfive');
 
-const DATALOADERS = ["image"];
+// All supported data loaders
+// image - imageset, data is directory of image files
+// catalog - h5 export of pandas dataframe
+const DATALOADERS = ["image", "catalog", "featurevector"];
+// Dataloaders that require a directory data root
+const DIRLOADERS = ["image"];
 
 
 class NavBar extends React.Component {
@@ -132,12 +138,12 @@ class ConfigParser extends React.Component {
     this.handleConfigSubmit = this.handleConfigSubmit.bind(this);
 
     this.state = {
-      dataLoader: this.props.configData["data_loader"]["name"],
+      dataLoader: this.props.configData["data_loader"]["name"].trim().toLowerCase(),
       dataRoot: "",
       dataRootPath: this.props.configData["data_to_score"],
       outDir: "",
       outDirPath: this.props.configData["out_dir"],
-      validDataLoader: DATALOADERS.includes(this.props.configData["data_loader"]["name"].trim()),
+      validDataLoader: DATALOADERS.includes(this.props.configData["data_loader"]["name"].trim().toLowerCase()),
       validDataRoot: false,
       validOutDir: false
     };
@@ -173,10 +179,17 @@ class ConfigParser extends React.Component {
 
   handleDataRootChange(e) {
     /* Handle data root input change */
-    this.setState({
-      dataRoot: e.target.value,
-      dataRootPath: path.dirname(e.target.files[0].path)
-    });
+    if (DIRLOADERS.includes(this.state["dataLoader"])) {
+      this.setState({
+        dataRoot: e.target.value,
+        dataRootPath: path.dirname(e.target.files[0].path)
+      });
+    } else {
+      this.setState({
+        dataRoot: e.target.value,
+        dataRootPath: e.target.files[0].path
+      });
+    }
   }
 
   handleDataRootSubmit(e) {
@@ -206,7 +219,8 @@ class ConfigParser extends React.Component {
     this.props.setConfig(
       this.state["dataLoader"],
       this.state["dataRootPath"],
-      this.state['outDirPath']);
+      this.state['outDirPath']
+    );
   }
   
   render() {
@@ -221,8 +235,9 @@ class ConfigParser extends React.Component {
     // Set data root output/input
     var componentDataRoot = null;
     if (this.state["validDataRoot"]) {
[…]

### 2021-10-20 dora_vis/src/Table.js

@@ -6,9 +6,25 @@ import { useTable, useSortBy, usePagination} from 'react-table'
 const fs = window.require('fs');
 const path = window.require('path');
 const {clipboard} = window.require('electron');
+const Plotly = window.require('plotly.js-dist');
 
+function percentRank(arr, v) {
+  // https://gist.github.com/IceCreamYou/6ffa1b18c4c8f6aeaad2
+  if (typeof v !== 'number') throw new TypeError('v must be a number');
+  for (var i = 0, l = arr.length; i < l; i++) {
+      if (v <= arr[i]) {
+          while (i < l && v === arr[i]) i++;
+          if (i === 0) return 0;
+          if (v !== arr[i-1]) {
+              i += (v - arr[i-1]) / (arr[i] - arr[i-1]);
+          }
+          return i / l;
+      }
+  }
+  return 1;
+}
 
-function Table({ columns, data }) {
+function Table({ columns, data, forceUpdate = () => ({}) }) {
   // Use the state and functions returned from useTable to build your UI
   const {
     getTableProps,
@@ -81,16 +97,16 @@ function Table({ columns, data }) {
         This is just a very basic UI implementation:
       */}
       <div className="pagination">
-        <button type="button" className="btn btn-outline-dark btn-sm" onClick={() => gotoPage(0)} disabled={!canPreviousPage}>
+        <button type="button" className="btn btn-outline-dark btn-sm" onClick={() => {gotoPage(0); forceUpdate();}} disabled={!canPreviousPage}>
           {'<<'}
         </button>{' '}
-        <button type="button" className="btn btn-outline-dark btn-sm" onClick={() => previousPage()} disabled={!canPreviousPage}>
+        <button type="button" className="btn btn-outline-dark btn-sm" onClick={() => {previousPage(); forceUpdate();}} disabled={!canPreviousPage}>
           {'<'}
         </button>{' '}
-        <button type="button" className="btn btn-outline-dark btn-sm" onClick={() => nextPage()} disabled={!canNextPage}>
+        <button type="button" className="btn btn-outline-dark btn-sm" onClick={() => {nextPage(); forceUpdate();}} disabled={!canNextPage}>
           {'>'}
         </button>{' '}
-        <button type="button" className="btn btn-outline-dark btn-sm" onClick={() => gotoPage(pageCount - 1)} disabled={!canNextPage}>
[…]

### 2021-10-20 dora_vis/src/App.js
catalog -> featurevector

@@ -10,8 +10,8 @@ const hdf5 = window.require('jsfive');
 
 // All supported data loaders
 // image - imageset, data is directory of image files
-// catalog - h5 export of pandas dataframe
-const DATALOADERS = ["image", "catalog", "featurevector"];
+// featurevector - h5 export of pandas dataframe
+const DATALOADERS = ["image", "featurevector"];
 // Dataloaders that require a directory data root
 const DIRLOADERS = ["image"];
 
@@ -386,7 +386,7 @@ class App extends React.Component {
           });
           this.setState({data: dataObj});
         });
-    } else if (this.state["dataLoader"] === "catalog" || this.state["dataLoader"] === "featurevector") {
+    } else if (this.state["dataLoader"] === "featurevector") {
       // jsfive
       var rawData = fs.readFileSync(this.state["dataRoot"]);
       var h5File = new hdf5.File(rawData.buffer);

### 2021-10-20 dora_vis/src/Table.js

@@ -222,8 +222,8 @@ class DataTable extends React.Component {
   }
 
   componentDidUpdate() {
-    // Catalog dataloader requires plotly plots
-    if (this.props.dataLoader === "catalog" || this.props.dataLoader === "featurevector") {
+    // Feature vector dataloader requires plotly plots
+    if (this.props.dataLoader === "featurevector") {
       // get length of data and dimension of each feature
       var dataLength = this.props.dataArray != null ? this.props.dataArray.length : 0;
       var featLength = this.props.featNames != null ? this.props.featNames.length : 1;
@@ -320,7 +320,7 @@ class DataTable extends React.Component {
           sortType: "number"
         }
       ];
-    } else if (this.props.dataLoader === "catalog" || this.props.dataLoader == "featurevector") {
+    } else if (this.props.dataLoader == "featurevector") {
       columns = [
         {
           Header: "Rank",

### 2021-10-21 dora_vis/src/App.js
Improved data rep, better featvec plot, better err

@@ -203,7 +203,7 @@ class ConfigParser extends React.Component {
     /* Handle output directory input change */
     this.setState({
       outDir: e.target.value,
-      outDirPath: path.dirname(e.target.files[0].path)
+      outDirPath: e.target.files != null ? path.dirname(e.target.files[0].path) : null
     });
   }
 
@@ -367,7 +367,6 @@ class App extends React.Component {
     const methodCSVPath = path.join(this.state["outDir"], methodDirName, methodCSVName);
 
     let data = [];
-
     if (this.state["dataLoader"] === "image") {
       fs.createReadStream(methodCSVPath)
         .pipe(parse({delimiter: ', '}))
@@ -390,7 +389,6 @@ class App extends React.Component {
       // jsfive
       var rawData = fs.readFileSync(this.state["dataRoot"]);
       var h5File = new hdf5.File(rawData.buffer);
-      var featDim = h5File.get("data/axis0").value.length;
 
       fs.createReadStream(methodCSVPath)
         .pipe(parse({delimiter: ', '}))
@@ -399,31 +397,39 @@ class App extends React.Component {
         })
         .on('end', () => {
           const dataObj = data.map(row => {
+            let elementId = parseInt(row[1]);
+            let flatArray = h5File.get("data/block0_values").value != null ? h5File.get("data/block0_values").value : [];
+            let featDim = h5File.get("data/axis0").value != null ? h5File.get("data/axis0").value.length : 0;
+            
             return {
               rank: parseInt(row[0]),
-              id: parseInt(row[1]),
+              id: elementId,
               fileName: row[2],
+              feature: flatArray.slice(elementId * featDim, (elementId + 1) * featDim),
               score: parseFloat(row[3])
             };
           });
           this.setState({
             featNames: h5File.get("data/axis0").value,
-            data: dataObj,
-            dataArray: h5File.get("data/block0_values").value
+            data: dataObj.sort((a,b) => a.rank < b.rank ? -1 : 1)
           });
         });
     }
   }
 
   loadAggData() {
+    if (this.state["dataLoader"] !== "image") {
+      this.setState({aggData: null});
+      return;
+    }
+
     // Get list of available methods
-    const methods = Object.keys(this.state["configData"]["outlier_detection"]);
-    let data = [];
+    const methods = this.state["configData"] != null ? Object.keys(this.state["configData"]["outlier_detection"]) : [];
+    let dataList = [];
 
     // For each method available
     for (let methodName of methods) {
       const currData = [];
-
       // Define and find CSV
[…]

### 2021-10-21 dora_vis/src/Table.js

@@ -161,27 +161,39 @@ class AggTable extends React.Component {
       headerClassName: "rankHeader",
       sortType: "number"
     }]
-
-    for (let methodName of Object.keys(this.props.configData["outlier_detection"])) {
-      columns.push({
-        Header: methodName,
-        className: "imageCell",
-        headerClassName: "imageHeader",
-        Cell: (row) => {
-          return(
-          <div>
-            <a className="imageLink" onClick={() => {clipboard.writeText(row.row.original[methodName+"Name"]);alert(row.row.original[methodName+"Name"]+ " copied to clipboard.");}}>
-              <img src={"data:image/png;base64,"+row.row.original[methodName]} title={row.row.original[methodName+"Name"]}/>
-            </a>
-          </div>
-          );
-        }
-      });
+    if (this.props.configData != null) {
+      for (let methodName of Object.keys(this.props.configData["outlier_detection"])) {
+        columns.push({
+          Header: methodName,
+          className: "imageCell",
+          headerClassName: "imageHeader",
+          Cell: (row) => {
+            return(
+            <div>
+              <a className="imageLink" onClick={() => {clipboard.writeText(row.row.original[methodName+"Name"]);alert(row.row.original[methodName+"Name"]+ " copied to clipboard.");}}>
+                <img src={"data:image/png;base64,"+row.row.original[methodName]} title={row.row.original[methodName+"Name"]}/>
+              </a>
+            </div>
+            );
+          }
+        });
+      }
     }
 
     let datapass = null;
-    if (this.props.data == null) {
-      datapass = <h1> No DORA configuration loaded </h1>;
+    if (this.props.data === null) {
+      datapass =
+      <>
+        <h1> Loading Aggregate Table... </h1>
+        <br/>
+        <p> If stuck, it may be due to: </p>
+        <ul>
+          <li><b>A dataloader that is not "image"</b></li>
+          <li>No configuration being loaded</li>
+          <li>Mismatch between the configuration, data, and results</li>
+          <li>A very large data set</li>
+        </ul>
+      </>;
     } else {
       datapass = <>
         <h1>Aggregate Results</h1>
@@ -224,40 +236,71 @@ class DataTable extends React.Component {
   componentDidUpdate() {
     // Feature vector dataloader requires plotly plots
     if (this.props.dataLoader === "featurevector") {
-      // get length of data and dimension of each feature
-      var dataLength = this.props.dataArray != null ? this.props.dataArray.length : 0;
[…]

### 2021-10-21 dora_vis/src/App.js
Support for time series

@@ -11,7 +11,7 @@ const hdf5 = window.require('jsfive');
 // All supported data loaders
 // image - imageset, data is directory of image files
 // featurevector - h5 export of pandas dataframe
-const DATALOADERS = ["image", "featurevector"];
+const DATALOADERS = ["image", "featurevector", "time series"];
 // Dataloaders that require a directory data root
 const DIRLOADERS = ["image"];
 
@@ -365,9 +365,9 @@ class App extends React.Component {
     const methodDirName = methodOptions.join('-');
     const methodCSVName = "selections-" + methodName + ".csv";
     const methodCSVPath = path.join(this.state["outDir"], methodDirName, methodCSVName);
-
-    let data = [];
+    console.log(methodCSVPath);
     if (this.state["dataLoader"] === "image") {
+      let data = [];
       fs.createReadStream(methodCSVPath)
         .pipe(parse({delimiter: ', '}))
         .on('data', csvrow => {
@@ -386,6 +386,7 @@ class App extends React.Component {
           this.setState({data: dataObj});
         });
     } else if (this.state["dataLoader"] === "featurevector") {
+      let data = [];
       // jsfive
       var rawData = fs.readFileSync(this.state["dataRoot"]);
       var h5File = new hdf5.File(rawData.buffer);
@@ -414,6 +415,39 @@ class App extends React.Component {
             data: dataObj.sort((a,b) => a.rank < b.rank ? -1 : 1)
           });
         });
+    } else if (this.state["dataLoader"] === "time series") {
+      let dataArray = [];
+      let data = [];
+
+      fs.createReadStream(this.state["dataRoot"])
+        .pipe(parse({delimiter: ','}))
+        .on('data', datarow => {
+          dataArray.push(datarow.slice(1));
+        })
+        .on('end', () => {
+          fs.createReadStream(methodCSVPath)
+            .pipe(parse({delimiter: ', '}))
+            .on('data', csvrow => {
+              data.push(csvrow);
+            })
+            .on('end', () => {
+              const dataObj = data.map(row => {
+                return {
+                  rank: parseInt(row[0]),
+                  id: parseInt(row[1]),
+                  fileName: row[2],
+                  feature: dataArray[parseInt(row[1])].map(num => Number(num)),
+                  score: parseFloat(row[3])
+                };
+              });
+              console.log(dataArray);
+              this.setState({
+                featNames: [...Array(dataArray[0].length).keys()],
+                data: dataObj
+              });
+            });
+        });
+
     }
   }
 

### 2021-10-21 dora_vis/src/Table.js

@@ -235,7 +235,7 @@ class DataTable extends React.Component {
 
   componentDidUpdate() {
     // Feature vector dataloader requires plotly plots
-    if (this.props.dataLoader === "featurevector") {
+    if (this.props.dataLoader === "featurevector" || this.props.dataLoader === "time series") {
 
       // get dimension of each feature
       var featLength = this.props.featNames != null ? this.props.featNames.length : 0;
@@ -363,7 +363,7 @@ class DataTable extends React.Component {
           sortType: "number"
         }
       ];
-    } else if (this.props.dataLoader === "featurevector") {
+    } else if (this.props.dataLoader === "featurevector" || this.props.dataLoader === "time series") {
       columns = [
         {
           Header: "Rank",

### 2021-10-21 dora_vis/src/App.js
Minor javascript edits

@@ -464,6 +464,7 @@ class App extends React.Component {
     // For each method available
     for (let methodName of methods) {
       const currData = [];
+      var currRow = {};
       // Define and find CSV
       const methodOptions = [methodName];
       if (this.state["configData"]["outlier_detection"][methodName] != null) {
@@ -484,7 +485,7 @@ class App extends React.Component {
         .on('end', () => {
           if (dataList.length === 0) {
             for (let row of currData) {
-              var currRow = {};
+              currRow = {};
               currRow["rank"] = parseInt(row[0]);
               currRow[methodName] = fs.readFileSync(path.join(this.state["dataRoot"], row[2])).toString('base64');
               currRow[methodName+"Name"] = row[2];
@@ -492,7 +493,7 @@ class App extends React.Component {
             }
           } else {
             for (let [ind, row] of currData.entries()) {
-              var currRow = {};
+              currRow = {};
               currRow["rank"] = parseInt(row[0]);
               currRow[methodName] = fs.readFileSync(path.join(this.state["dataRoot"], row[2])).toString('base64');
               currRow[methodName+"Name"] = row[2];

### 2021-10-21 dora_vis/src/Table.js

@@ -3,8 +3,6 @@ import './Table.css';
 
 import { useTable, useSortBy, usePagination} from 'react-table'
 
-const fs = window.require('fs');
-const path = window.require('path');
 const {clipboard} = window.require('electron');
 const Plotly = window.require('plotly.js-dist');
 

### 2021-10-21 dora_vis/src/App.js
inline comments

@@ -8,22 +8,27 @@ const fs = window.require('fs');
 const path = window.require('path');
 const hdf5 = window.require('jsfive');
 
-// All supported data loaders
-// image - imageset, data is directory of image files
-// featurevector - h5 export of pandas dataframe
+/* Define Supported Data Loaders
+image: data is directory of image files
+featurevector: data is hdf5 export of pandas dataframe
+time series: data is a csv file, first column is dropped
+*/
 const DATALOADERS = ["image", "featurevector", "time series"];
-// Dataloaders that require a directory data root
+// These dataloaders require a directory path for data
 const DIRLOADERS = ["image"];
 
 
 class NavBar extends React.Component {
+  /* Navigation bar component */
   constructor(props) {
     super(props);
 
+    // Handle changing views
     this.handleNavClick = this.handleNavClick.bind(this);
   }
 
   handleNavClick(e) {
+    /* Handle changing views */
     e.preventDefault();
     switch(e.target.innerText) {
       case "Configure":
@@ -41,6 +46,7 @@ class NavBar extends React.Component {
   }
 
   render() {
+    // Logic to bold current view
     let configureActive = "";
     let aggregateActive = "";
     let tableActive = "";
@@ -56,6 +62,8 @@ class NavBar extends React.Component {
     } else {
       configureActive = aggregateActive = tableActive = "";
     }
+
+    // NavBar JSX
     return(
       <nav className="navbar navbar-expand-lg navbar-light bg-light">
         <div className="container-fluid">
@@ -83,13 +91,22 @@ class NavBar extends React.Component {
 }
 
 class ConfigLoader extends React.Component {
+  /* Initial configuration loader component */
   constructor(props) {
     super(props);
-    this.handleConfigPathChange = this.handleConfigPathChange.bind(this);
-    this.handleConfigPathSubmit = this.handleConfigPathSubmit.bind(this);
 
+    // Handle config file input field change
+    this.handleConfigPathChange = this.handleConfigPathChange.bind(this);
+    // Handle submission of file input field
+    this.handleConfigPathSubmit = this.handleConfigPathSubmit.bind(this);
+    
+    /* Component states
+    configFile:   Filename of config
+    configPath:   Filepath of config
+    */
     this.state = {configFile: "",
-                  configPath: ""};
+                  configPath: ""
+    };
   }
 
   handleConfigPathChange(e) {
@@ -108,6 +125,7 @@ class ConfigLoader extends React.Component {
   }
 
   render() {
+    /* Configuration loader JSX */
     return(
[…]

### 2021-10-21 dora_vis/src/Table.js

@@ -7,7 +7,8 @@ const {clipboard} = window.require('electron');
 const Plotly = window.require('plotly.js-dist');
 
 function percentRank(arr, v) {
-  // https://gist.github.com/IceCreamYou/6ffa1b18c4c8f6aeaad2
+  /* Get the percentile of a value given a sorted array
+     From https://gist.github.com/IceCreamYou/6ffa1b18c4c8f6aeaad2 */
   if (typeof v !== 'number') throw new TypeError('v must be a number');
   for (var i = 0, l = arr.length; i < l; i++) {
       if (v <= arr[i]) {
@@ -23,16 +24,15 @@ function percentRank(arr, v) {
 }
 
 function Table({ columns, data, forceUpdate = () => ({}) }) {
-  // Use the state and functions returned from useTable to build your UI
+  /* Table for showing results from a single method
+     ReactTable template, taken from the documentation example */
   const {
     getTableProps,
     getTableBodyProps,
     headerGroups,
     prepareRow,
-    page, // Instead of using 'rows', we'll use page,
-    // which has only the rows for the active page
+    page,
 
-    // The rest of these things are super handy, too ;)
     canPreviousPage,
     canNextPage,
     pageOptions,
@@ -52,7 +52,7 @@ function Table({ columns, data, forceUpdate = () => ({}) }) {
     usePagination
   )
 
-  // Render the UI for your table
+  // Render the UI for the table
   return (
     <>
       <table className="table table-striped" {...getTableProps()}>
@@ -62,7 +62,6 @@ function Table({ columns, data, forceUpdate = () => ({}) }) {
               {headerGroup.headers.map(column => (
                 <th {...column.getHeaderProps({...column.getSortByToggleProps(),...{className: column.headerClassName}})}>
                 {column.render('Header')}
-                {/* Add a sort direction indicator */}
                 <span>
                   {column.canSort
                     ? column.isSorted
@@ -90,10 +89,6 @@ function Table({ columns, data, forceUpdate = () => ({}) }) {
           })}
         </tbody>
       </table>
-      {/* 
-        Pagination can be built however you'd like. 
-        This is just a very basic UI implementation:
-      */}
       <div className="pagination">
         <button type="button" className="btn btn-outline-dark btn-sm" onClick={() => {gotoPage(0); forceUpdate();}} disabled={!canPreviousPage}>
           {'<<'}
@@ -143,15 +138,18 @@ function Table({ columns, data, forceUpdate = () => ({}) }) {
 }
[…]

### 2021-12-06 publications/iaai21/dora_outlier_detection_iaai.tex
Fix typo in manuscript.

@@
  benchmark dataset that uses MNIST and Fashion-MNIST. Table \ref{tab:datasets}
- summarizes the number of unlabeled samples used for training evaluation for
+ summarizes the number of unlabeled samples used for training and evaluation for
  each dataset. We describe each dataset in detail below.
@@
 and LRX was omitted as it applies only to image data.  Of the remaining methods, 
-PCA achieved the highest precision, following by RX.  Negative sampling 
+PCA achieved the highest precision, followed by RX.  Negative sampling 
 performs well initially before its performance drops off.  The PAE finds the most outliers 

### 2021-12-06 dora_exp_pipeline/dora_data_loader.py
Add csv to list of supported file types for FeatureVectorLoader.

@@ -246,7 +246,7 @@ class FeatureVectorLoader(DataLoader):
                                f'{os.path.abspath(dir_path)}')
 
         # List of supported file types
-        file_types = ['.h5']
+        file_types = ['.h5', '.csv']
 
         data_dict = dict()
         data_dict.setdefault('id', [])

### 2021-12-09 dora_exp_pipeline/dora_results_organization.py
Rename from rastermdr to raster metrics:

@@ -145,9 +145,9 @@ save_comparison_plot = SaveComparisonPlot()
 register_org_method(save_comparison_plot)
 
 
-class RasterMDR(ResultsOrganization):
+class RasterMetrics(ResultsOrganization):
     def __init__(self):
-        super(RasterMDR, self).__init__('raster_mdr')
+        super(RasterMetrics, self).__init__('raster_metrics')
 
     def _run(self, data_ids, dts_scores, dts_sels, data_to_score,
              outlier_alg_name, out_dir, logger, seed, top_n, raster_path,
@@ -183,11 +183,7 @@ class RasterMDR(ResultsOrganization):
         # Only keep pixels where we had scores
         labels = labels[np.isfinite(max_scores)]
         max_scores = max_scores[np.isfinite(max_scores)]
-
-        # Sort scores and labels together, by score
-        sorted_scores, sorted_labels = zip(*sorted(zip(max_scores.flatten(),
-                                                       labels.flatten()),
-                                                   reverse=True))
+        sorted_labels = labels[max_scores.argsort()[::-1]]
 
         # Calculate MDR
         y = [0]  # Correct selections at each iteration
@@ -203,19 +199,18 @@ class RasterMDR(ResultsOrganization):
 
         # Calculate other metrics
         precision_at_n = y[-1]/x[-1]
-        max_recall = average_precision_score(labels.flatten(),
-                                             max_scores.flatten())
+        ap_score = average_precision_score(labels, max_scores)
 
         # Save results to txt file
         fname = str(datetime.datetime.now())
         with open(os.path.join(out_dir, f'{fname}.txt'), 'w') as f:
             res = " ".join([outlier_alg_name, shp_path, str(patch_size),
-                            str(mdr), str(max_recall), str(precision_at_n)])
+                            str(mdr), str(ap_score), str(precision_at_n)])
             f.write(res)
 
 
-raster_mdr = RasterMDR()
-register_org_method(raster_mdr)
+raster_metrics = RasterMetrics()
+register_org_method(raster_metrics)
 
 
 class KmeansCluster(ResultsOrganization):

### 2022-01-05 publications/frontiers/dora_outlier_detection_frontiers.tex
Convert manuscript from IAAI to Frontiers format. #79

--- 2021-12-06 publications/iaai21/dora_outlier_detection_iaai.tex
@@
-\def\year{2021}\relax
-%File: formatting-instructions-latex-2021.tex
-%release 2021.2
-\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
-\usepackage{aaai21}  % DO NOT CHANGE THIS
-\usepackage{times}  % DO NOT CHANGE THIS
-\usepackage{helvet} % DO NOT CHANGE THIS
-\usepackage{courier}  % DO NOT CHANGE THIS
-\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
-\usepackage{graphicx} % DO NOT CHANGE THIS
-\urlstyle{rm} % DO NOT CHANGE THIS
-\def\UrlFont{\rm}  % DO NOT CHANGE THIS
-\usepackage{natbib}  % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
-\usepackage{caption} % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
-\frenchspacing  % DO NOT CHANGE THIS
-\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
-\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS
-%\nocopyright
-%PDF Info Is REQUIRED.
-% For /Author, add all authors within the parentheses, separated by commas. No accents or commands.
-% For /Title, add Title in Mixed Case. No accents or commands. Retain the parentheses.
-\pdfinfo{
-/Title (Domain-agnostic Outlier Ranking Algorithms (DORA)---A Configurable Pipeline for Facilitating Outlier Detection in Scientific Datasets)
-/Author (AAAI Press Staff, Pater Patel Schneider, Sunil Issar, J. Scott Penberthy, George Ferguson, Hans Guesgen, Francisco Cruz, Marc Pujol-Gonzalez)
-/TemplateVersion (2021.2)
-} %Leave this
-% /Title ()
-% Put your actual complete title (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case
-% Leave the space between \Title and the beginning parenthesis alone
-% /Author ()
-% Put your actual complete list of authors (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case.
-% Each author should be only by a comma. If the name contains accents, remove them. If there are any LaTeX commands,
-% remove them.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% This is just an example/guide for you to refer to when submitting manuscripts to Frontiers, it is not mandatory to use Frontiers .cls files nor frontiers.tex  %
+% This will only generate the Manuscript, the final article will be typeset by Frontiers after acceptance.   
+%                                              %
+%                                                                                                                                                         %
+% When submitting your files, remember to upload this *tex file, the pdf generated with it, the *bib file (if bibliography is not within the *tex) and all the figures.
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
-% DISALLOWED PACKAGES
-% \usepackage{authblk} -- This package is specifically forbidden
-% \usepackage{balance} -- This package is specifically forbidden
-% \usepackage{color (if used in text)
-% \usepackage{CJK} -- This package is specifically forbidden
-% \usepackage{float} -- This package is specifically forbidden
-% \usepackage{flushend} -- This package is specifically forbidden
-% \usepackage{fontenc} -- This package is specifically forbidden
-% \usepackage{fullpage} -- This package is specifically forbidden
-% \usepackage{geometry} -- This package is specifically forbidden
-% \usepackage{grffile} -- This package is specifically forbidden
-% \usepackage{hyperref} -- This package is specifically forbidden
-% \usepackage{navigator} -- This package is specifically forbidden
-% (or any other package that embeds links such as navigator or hyperref)
-% \indentfirst} -- This package is specifically forbidden
-% \layout} -- This package is specifically forbidden
-% \multicol} -- This package is specifically forbidden
-% \nameref} -- This package is specifically forbidden
-% \usepackage{savetrees} -- This package is specifically forbidden
-% \usepackage{setspace} -- This package is specifically forbidden
-% \usepackage{stfloats} -- This package is specifically forbidden
-% \usepackage{tabu} -- This package is specifically forbidden
-% \usepackage{titlesec} -- This package is specifically forbidden
-% \usepackage{tocbibind} -- This package is specifically forbidden
-% \usepackage{ulem} -- This package is specifically forbidden
-% \usepackage{wrapfig} -- This package is specifically forbidden
-% DISALLOWED COMMANDS
-% \nocopyright -- Your paper will not be published if you use this command
-% \addtolength -- This command may not be used
-% \balance -- This command may not be used
-% \baselinestretch -- Your paper will not be published if you use this command
-% \clearpage -- No page breaks of any kind may be used for the final version of your paper
-% \columnsep -- This command may not be used
-% \newpage -- No page breaks of any kind may be used for the final version of your paper
-% \pagebreak -- No page breaks of any kind may be used for the final version of your paperr
-% \pagestyle -- This command may not be used
-% \tiny -- This is not an acceptable font size.
-% \vspace{- -- No negative value may be used in proximity of a caption, figure, table, section, subsection, subsubsection, or reference
-% \vskip{- -- No negative value may be used to alter spacing above or below a caption, figure, table, section, subsection, subsubsection, or reference
+%%% Version 3.4 Generated 2018/06/15 %%%
+%%% You will need to have the following packages installed: datetime, fmtcount, etoolbox, fcprefix, which are normally inlcuded in WinEdt. %%%
+%%% In http://www.ctan.org/ you can find the packages and how to install them, if necessary. %%%
+%%%  NB logo1.jpg is required in the path in order to correctly compile front page header %%%
 
-\setcounter{secnumdepth}{0} %May be changed to 1 or 2 if section numbers are desired.
+%\documentclass[utf8]{frontiersSCNS} % for Science, Engineering and Humanities and Social Sciences articles
+%\documentclass[utf8]{frontiersHLTH} % for Health articles
+\documentclass[utf8]{frontiersFPHY} % for Physics and Applied Mathematics and Statistics articles
 
+\setcitestyle{square} % for Physics and Applied Mathematics and Statistics articles
+\usepackage{url,hyperref,lineno,microtype,subcaption}
+\usepackage[onehalfspacing]{setspace}
+
+% packages added by authors
+\usepackage{graphicx}
 \usepackage{siunitx}
@@
 
-% The file aaai21.sty is the style file for AAAI Press
-% proceedings, working notes, and technical reports.
-%
+\linenumbers
 
-% Title
+% Leave a blank line between paragraphs instead of using \\
 
-% Your title must be in mixed case, not sentence case.
-% That means all verbs (including short verbs like be, is, using,and go),
-% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
-% articles, conjunctions, and prepositions are lower case unless they
-% directly follow a colon or long dash
-\iffalse
-\title{Domain-agnostic Outlier Ranking Algorithms (DORA)---A Configurable Pipeline for Facilitating Outlier Detection in Scientific Datasets}
-\author{
-    %Authors
-    % All authors must be in the same font size and format.
-    Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{With help from the AAAI Publications Committee.}\\
-    AAAI Style Contributions by Pater Patel Schneider,
-    Sunil Issar,  \\
-    J. Scott Penberthy,
-    George Ferguson,
-    Hans Guesgen,
-    Francisco Cruz,
-    Marc Pujol-Gonzalez
-    \\
-}
-\affiliations{
-    %Afiliations
-    \textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\
-    %If you have multiple authors and multiple affiliations
-    % use superscripts in text and roman font to identify them.
-    %For example,
+\def\keyFont{\fontsize{8}{11}\helveticabold }
+\def\firstAuthorLast{Kerner {et~al.}} %use et al only if is more than 1 author
+\def\Authors{Hannah R. Kerner\,$^{1,*}$, Umaa Rebbapragada\,$^{2}$, 
+Kiri L. Wagstaff\,$^{2}$, Steven Lu\,$^{2}$, Bryce Dubayah\,$^{1}$,
+Eric Huff\,$^{2}$, Jake Lee\,$^{2}$, Vinay Raman\,$^{3}$, and
+Sakshum Kulshrestha\,$^{1}$}
+% Affiliations should be keyed to the author's name with superscript numbers and be listed as follows: Laboratory, Institute, Department, Organization, City, State abbreviation (USA, Canada, Australia), and Country (without detailed address information such as city zip codes or street names).
+% If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
+\def\Address{$^{1}$Department of Geographical Sciences, University of Maryland, College Park, MD, USA \\
+$^{2}$Jet Propulsion Laboratory, California Institute of Technology, Pasadena, California, USA \\
+$^{3}$Montgomery Blair High School, Silver Spring, MD, USA}
+% The Corresponding Author should be marked with an asterisk
+% Provide the exact contact address (this time including street name and city zip code) and email of the corresponding author
+\def\corrAuthor{Corresponding Author}
 
-    % Sunil Issar, \textsuperscript{\rm 2}
-    % J. Scott Penberthy, \textsuperscript{\rm 3}
-    % George Ferguson,\textsuperscript{\rm 4}
-    % Hans Guesgen, \textsuperscript{\rm 5}.
-    % Note that the comma should be placed BEFORE the superscript for optimum readability
-
-    2275 East Bayshore Road, Suite 160\\
-    Palo Alto, California 94303\\
-    % email address must be in roman text type, not monospace or sans serif
-    publications21@aaai.org
-
-    % See more examples next
-}
-\fi
-\iffalse
-%Example, Single Author, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it
-\title{My Publication Title --- Single Author}
-\author {
-    % Author
-    Author Name \\
-}
-
-\affiliations{
-    Affiliation \\
-    Affiliation Line 2 \\
-    name@example.com
-}
-\fi
-
-%Example, Multiple Authors, ->> remove \iffalse,\fi and place them surrounding AAAI title to use it
-\title{Domain-agnostic Outlier Ranking Algorithms (DORA)---A Configurable Pipeline for Facilitating Outlier Detection in Scientific Datasets}
-\author {
-    % Authors
-    Hannah Kerner,\textsuperscript{\rm 1}
-    Umaa Rebbapragada, \textsuperscript{\rm 2}
-    Kiri Wagstaff, \textsuperscript{\rm 2} 
-    Steven Lu, \textsuperscript{\rm 2} 
-    Bryce Dubayah, \textsuperscript{\rm 1} \\
-    Eric Huff, \textsuperscript{\rm 2} 
-    Jake Lee, \textsuperscript{\rm 2} 
-    Vinay Raman, \textsuperscript{\rm 3}
-    Sakshum Kulshrestha \textsuperscript{\rm 1}  \\
-}
-\affiliations {
-    % Affiliations
-    \textsuperscript{\rm 1} University of Maryland, College Park \\
-    \textsuperscript{\rm 2} Jet Propulsion Laboratory, 
-    									   California Institute of Technology \\
-    \textsuperscript{\rm 3} Montgomery Blair High School \\
-    hkerner@umd.edu, \{umaa.d.rebbapragada,
-    										kiri.l.wagstaff,
-    										you.lu,
-    										eric.m.huff,
-    										jake.h.lee\}@jpl.nasa.gov
-}
+\def\corrEmail{hkerner@umd.edu}
 
 \begin{document}
+\onecolumn
+\firstpage{1}
+
+\title[Domain-agnostic Outlier Ranking Algorithms (DORA)]{Domain-agnostic Outlier Ranking Algorithms (DORA)---A Configurable Pipeline for Facilitating Outlier Detection in Scientific Datasets}
+
+\author[\firstAuthorLast ]{\Authors} %This field will be automatically populated
+\address{} %This field will be automatically populated
+\correspondance{} %This field will be automatically populated
+
+\extraAuth{}% If there are more than 1 corresponding author, comment this line and uncomment the next one.
+%\extraAuth{corresponding Author2 \\ Laboratory X2, Institute X2, Department X2, Organization X2, Street X2, City X2 , State XX2 (only USA, Canada and Australia), Zip Code2, X2 Country X2, email2@uni2.edu}
 
@@
 \begin{abstract}
-Automatic detection of outliers is
-universally needed when working with scientific datasets, e.g., for cleaning 
-datasets or flagging novel samples to guide instrument acquisition or
- scientific analysis.
+
+Automatic detection of outliers is universally needed when working with 
+scientific datasets, e.g., for cleaning datasets or flagging novel samples to 
+guide instrument acquisition or scientific analysis.
 We present Domain-agnostic Outlier Ranking Algorithms (DORA),
@@
 or feature vector input data types and outlier detection methods that include 
- Isolation Forest, DEMUD, PCA, RX detector, 
- Local RX, negative sampling, and probabilistic autoencoder. Each algorithm assigns an outlier score to each data sample. DORA
-   provides results interpretation modules to help users 
+Isolation Forest, DEMUD, PCA, RX detector, Local RX, negative sampling, 
+and probabilistic autoencoder. Each algorithm assigns an outlier score to 
+each data sample. DORA provides results interpretation modules to help users 
 process the results, including sorting samples by outlier score, evaluating 
@@
 facilitates application, evaluation, and interpretation of outlier detection
-   methods by performing
-    experiments for three real-world datasets from Earth science, planetary
-  science, and astrophysics, as well as one benchmark dataset 
-  (MNIST/Fashion-MNIST). We found that no single algorithm
-      performed best across all datasets, underscoring the need for a tool 
-      that enables comparison of multiple algorithms.
+methods by performing experiments for three real-world datasets from Earth 
+science, planetary science, and astrophysics, as well as one benchmark dataset 
+(MNIST/Fashion-MNIST). We found that no single algorithm performed best across 
+all datasets, underscoring the need for a tool that enables comparison of 
+multiple algorithms.
+
+\tiny
+ \keyFont{ \section{Keywords:} astrophysics, planetary science, Earth science, outlier detection, novelty detection, out-of-distribution detection, anomaly detection} %All article types: you may provide up to 8 keywords; at least 5 are mandatory.
 \end{abstract}
@@
 \section{Introduction}
+
 The ability to automatically detect out-of-distribution samples in large data 
@@
 In machine learning, these methods are commonly used for identifying 
-   mislabeled or otherwise invalid samples in a dataset. When working with science datasets, OOD detection can be used for 
+mislabeled or otherwise invalid samples in a dataset. When working with 
+science datasets, OOD detection can be used for 
 cleaning datasets, e.g., flagging ground-truth labels with GPS or human
@@
 \section{Related Work}
+
 Methods for outlier detection have been 
@@
 
-\paragraph{Data types.} We chose to implement data loaders for four
+\subsection{Data loaders} 
+We chose to implement data loaders for four
 data types that are commonly used by the machine learning and domain science
@@
 defined in the configuration file and loads samples into a dictionary of numpy
- arrays indexed by the sample id. This \texttt{data\_dict} is then passed to each of the ranking algorithms.
+arrays indexed by the sample id. This \texttt{data\_dict} is then passed to 
+each of the ranking algorithms.
 
-\paragraph{Outlier ranking algorithms.} We implemented 7 unsupervised 
+\subsection{Outlier ranking algorithms} 
+We implemented 7 unsupervised 
 algorithms for scoring and 
@@
 perform better for different datasets and use cases. 
-%We describe each
-%approach to scoring outliers and the associated methods below.
+We describe each
+approach to scoring outliers and the associated methods below.
 
@@
 
-\subparagraph{Reconstruction error.}
+\subsubsection{Reconstruction error}
 Principal component analysis (PCA) has been used for outlier detection by
@@
 
-\subparagraph{Distance.}
+\subsubsection{Distance}
 The 
@@
 
-\subparagraph{Sparsity.}
+\subsubsection{Sparsity}
 Isolation forests are a common sparsity-based method
@@
 
-\subparagraph{Probability.}
+\subsubsection{Probability}
 The negative sampling algorithm is implemented by converting the unsupervised 
@@
 
-\paragraph{Results interpretation.}
+\subsection{Results interpretation}
 Each of the outlier ranking algorithms returns an array containing the sample
@@
 This enables fast and easy comparison of the results from
+different methods. 
+%Figure \ref{fig:doravis} shows a screenshot of the
+%``Aggregate Table'' view, which displays all algorithm results 
+%side-by-side.
+%
+We developed a desktop application to easily visualize DORA results with the
+Electron application framework and React frontend library. 
+The application loads
+the DORA configuration file to locate the dataset and result CSVs. Then,
+it displays the ranked samples and their scores in a table sorted by their
+selection order. This allows for fast and easy comparison of the results of
 different methods. Figure \ref{fig:doravis} shows a screenshot of the
-``Aggregate Table'' view, which displays all algorithm results 
+``Aggregate Table'' view, which displays all results from different algorithms
 side-by-side.
-
-%We developed a desktop application to easily visualize DORA results with the
-%Electron application framework and React frontend library. The application loads
-%the DORA configuration file to locate the dataset and result CSVs. Then,
-%it displays the ranked samples and their scores in a table sorted by their
-%selection order. This allows for fast and easy comparison of the results of
-%different methods. Figure \ref{fig:doravis} shows a screenshot of the
-%``Aggregate Table'' view, which displays all results from different algorithms
-%side-by-side.
 
@@
 
-\subsubsection{Astrophysics: objects in Dark Energy Survey.}
+\subsection{Astrophysics: objects in Dark Energy Survey}
 Astronomical data sets are large and growing. Large modern optical
@@
 $5000~{\rm deg}^2$ of the southern sky from the Cerro-Tololo
-Inter-American Observatory in Chile~\cite{zuntz2018dark}. The resulting galaxy catalogs produced have provided some of the strongest constraints to date on
+Inter-American Observatory in Chile~\cite{zuntz2018dark}. The resulting galaxy 
+catalogs produced have provided some of the strongest constraints to date on
 the physical properties of dark energy and accelerated expansion
@@
 observed by the Dark Energy Survey (DES) sampled from the initial June 2018 release.
-We labeled the 25,339 objects from this 100K set that did not appear in the later December 2019
-release, thus were likely eliminated during the manual vetting process, 
-as outliers.
-While the remaining 74,661 objects may also contain outliers, we assume them to be inliers in this experiment.
-We used publicly-available photometry from the $g-$, $r-$, $i-$ and $z-$band DES exposures.
-We
+We labeled the 25,339 objects from this 100K set that did not appear in the 
+later December 2019 release, thus were likely eliminated during the manual 
+vetting process, as outliers.
+While the remaining 74,661 objects may also contain outliers, we assume them 
+to be inliers in this experiment. We used publicly-available photometry from 
+the $g-$, $r-$, $i-$ and $z-$band DES exposures.  We
 transformed the photometry into luptitudes\footnote{A
@@
 were the $r$-band luptitude, colors computed as band differences between
-$g-r$, $i-r$, and $z-r$, and associated observational errors, for a total of eight features.
+$g-r$, $i-r$, and $z-r$, and associated observational errors, for a total of 
+eight features.
 %
 
-\subsubsection{Planetary: targets in Mars rover images.}
+\subsection{Planetary: targets in Mars rover images}
 Mars exploration is fundamentally an exercise in discovery with the
@@
 
-\subsubsection{Earth: satellite time series for ground observations.}
+\subsection{Earth: satellite time series for ground observations}
 Many Earth science applications using satellite Earth observation (EO) data
@@
 learning models. For example, a model trained to classify crop types in
- satellite observations requires ground-annotated labels of crop type. A widespread challenge for ground-
+satellite observations requires ground-annotated labels of crop type. 
+A widespread challenge for ground-
 annotated labels is that there are often points with erroneous location or 
@@
 
-\subsubsection{Benchmark: MNIST and Fashion-MNIST.}
+\subsection{Benchmark: MNIST and Fashion-MNIST}
 We used MNIST and F-MNIST to demonstrate DORA with a
 traditional benchmark dataset. We used 60,000 images from 
-F-MNIST as the training set and a test set of 1,000 images each from MNIST (outliers) and F-MNIST (inliers). 
+F-MNIST as the training set and a test set of 1,000 images 
+each from MNIST (outliers) and F-MNIST (inliers). 
 %Well-discriminating algorithms should assign 
@@
 \section{Results}
-\begin{figure}[h]
+
+% TODO: Tried to make these a 2x2 panel of subfigures, but was getting
+% errors using the subfigure package (Error: 'Not in outer par mode'). It seems
+% this option may not be available for Frontiers articles 
+% (see bottom of document). May combine them as a single figure manually.
+
+\begin{figure}
 \centering
-    \includegraphics[width=\linewidth]{figures/planetary_combined_plot.pdf}
+\includegraphics[width=0.6\linewidth]{figures/planetary_combined_plot.pdf}
 \caption{Number of known outliers ranked in top $N$ selections for 
@@
 
-\begin{figure}[h]
+\begin{figure}
 \centering
-    \includegraphics[width=\linewidth]{figures/faw_combined_plot.pdf}
+\includegraphics[width=0.6\linewidth]{figures/faw_combined_plot.pdf}
 \caption{Number of known outliers ranked in top $N$ selections for 
@@
 
-\begin{figure}[h]
+\begin{figure}
 \centering
-    \includegraphics[width=\linewidth]{figures/des_combined_plot.pdf}
+\includegraphics[width=0.6\linewidth]{figures/des_combined_plot.pdf}
 \caption{Number of known outliers ranked in top $N$ selections for 
@@
 \label{fig:des_results}
+\end{figure}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=0.6\linewidth]{figures/fmnist_combined_plot.pdf}
+    \caption{Number of known outliers ranked in top $N$ selections for 
+    F-MNIST dataset.}
+    \label{fig:fmnist_results}
 \end{figure}
@@
 precision at $N=n_{outlier}$
-while DEMUD, PCA, and PAE tied for the highest MDR. Negative sampling had the 
+while DEMUD, PCA, and PAE tied for the highest MDR 
+(Figure \ref{fig:fmnist_results}). 
+Negative sampling had the 
 lowest performance in both metrics.
@@
 
-\begin{figure}
-    \centering
-    \includegraphics[width=\linewidth]{figures/fmnist_combined_plot.pdf}
-    \caption{Number of known outliers ranked in top $N$ selections for 
-    F-MNIST dataset.}
-    \label{fig:fmnist_results}
-\end{figure}
-
-\paragraph{Algorithm performance.} 
+\subsection{Algorithm performance} 
 No one algorithm  
@@
 
-\paragraph{Evaluation in outlier detection.} 
+\subsection{Evaluation in outlier detection} 
 Prior work has emphasized
@@
 selections is equivalent to the number of outliers, and
- Mean Discovery Rate, which measures the fraction of selections that are known outliers on average. 
+Mean Discovery Rate, which measures the fraction of selections that are known 
+outliers on average. 
 Designing experiments to evaluate outlier detection methods for real-world
@@
 
-\paragraph{Path to deployment.} 
+% This section was specific to IAAI, can change the name, remove, or keep
+% as is if we want.
+\subsection{Path to deployment} 
 Our goal is for DORA to enable increased application and benefit of
@@
 three use cases we demonstrated results for in this study.
+\todo{Add link to github since open now.}
 %the use cases from this will also lead to infusion in real scenarios
@@
 
-%\section{Acknowledgments}
-%We thank the NASA Center Innovation Fund and the NASA Science Mission 
-%Directorate (under the NASA Cross-Divisional AI Use Case Demonstration grants)
-% for supporting this work.
-%
-%We thank the UN Food and Agriculture Organization (FAO) and Dr. Ritvik Sahajpal
-%(University of Maryland/NASA Harvest) for providing the maize ground
-% observation dataset.
-% 
-% We thank Raymond Francis for analyzing the Navcam rover images to
-%identify targets considered to be ``novel'' for evaluating outlier
-%detection in the planetary rover scenario.
-%
-%Part of this research was carried out at the Jet Propulsion
-%Laboratory, California Institute of Technology, under a contract with
-%the National Aeronautics and Space Administration.
+\section*{Conflict of Interest Statement}
+%All financial, commercial or other relationships that might be perceived by the academic community as representing a potential conflict of interest must be disclosed.
+If no such relationship exists, authors will be asked to confirm the following statement:
 
+The authors declare that the research was conducted in the absence of any commercial or financial relationships that could be construed as a potential conflict of interest.
+
+\section*{Author Contributions}
+
+The Author Contributions section is mandatory for all articles, including articles by sole authors.
+If an appropriate statement is not provided on submission, a standard one will be inserted during the production process.
+The Author Contributions statement must describe the contributions of individual authors referred to by their initials and, in doing so, all authors agree to be accountable for the content of the work.
+Please see  \href{http://home.frontiersin.org/about/author-guidelines#AuthorandContributors}{here} for full authorship criteria.
+\todo{Fill this in.}
+
+\section*{Funding}
+We thank the NASA Center Innovation Fund and the NASA Science Mission 
+Directorate (under the NASA Cross-Divisional AI Use Case Demonstration grants)
+ for supporting this work. \todo{Add grant numbers if possible.}
+
+\section*{Acknowledgments}
+We thank the UN Food and Agriculture Organization (FAO) and Dr. Ritvik Sahajpal
+(University of Maryland/NASA Harvest) for providing the maize ground
+ observation dataset.
+
+ We thank Raymond Francis for analyzing the Navcam rover images to
+identify targets considered to be ``novel'' for evaluating outlier
+detection in the planetary rover scenario.
+
+Part of this research was carried out at the Jet Propulsion
+Laboratory, California Institute of Technology, under a contract with
+the National Aeronautics and Space Administration.
+
+%\subsection{Tables}
+%Tables should be inserted at the end of the manuscript.
+Please build your table directly in LaTeX.Tables provided as jpeg/tiff files will not be accepted.
+Please note that very large tables (covering several pages) cannot be included in the final PDF for reasons of space.
+These tables will be published as \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} on the online article page at the time of acceptance.
+The author will be notified during the typesetting of the final article if this is the case.
+
+%\section*{Supplemental Data}
+% \href{http://home.frontiersin.org/about/author-guidelines#SupplementaryMaterial}{Supplementary Material} should be uploaded separately on submission, if there are Supplementary Figures, please include the caption in the same file as the figure.
+LaTeX Supplementary Material templates can be found in the Frontiers LaTeX folder.
+
+\section*{Data Availability Statement}
+The datasets [GENERATED/ANALYZED] for this study can be found in the [NAME OF REPOSITORY] [LINK]. 
+\todo{Can we make our datasets available here? We can at least add the github link to the code.}
+% Please see the availability of data guidelines for more information, at https://www.frontiersin.org/about/author-guidelines#AvailabilityofData
+
+%\bibliographystyle{frontiersinSCNS_ENG_HUMS} % for Science, Engineering and Humanities and Social Sciences articles, for Humanities and Social Sciences articles please include page numbers in the in-text citations
+\bibliographystyle{frontiersinHLTH&FPHY} % for Health, Physics and Mathematics articles
 \bibliography{dora_references}
 
+%%% Make sure to upload the bib file along with the tex file and PDF
+%%% Please see the test.bib file for some examples of references
+
+%\section*{Figure captions}
+
+%%% Please be aware that for original research articles we only permit a combined number of 15 figures and tables, one figure with multiple subfigures will count as only one figure.
+%%% Use this if adding the figures directly in the mansucript, if so, please remember to also upload the files when submitting your article
+%%% There is no need for adding the file termination, as long as you indicate where the file is saved. In the examples below the files (logo1.eps and logos.eps) are in the Frontiers LaTeX folder
+%%% If using *.tif files convert them to .jpg or .png
+%%%  NB logo1.eps is required in the path in order to correctly compile front page header %%%
+
+%\begin{figure}[h!]
+%\begin{center}
+%\includegraphics[width=10cm]{logo1}% This is a *.eps file
+%\end{center}
+%\caption{ Enter the caption for your figure here.  Repeat as  necessary for each of your figures}\label{fig:1}
+%\end{figure}
+%
+%
+%\begin{figure}[h!]
+%\begin{center}
+%\includegraphics[width=15cm]{logos}
+%\end{center}
+%\caption{This is a figure with sub figures, \textbf{(A)} is one logo, \textbf{(B)} is a different logo.}\label{fig:2}
+%\end{figure}
+
+%%% If you are submitting a figure with subfigures please combine these into one image file with part labels integrated.
+%%% If you don't add the figures in the LaTeX files, please upload them when submitting the article.
+%%% Frontiers will add the figures at the end of the provisional pdf automatically
+%%% The use of LaTeX coding to draw Diagrams/Figures/Structures should be avoided. They should be external callouts including graphics.
+
 \end{document}
-
-% LocalWords:  tex OOD Eigenbasis DEMUD PCA Xiaoli RX autoencoder al
-% LocalWords:  Pimentel mislabeled Liang Böhm wagstaff demud Xiong
-% LocalWords:  Storey Hayat des anom Lochner Chiang Zhou Liu Alvera
-% LocalWords:  Azcárate Touati Prochaska Kerner kerner PyOD zhao pyod
-% LocalWords:  API Umaa Vinay Navcam MNIST FashionMNIST installable
-% LocalWords:  dora CIF SMD FAO Pritchard

### 2022-01-11 publications/frontiers/dora_outlier_detection_frontiers.tex
Fill in todos. #79

@@
 \usepackage{amsmath}
+\usepackage{booktabs}
 \usepackage[dvipsnames]{xcolor}
@@
 % If one of the authors has a change of address, list the new address below the correspondence details using a superscript symbol and use the same symbol to indicate the author in the author list.
-\def\Address{$^{1}$Department of Geographical Sciences, University of Maryland, College Park, MD, USA \\
+\def\Address{$^{1}$University of Maryland, College Park, MD, USA \\
 $^{2}$Jet Propulsion Laboratory, California Institute of Technology, Pasadena, California, USA \\
@@
 \begin{tabular}{l|c|cc}
-  \hline
+  \toprule
   Dataset & Training & \multicolumn{2}{c}{Test}\\
@@
 
-  \hline
+  \midrule
   Planetary & 992 &  9 & 49 \\
@@
   Astrophysics & 100,000   &  25,339 &  74,661 \\
-  \hline
+  \bottomrule
 \end{tabular}
@@
 \begin{tabular}{l|cccc}
-  \hline
+  \toprule
   Algorithm & Astro. & Earth & Planetary & F-MNIST \\
-  \hline
+  \midrule
   PCA               & \textbf{0.42} & 0.41 & 0.44 & \textbf{0.84} \\
   DEMUD             & ---- & 0.41 & 0.44 & \textbf{0.84} \\
-  \hline
+  \midrule
   RX                & 0.40 & 0.43 & ---- & 0.82 \\
   LRX               & ----& ---- & 0.33 & 0.56 \\
-  \hline
+  \midrule
   Iso. Forest  & 0.34 & 0.46 & \textbf{0.56} & 0.74 \\
-  \hline
+  \midrule
   PAE               & 0.35 & 0.41 & 0.44 & 0.83 \\
   Neg. Sampling & 0.32 & \textbf{0.49} & 0.33 & 0.43 \\
-  \hline
+  \midrule
   Random            & 0.25 & 0.32 & 0.14 & 0.50 \\
-  \hline
+  \bottomrule
 \end{tabular}
@@
 three use cases we demonstrated results for in this study.
-\todo{Add link to github since open now.}
+The DORA code is openly accessible via Github at
+\url{https://github.com/nasaharvest/dora}.
 %the use cases from this will also lead to infusion in real scenarios
@@
 \section*{Author Contributions}
-
-The Author Contributions section is mandatory for all articles, including articles by sole authors.
-If an appropriate statement is not provided on submission, a standard one will be inserted during the production process.
-The Author Contributions statement must describe the contributions of individual authors referred to by their initials and, in doing so, all authors agree to be accountable for the content of the work.
-Please see  \href{http://home.frontiersin.org/about/author-guidelines#AuthorandContributors}{here} for full authorship criteria.
-\todo{Fill this in.}
+HRK, UR, and KLW developed the conception of the project.
+HK oversaw the overall implementation and led the analysis of the Earth science dataset.
+UR led the analysis of the astrophysics dataset.
+KLW led the analysis of the planetary science dataset.
+SL led the development of the DORA software architecture.
+BD led the analysis of the FMNIST/MNIST dataset.
+EH supported the analysis of the astrophysics dataset.
+JL developed the desktop application for results visualization.
+VR supported the analysis of the Earth science dataset.
+SK supported the analysis of the planetary science dataset.
+All authors contributed to the overall system implementation and manuscript writing/editing.
 
 \section*{Funding}
-We thank the NASA Center Innovation Fund and the NASA Science Mission 
-Directorate (under the NASA Cross-Divisional AI Use Case Demonstration grants)
- for supporting this work. \todo{Add grant numbers if possible.}
+We thank the NASA Center Innovation Fund (\todo{Award number?})
+and the NASA Science Mission Directorate 
+(under the NASA Cross-Divisional AI Use Case Demonstration grants;
+Award \#80NSSC21K0720) for supporting this work.
 
@@
  observation dataset. 
-
- We thank Raymond Francis for analyzing the Navcam rover images to
-identify targets considered to be ``novel'' for evaluating outlier
-detection in the planetary rover scenario.
-
+%
+We thank Raymond Francis for analyzing the Navcam 
+rover images to identify targets considered to be ``novel'' for evaluating
+outlier detection in the planetary rover scenario.
+%
 Part of this research was carried out at the Jet Propulsion
@@
 The datasets [GENERATED/ANALYZED] for this study can be found in the [NAME OF REPOSITORY] [LINK]. 
-\todo{Can we make our datasets available here? We can at least add the github link to the code.}
 % Please see the availability of data guidelines for more information, at https://www.frontiersin.org/about/author-guidelines#AvailabilityofData

### 2022-01-22 publications/frontiers/dora_outlier_detection_frontiers.tex
Suggested minor wording updates. (#79)

@@
 (DEMUD)~\citep{wagstaff:demud13}, principal component analysis (PCA),
-Reed-Xiaoli (RX) detector,  
+Reed-Xiaoli (RX) detector~\citep{reed1990adaptive},  
 Local RX, negative sampling~\cite{sipple:neg-sampling20}, and probabilistic 
@@
 limited data formats and algorithms. DORA aims to fill the need 
-for tools that faciliate application, evaluation, and interpretation of outlier
+for tools that facilitate application, evaluation, and interpretation of outlier
 detection methods.
@@
 \subsection{Outlier ranking algorithms} 
-We implemented 7 unsupervised 
+We implemented seven unsupervised 
 algorithms for scoring and 
@@
 the Mahalanobis
-distance between a sample and a background mean and covariance
-~\citep{reed1990adaptive}. The local variant of RX (Local
+distance between a sample and a background mean and covariance~\citep{reed1990adaptive}. The local variant of RX (Local
 RX or LRX) can be used for image or raster data and scores each pixel in
@@
 a sample is quantified as the average distance from the root to the item’s 
-leaf. Shorter distances are indicative of outliers because fewer splits were 
-required to isolate the sample.
+leaf. Shorter distances are indicative of outliers because the number
+of random splits required to isolate the sample is small.
 
@@
 \subsection{Benchmark: MNIST and Fashion-MNIST}
-We used MNIST and F-MNIST to demonstrate DORA with a
+We used MNIST and Fashion-MNIST (F-MNIST) to demonstrate DORA with a
 traditional benchmark dataset. We used 60,000 images from 
@@
 HRK, UR, and KLW developed the conception of the project.
-HK oversaw the overall implementation and led the analysis of the Earth science dataset.
+HRK oversaw the overall implementation and led the analysis of the Earth science dataset.
 UR led the analysis of the astrophysics dataset.
@@
 \section*{Funding}
-We thank the NASA Center Innovation Fund (\todo{Award number?})
+We thank the NASA Center Innovation Fund 
 and the NASA Science Mission Directorate 
@@
 %%% Please be aware that for original research articles we only permit a combined number of 15 figures and tables, one figure with multiple subfigures will count as only one figure.
-%%% Use this if adding the figures directly in the mansucript, if so, please remember to also upload the files when submitting your article
+%%% Use this if adding the figures directly in the manuscript, if so, please remember to also upload the files when submitting your article
 %%% There is no need for adding the file termination, as long as you indicate where the file is saved. In the examples below the files (logo1.eps and logos.eps) are in the Frontiers LaTeX folder
@@
 \end{document}
+
+% LocalWords:  cls

### 2022-01-26 publications/frontiers/dora_outlier_detection_frontiers.tex
Standardize dataset ordering and combine results figs. #79

@@
 the Mahalanobis
-distance between a sample and a background mean and covariance~\citep{reed1990adaptive}. The local variant of RX (Local
+distance between a sample and a background mean and 
+covariance~\citep{reed1990adaptive}. The local variant of RX (Local
 RX or LRX) can be used for image or raster data and scores each pixel in
@@
 \subsubsection{Sparsity}
-Isolation forests are a common sparsity-based method
+Isolation forest (iForest) is a common sparsity-based method
 that constructs many random binary trees from a 
@@
 \begin{table}
-\caption{Number of samples in the training and test sets for each dataset,
-ordered by dataset size.}
+\caption{Number of samples in the training and test sets for each dataset.}
 \label{tab:datasets}
@@
    & Unlabeled &  Outliers &  Inliers \\
-
   \midrule
+  Astrophysics & 100,000   &  25,339 &  74,661 \\
   Planetary & 992 &  9 & 49 \\
@@
   F-MNIST & 60,000 & 1,000 & 1,000  \\
-  Astrophysics & 100,000   &  25,339 &  74,661 \\
   \bottomrule
@@
   \toprule
-  Algorithm & Astro. & Earth & Planetary & F-MNIST \\
+  Algorithm & Astro. & Planetary & Earth & F-MNIST \\
   \midrule
-  PCA               & \textbf{0.42} & 0.41 & 0.44 & \textbf{0.84} \\
-  DEMUD             & ---- & 0.41 & 0.44 & \textbf{0.84} \\
+  PCA               & \textbf{0.42} & 0.44 & 0.41 & \textbf{0.84} \\
+  DEMUD             & --- & 0.44 & 0.41 & \textbf{0.84} \\
   \midrule
-  RX                & 0.40 & 0.43 & ---- & 0.82 \\
-  LRX               & ----& ---- & 0.33 & 0.56 \\
+  RX                & 0.40 & --- & 0.43 & 0.82 \\
+  LRX               & --- & 0.33 & --- & 0.56 \\
   \midrule
-  Iso. Forest  & 0.34 & 0.46 & \textbf{0.56} & 0.74 \\
+  iForest  & 0.34 & \textbf{0.56} & 0.46 & 0.74 \\
   \midrule
-  PAE               & 0.35 & 0.41 & 0.44 & 0.83 \\
-  Neg. Sampling & 0.32 & \textbf{0.49} & 0.33 & 0.43 \\
+  PAE               & 0.35 & 0.44 & 0.41 & 0.83 \\
+  Neg. Sampling & 0.32 & 0.33 & \textbf{0.49} & 0.43 \\
   \midrule
-  Random            & 0.25 & 0.32 & 0.14 & 0.50 \\
+  Random            & 0.25 & 0.14 & 0.32 & 0.50 \\
   \bottomrule
@@
 
-% TODO: Tried to make these a 2x2 panel of subfigures, but was getting
+% Tried to make these a 2x2 panel of subfigures, but was getting
 % errors using the subfigure package (Error: 'Not in outer par mode'). It seems
@@
 \centering
-\includegraphics[width=0.6\linewidth]{figures/planetary_combined_plot.pdf}
+\includegraphics[width=\linewidth]{figures/combined_results.pdf}
 \caption{Number of known outliers ranked in top $N$ selections for 
-planetary rover dataset.}
-\label{fig:planetary_results}
+the (a) astrophysics, (b) planetary, (c) Earth, and (d) FMNIST datasets.}
+\label{fig:combined_results}
 \end{figure}
 
-\begin{figure}
-\centering
-\includegraphics[width=0.6\linewidth]{figures/faw_combined_plot.pdf}
-\caption{Number of known outliers ranked in top $N$ selections for 
-Earth science dataset.}
-\label{fig:faw_results}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics[width=0.6\linewidth]{figures/des_combined_plot.pdf}
-\caption{Number of known outliers ranked in top $N$ selections for 
-the astrophysics (DES) dataset.}
-\label{fig:des_results}
-\end{figure}
-
-\begin{figure}
-    \centering
-    \includegraphics[width=0.6\linewidth]{figures/fmnist_combined_plot.pdf}
-    \caption{Number of known outliers ranked in top $N$ selections for 
-    F-MNIST dataset.}
-    \label{fig:fmnist_results}
-\end{figure}
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.6\linewidth]{figures/des_combined_plot.pdf}
+%\caption{Number of known outliers ranked in top $N$ selections for 
+%the astrophysics (DES) dataset.}
+%\label{fig:des_results}
+%\end{figure}
+%
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.6\linewidth]{figures/planetary_combined_plot.pdf}
+%\caption{Number of known outliers ranked in top $N$ selections for 
+%planetary rover dataset.}
+%\label{fig:planetary_results}
+%\end{figure}
+%
+%\begin{figure}
+%\centering
+%\includegraphics[width=0.6\linewidth]{figures/faw_combined_plot.pdf}
+%\caption{Number of known outliers ranked in top $N$ selections for 
+%Earth science dataset.}
+%\label{fig:faw_results}
+%\end{figure}
+%
+%\begin{figure}
+%    \centering
+%    \includegraphics[width=0.6\linewidth]{figures/fmnist_combined_plot.pdf}
+%    \caption{Number of known outliers ranked in top $N$ selections for 
+%    F-MNIST dataset.}
+%    \label{fig:fmnist_results}
+%\end{figure}
 
@@
 reported the Mean Discovery Rate (MDR) in the legend
-for each algorithm to give a quantitative comparison across the datasets. We defined MDR as:
+for each algorithm to give a quantitative comparison across the datasets. 
+We defined MDR as:
 \begin{equation}
@@
 
-%\todo{Hannah: Consider merging all three data set discussions since they are
-%in one table? with their own figures? The point here isn't really to
-%compare algs or data sets but instead to show that we CAN compare
-%them, using DORA.}
-% We initialized each algorithm with $D_s$ and requested a ranking of
-% targets in $D_{s+}$.
-%
+For the astrophysics dataset (Figure~\ref{fig:combined_results}a), 
+DEMUD was omitted due to computational time 
+and LRX was omitted as it applies only to image data.  
+Of the remaining methods, 
+PCA achieved the highest precision, followed by RX.  Negative sampling 
+performs well initially before its performance drops off.  
+The PAE finds the most outliers overall.
 
@@
 detection) when allowed to select only \num{9} images.
-Figure~\ref{fig:planetary_results} shows the complete (cumulative)
+Figure~\ref{fig:combined_results}b shows the complete (cumulative)
 outlier detection performance for each algorithm when ranking
@@
 $N=n_{outlier}$ while DEMUD and PCA tied for the lowest
-MDR (Figure \ref{fig:faw_results}). We did not evaluate LRX for this time 
+MDR (Figure \ref{fig:combined_results}c). We did not evaluate LRX for this time 
 series dataset because LRX can only be applied to gridded image or raster 
 data types.
-
-For the astrophysics dataset (Figure~\ref{fig:des_results}), DEMUD was omitted due to computational time 
-and LRX was omitted as it applies only to image data.  Of the remaining methods, 
-PCA achieved the highest precision, followed by RX.  Negative sampling 
-performs well initially before its performance drops off.  The PAE finds the most outliers 
-overall.
 
@@
 while DEMUD, PCA, and PAE tied for the highest MDR 
-(Figure \ref{fig:fmnist_results}). 
+(Figure \ref{fig:combined_results}d). 
 Negative sampling had the 

### 2022-02-01 publications/frontiers/dora_outlier_detection_frontiers.tex
Add link to dataset on Zenodo.

@@
 \section*{Data Availability Statement}
-The datasets [GENERATED/ANALYZED] for this study can be found in the [NAME OF REPOSITORY] [LINK]. 
+The datasets generated and analyzed for this study can be found in the
+``Multi-Domain Outlier Detection Dataset'' repository on Zenodo 
+(\url{https://doi.org/10.5281/zenodo.5941339}) (dataset will be made public
+with final publication). 
 % Please see the availability of data guidelines for more information, at https://www.frontiersin.org/about/author-guidelines#AvailabilityofData

### 2022-03-29 publications/frontiers/dora_outlier_detection_frontiers.tex
Update Zenodo link to DOI for all versions. #79

--- 2022-01-26 publications/frontiers/dora_outlier_detection_frontiers.tex
@@
 \section*{Data Availability Statement}
-The datasets [GENERATED/ANALYZED] for this study can be found in the [NAME OF REPOSITORY] [LINK]. 
+The datasets generated and analyzed for this study can be found in the
+``Multi-Domain Outlier Detection Dataset'' repository on Zenodo 
+(\url{https://doi.org/10.5281/zenodo.5941338}). 
 % Please see the availability of data guidelines for more information, at https://www.frontiersin.org/about/author-guidelines#AvailabilityofData

### 2022-03-30 publications/frontiers/dora_outlier_detection_frontiers.tex
Minor revisions from reviewers. #79

@@
 In machine learning, these methods are commonly used for identifying 
-mislabeled or otherwise invalid samples in a dataset. When working with 
-science datasets, OOD detection can be used for 
+mislabeled or otherwise invalid samples in a dataset 
+~\citep{liang2018enhancing,bohm2020probabilistic}. 
+%
+When working with science datasets, OOD detection can be used for 
 cleaning datasets, e.g., flagging ground-truth labels with GPS or human
-entry error or identifying wrongly categorized objects in a catalog. 
+entry error or identifying wrongly categorized objects in a catalog
+~\citep{wagstaff:des-anom20,lochner2021astronomaly}. 
 It could also be used for discovery, e.g., to flag novel samples in order 
-to guide instrument acquisition or scientific analysis. Another application
+to guide instrument acquisition or scientific analysis
+~\citep{wagstaff:demud13,wagstaff:rover-novelty20,
+kerner2020comparison,kerner2020analysis}.
+Another application
 is the detection of rare objects that are known to exist but the known
 examples are too few to create a large enough labeled dataset for 
-supervised classification algorithms.
+supervised classification algorithms~\citep{chang2002anomaly,zhou2016novel}.
 
@@
 science, and astrophysics, as well as one benchmark dataset 
-(MNIST/Fashion-MNIST). We found that no single algorithm
-performed best across all datasets, underscoring the need for a tool 
-that enables comparison of multiple algorithms.
+(MNIST/Fashion-MNIST). 
+%We found that no single algorithm
+%performed best across all datasets, underscoring the need for a tool 
+%that enables comparison of multiple algorithms.
+
+The key contributions of this paper are:
+\begin{itemize}
+\item A new pipeline, DORA, for performing outlier detection experiments using 
+several AI algorithms that reduces the effort and expertise required for 
+performing experiments and comparing results from multiple algorithms
+\item Using experiments for a diverse set of real world datasets and 
+application areas, we show that no single algorithm performs best for all 
+datasets and use cases, underscoring the need for a tool that compares 
+multiple algorithms
+\item We provide publicly available code for running and contributing to the 
+DORA pipeline and datasets that can be used for reproducing experiments or 
+benchmarking outlier detection methods 
+\end{itemize}
 
@@
 as well as data or modeling 
-artifacts in astronomical surveys~\citep{wagstaff:des-anom20}. 
+artifacts in astronomical 
+surveys~\citep{wagstaff:des-anom20,lochner2021astronomaly}. 
 Example outlier detection applications in 
@@
 
+\subsection{Astrophysics: objects in Dark Energy Survey}
+Astronomical data sets are large and growing. Large modern optical
+imaging surveys are producing catalogs of order $10^8$ stars
+and galaxies, with dozens or hundreds of distinct measured features
+for each entry. Discovery science becomes difficult at this data
+volume: the scale is too large for expert human inspection, and separating real
+astrophysical anomalies from non-astrophysical sources like detector artifacts or
+satellite trails is a challenging problem for current methods.
+
 \begin{table}
@@
 \end{table}
-
-\subsection{Astrophysics: objects in Dark Energy Survey}
-Astronomical data sets are large and growing. Large modern optical
-imaging surveys are producing catalogs of order $10^8$ stars
-and galaxies, with dozens or hundreds of distinct measured features
-for each entry. Discovery science becomes difficult at this data
-volume: the scale is too large for expert human inspection, and separating real
-astrophysical anomalies from non-astrophysical sources like detector artifacts or
-satellite trails is a challenging problem for current methods.
 
@@
 % as is if we want.
-\subsection{Path to deployment} 
+\subsection{Open code and data} 
 Our goal is for DORA to enable increased application and benefit of
@@
 designed DORA to make it as easy as possible for scientists to apply
-algorithms and to compare and interpret their results; users need only
+algorithms and to compare and interpret their results. Users need only
 to specify the specifics of their data (e.g., path, data type) in a 
-configuration file to start running experiments and seeing results. We will
-make DORA publicly available as a pip-installable package to make it
+configuration file to start running experiments and seeing results for 
+their own datasets and use cases. DORA is
+publicly available and can be installed using pip via Github, making it
 easy to integrate into existing scientific workflows. 
-The source code and example datasets used in this study will also be 
-open-sourced, enabling DORA to be improved and expanded by the 
+The datasets used in this study are also publicly available via Zenodo. 
+This enables DORA to be improved and expanded by the 
 machine learning and domain science communities. 
+If a researcher wants to use DORA for a dataset with a type that is not yet
+supported, they can contribute a new data loader by creating a subclass
+that extends the \texttt{DataLoader} abstract base class. Similarly, new 
+results interpretation modules can be added by creating a subclass of the 
+\texttt{ResultsOrganization} abstract base class. A new outlier ranking
+algorithm can be added by writing a new python module that defines
+a subclass of the \texttt{OutlierDetection} abstract base class and implements 
+the required functions for scoring and ranking samples, following the 
+existing algorithm modules named \texttt{*\_outlier\_detection.py}.
 In addition, DORA will be infused into the scientific workflows for the
 three use cases we demonstrated results for in this study.
-The DORA code is openly accessible via Github at
-\url{https://github.com/nasaharvest/dora}.
+The DORA code can be accessed at
+\url{https://github.com/nasaharvest/dora} and datasets 
+at \url{https://doi.org/10.5281/zenodo.5941338}.
 %the use cases from this will also lead to infusion in real scenarios

### 2022-03-31 publications/frontiers/reviewer-responses.md
Add reviewer comments and responses. #79

--------------------------------------------------------------
 01 March 2022
--------------------------------------------------------------

Dear Dr. Kerner,

I am attaching the reports of the two reviewers below. Please make the necessary corrections/changes as suggested in the reports and submit a revised manuscript.

Please do not hesitate to contact me if you have any questions.

Sincerely,
Bala Poduval
-------------------------

Reviewer 2:

1) Please provide references supporting these arguments "OOD detection can be used for cleaning datasets, e.g., flagging ground-truth labels with GPS or human entry error or identifying wrongly categorized objects in a catalog. It could also be used for discovery, e.g., to flag novel samples in order to guide instrument acquisition or scientific analysis. Another application is the detection of rare objects that are known to exist but the known examples are too few to create a large enough labeled dataset for supervised classification algorithms."

2) Please provide the contributions of the paper in a highlighted manner. For example, you can use bullets for listing down the contribution. Ex: 1) a tool for applying outlier detection algorithms to various scientific data, 2) outlier score to each sample in a given dataset using interactive visualizations, etc.

3) Can you add a short, high-level description of how to integrate more datasets and algorithms into your framework. It will be an important section as it allows researchers to utilize your framework for their specific domains. You can add a more detailed description in the git repository.

Reviewer 3:

This article is particularly well written and very pleasant to read. The proposed python package, called DORA and already available to the scientific community, allowing the detection of outliers via many approaches already available in the literature, is of great interest for users. As mentioned by the authors and observed on the 4 datasets considered, obviously no method dominates the others, each one having its own numerical qualities in the detection of outliers.

There is just one clarification that I missed when reading the article: it concerns the outlier score. Has this been standardized so that it is comparable for all methods, or does each method remain on its own score scale (the latter potentially changing depending on the data set considered)? And how, if at all?
----------------

--------------------------------------------------------------
 30 March 2022
--------------------------------------------------------------

Dear Dr. Poduval and anonymous reviewers,

Thank you for your review of our paper and for your helpful comments and suggestions. We have addressed them all and I am attaching the revised manuscript here. Below are responses to individual reviewer comments:

Reviewer 2

1) Please provide references supporting these arguments "OOD detection can be used for cleaning datasets, e.g., flagging ground-truth labels with GPS or human entry error or identifying wrongly categorized objects in a catalog. It could also be used for discovery, e.g., to flag novel samples in order to guide instrument acquisition or scientific analysis. Another application is the detection of rare objects that are known to exist but the known examples are too few to create a large enough labeled dataset for supervised classification algorithms."

We have added references to multiple relevant papers (lines 26-33).

2) Please provide the contributions of the paper in a highlighted manner. For example, you can use bullets for listing down the contribution. Ex: 1) a tool for applying outlier detection algorithms to various scientific data, 2) outlier score to each sample in a given dataset using interactive visualizations, etc.

Thank you for the suggestion. We have added a bulleted list of the key contributions at the end of the Introduction on lines 53-61.

3) Can you add a short, high-level description of how to integrate more datasets and algorithms into your framework. It will be an important section as it allows researchers to utilize your framework for their specific domains. You can add a more detailed description in the git repository.

Thank you for the suggestion. We have added a description of how to add new data loaders, outlier ranking algorithms, and results interpretation modules to DORA as well as apply DORA to other datasets (lines 275-292).

Reviewer 3

This article is particularly well written and very pleasant to read. The proposed python package, called DORA and already available to the scientific community, allowing the detection of outliers via many approaches already available in the literature, is of great interest for users. As mentioned by the authors and observed on the 4 datasets considered, obviously no method dominates the others, each one having its own numerical qualities in the detection of outliers.

1) There is just one clarification that I missed when reading the article: it concerns the outlier score. Has this been standardized so that it is comparable for all methods, or does each method remain on its own score scale (the latter potentially changing depending on the data set considered)? And how, if at all?

Thank you for your positive feedback. The outlier scores are not standardized; we use the scores that are output by the algorithms which have different scales across the algorithms and datasets. To inter-compare the algorithms directly, we compare the relative ranking order rather than the absolute scores of each dataset in our experiments and the results modules in DORA.
